### 2019-08-22 ICES poster/ICES poster.Rmd
poster creation file

---
title: "**Short title:** Flexdashboard poster template"
output: 
  postr::flex_dashboard_poster:
    render_print: false
---

## Header row {data-height=15}

###

<table style="width:100%">
<tr>
  <td><center><h3>
  **Poster title**
  </h3></center></td>
  <td rowspan="2">
  <img src="resources/logo.png" style="width:128px"></img>
  </td>
</tr>
<tr><td><center>
  **Author**
  </center></td>
</tr>
</table>

<hr>

**Abstract** This paragraph will be rendered inside a wide box right below the
poster title and authors information.
The `render_print` format option is set to `TRUE` by default in this poster
sample, and will result in `postr::render` being called when knitting the
document (see `help("flex_dashboard_poster", package = postr)`).
To disable this (recommended in the initial stages of poster development)
set the `render_print` option to `FALSE`.

## Body row {data-height=75}

### **Column 1 title**

#### Heading

This section defines the first column of the body row of the poster.
You can include regular markdown text and R `code chunks`.

```{r}
hist(volcano)
```

### **Column 2 title**

Content for the second column.

To show the code in the chunks you have to ask for it with `echo=TRUE`.

```{r volcano-heat, echo=TRUE, eval=FALSE}
filled.contour(volcano, color.palette = terrain.colors, asp = 1)
```

You can also include data frames as tables just by printing them

```{r}
head(iris)
```

### **Column 3 title**

This template shows 3 columns in the body row, you can change this by adding
or removing third level headings (i.e. lines starting with ###).

## Footer row {data-height=10}

### {data-width=50}

### {data-width=50 .small}

### 2019-08-22 Figures.R
Make participant plot

@@ -389,7 +389,9 @@ freq.data_CC<-mutate(freq.data_CC,Analysis="Climate Change")
 freq.data<-rbind(freq.data_pub,freq.data_CC)
 part.data_pub<-mutate(part.data_pub,Analysis="Publication")
 part.data_CC<-mutate(part.data_CC,Analysis="Climate Change")
-part.data<-rbind(part.data_pub,part.data_CC)
+part.data<-rbind(part.data_pub,part.data_CC) %>% 
+  arrange(Stage, Percent) %>% 
+  mutate(order = row_number())
 
   
 ##### Data Analysis Outputs #####
@@ -453,8 +455,9 @@ Part.plot<-ggplot(part.data,aes(x=order,y=Percent,fill=Analysis)) +
   scale_x_continuous(breaks = part.data$order,
                      labels = part.data$Participants)+
   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
-  ylab("Percent")+coord_flip() + 
-  guides(fill = guide_legend(reverse=T)) + theme_bw()
+  ylab("Percent")+xlab(NULL)+coord_flip() + 
+  guides(fill = guide_legend(reverse=T)) + theme_bw() +
+  theme(legend.position = c(0.875,0.085),legend.title = element_blank())
 
 as_tibble(summary.data_pub)
 as_tibble(summary.data_CC)
@@ -628,3 +631,6 @@ dbExecute(con, "UPDATE tblStudyObjectives
 dbExecute(con, "DELETE FROM tblStudyManagement")
 # 
 dbGetQuery(con, "SELECT * FROM tblStudyObjectives WHERE ID = 51")
+
+part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,16,17,18,20,21,22,23,24,25,21,23,
+26,25,26,27,27,28,29,30,31,29,32,30,33,32,33,34,35,36,37,38,39,39)

### 2019-08-22 Figures.R
rename two data tables to use dataTable notation, fix stringsplit for drive.data, add poster result outputs, remove unneeded shiny app code

@@ -262,13 +262,13 @@ part.data_CC <- part.data_CC %>%
   # 3. Add order column of row numbers
   mutate(order = row_number())
 
-part.data.table_pub<-part.data_pub %>%
+part.dataTable_pub<-part.data_pub %>%
     as_tibble() %>%
     select(Stage,Participants,Number) %>%
     rename("Participant Group"="Participants") %>%
     spread(Stage,Number,fill=0)
 
-part.data.table_CC<-part.data_CC %>%
+part.dataTable_CC<-part.data_CC %>%
   as_tibble() %>%
   select(Stage,Participants,Number) %>%
   rename("Participant Group"="Participants") %>%
@@ -277,7 +277,7 @@ part.data.table_CC<-part.data_CC %>%
 # What drivers are considered
 drive.data_pub<-data_pub %>%
     select(Drivers) %>%
-    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+    purrr::map(~ strsplit(as.character(.),split=",")) %>%
     purrr::map(unlist) %>%
     purrr::map(table) %>%
     plyr::ldply(data.frame) %>%
@@ -288,7 +288,7 @@ drive.data_pub<-data_pub %>%
 
 drive.data_CC<-data_CC %>%
   select(Drivers) %>%
-  purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+  purrr::map(~ strsplit(as.character(.),split=",")) %>%
   purrr::map(unlist) %>%
   purrr::map(table) %>%
   plyr::ldply(data.frame) %>%
@@ -391,7 +391,9 @@ part.data_pub<-mutate(part.data_pub,Analysis="Publication")
 part.data_CC<-mutate(part.data_CC,Analysis="Climate Change")
 part.data<-rbind(part.data_pub,part.data_CC) %>% 
   arrange(Stage, Percent) %>% 
-  mutate(order = row_number())
+  mutate(order = row_number()) %>% 
+  part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,16,17,18,20,21,22,23,24,25,21,23,
+                   26,25,26,27,27,28,29,30,31,29,32,30,33,32,33,34,35,36,37,38,39,39)
 
   
 ##### Data Analysis Outputs #####
[…]

### 2019-08-22 Poster_ICES.Rmd
Ices poster creation markdown file

---
main_topsize: 0.2 #percent coverage of the poster
main_bottomsize: 0.1
#ESSENTIALS
title: 'Use of management strategy evaluation to inform in fisheries management in a changing climate'
author:
  - name: '**Jonathan W. Cummings**'
    affil: 1
    main: true
    orcid: '0000-0001-8028-5787'
    twitter: jonwcummings 
    email: 'jcummings@umassd.edu'
  - name: Amanda Hart
    affil: 1
    main: true
    orcid: '0000-0002-6449-9328'
    email: 'ahart2@umassd.edu'
  - name: Gavin Fay
    main: true
    affil: 1
    email: 'gfay@umassd.edu'
affiliation:
  - num: 1
    address: School for Marine Science and Technology, University of Massachusetts - Dartmouth 836 South Rodney French Blvd,New Bedford, MA, USA 02744
main_findings:
  - "Make **better posters** with RMarkdown + **posterdown**."
logoleft_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
logoright_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
logocenter_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/qr-code-black.png
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
bibliography: packages.bib
link-citations: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
```

# Introduction

For most of its history fisheries management has taken place against the backdrop of ecosystem status with fishing impacts occupying center stage.  However, the increased rate and degree of ecosystem climatic change is also bringing ecosystem status to the forefront of fisheries management.
Management strategy evaluation (MSE) is commonly considered the gold standard of fisheries management approaches, and is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” (citation). Thus, MSE is a compelling tool to address the impact of climate change on fisheries management.
Despite numerous applications of MSE, the application that include climate change are limited, leaving uncertainty about the best approach to applying MSEs that account for a changing environment.

```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

## Objectives

1. Pick a template layout.
2. Write/ create your poster content distraction free.
3. Let posterdown do its thing!

# Methods

I will show here how to include poster elements that may be useful, such as an equation using mathjax:

$$
E = mc^2
$$

To reference a citation you can add your `.bib` file to the working directory and name it in the YAML metadata or generate an automated one as done here, then you only need to reference the label value in the `.bib` file. For example this package is built on top of the wonderful {pagedown} package and I will cite it at the end of this sentance using this in the rmd `[@R-pagedown]` [@R-pagedown].

To get a better understanding of how to include features like these please refer to the {posterdown} [wiki](https://github.com/posterdown/wiki).

**_Now on to the results!_**

# Results

Here you may have some figures to show off, bellow I have made a scatterplot with the infamous Iris dataset and I can even reference to the figure automatically like this, `Figure \@ref(fig:irisfigure)`, Figure \@ref(fig:irisfigure).

```{r, irisfigure, fig.cap='Here is a caption for the figure. This can be added by using the "fig.cap" option in the r code chunk options, see this [link](https://yihui.name/knitr/options/#plots) from the legend himself, [Yihui Xie](https://twitter.com/xieyihui).', out.width="80%"}
par(mar=c(2,2,0,1))
plot(x = iris$Sepal.Length, y = iris$Sepal.Width,
     col = iris$Species, pch = 19, xlab = "Sepal Length",
     ylab = "Sepal Width")
```

Maybe you want to show off some of that fancy code you spent so much time on to make that figure, well you can do that too! Just use the `echo=TRUE` option in the r code chunk options, Figure \@ref(fig:myprettycode)!

```{r myprettycode, echo=TRUE, fig.cap='Boxplots, so hot right now!', fig.height=6, out.width="80%"}
#trim whitespace
par(mar=c(2,2,0,0))
#plot boxplots
boxplot(iris$Sepal.Width~iris$Species,
        col = "#008080", 
        border = "#0b4545",
        ylab = "Sepal Width (cm)",
        xlab = "Species")
```

How about a neat table of data? See, Table \@ref(tab:iristable):

```{r, iristable}
knitr::kable(
  iris[1:15,1:5], format = "html",
  caption = "A table made with the **knitr::kable** function.",
  align = "c", col.names = c("Sepal <br> Length",
                             "Sepal <br> Width",
                             "Petal <br> Length",
                             "Petal <br> Width",
                             "Species"),
  escape = FALSE)
```

# References

### 2019-08-22 Poster_ICES.Rmd
Draft objectives section

@@
 
-1. Pick a template layout.
-2. Write/ create your poster content distraction free.
-3. Let posterdown do its thing!
+Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives.
+With an eye towards adaptive management, we seek to note the lessons learned from successful MSEs to:
+
+* Highlight unique aspects of MSEs that model climate change as a driver
+* Identify shortcomings in MSE documentation that delays learning
+* Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate
+
+Need climate change imagery
+
+Need MSE imagery
+
+Need learning and adaptive management imagery
 

### 2019-08-23 Figures.R
fix broken pipe, suppress output

@@ -391,14 +391,13 @@ part.data_pub<-mutate(part.data_pub,Analysis="Publication")
 part.data_CC<-mutate(part.data_CC,Analysis="Climate Change")
 part.data<-rbind(part.data_pub,part.data_CC) %>% 
   arrange(Stage, Percent) %>% 
-  mutate(order = row_number()) %>% 
-  part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,16,17,18,20,21,22,23,24,25,21,23,
+  mutate(order = row_number())
+part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,16,17,18,20,21,22,23,24,25,21,23,
                    26,25,26,27,27,28,29,30,31,29,32,30,33,32,33,34,35,36,37,38,39,39)
 
-  
 ##### Data Analysis Outputs #####
 # number of climate change MSE articles
-n_CC
+# n_CC
 
 # plot MSEs on map
 MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +

### 2019-08-23 Poster_ICES.Rmd
Play with title and header section; add data analysis and figure creation code

@@
 #ESSENTIALS
-title: 'Use of management strategy evaluation to inform in fisheries management in a changing climate'
+title: Use of management strategy evaluation to inform in fisheries management in a changing climate
 author:
@@
 main_findings:
-  - "Make **better posters** with RMarkdown + **posterdown**."
+  - "Make **better posters** with RMarkdown + **posterdown**.
+  Use of management strategy evaluation to inform in fisheries management in a changing climate"
 logoleft_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
@@
 ---
-
 ```{r, include=FALSE}
@@
 options(knitr.table.format = "html")
+##### load libraries #####
+library(tidyverse) # upgrade to base R
+library(readxl) # read in excel files
+library(shiny) # Shiny App
+library(shinythemes) # web themes for Shiny
+library(DT) # Data Table
+library(purrr) # for loop alternative
+library(ggmap) # map plotting
+library(mapdata) # basemap generation
+library(odbc) # database connection 
+library(RMySQL) # MySQL scripting in R
+library(DBI) # database interface
+
+load("C:/Users/jcummings/OneDrive - UMASS Dartmouth/SMAST/Research/MSE Problem Framing/MSEreview/MSEreview.RData")
+
+study<-study %>% 
+  unite(Citation,c(Authors,YearPub),sep=" ",remove=F)
+
+mgmt.join<-mgmt %>% 
+  group_by(fkStudyID) %>%
+  summarise(ManagementType = toString(sort(unique(ManagementType))),
+            AlternativesEvaluated = toString(sort(unique(AlternativesEvaluated))))
+
+obj.join<-obj %>%
+  group_by(fkStudyID) %>%
+  summarise(ObjectiveCategories = toString(sort(unique(ObjCategory))))
+
+# Join tables
+data<-full_join(study,mgmt.join,by=c("ID"="fkStudyID"))
+data<-full_join(data,obj.join,by=c("ID"="fkStudyID"))
+
+# Move comment column to the end
+data<-data %>%
+  select(-'Comments', 'Comments')
+
+# Filter data
+data_pub<-filter(data,IncludeInPublication==TRUE)
+data_CC<-filter(data,str_detect(Drivers,"Climate Change"))
+
+data_pub.join<-data_pub %>%
+  select(ID,Citation)
+data_CC.join<-data_CC %>%
+  select(ID,Citation)
+
+obj.data_pub<-left_join(data_pub.join,obj,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID","ID.y"))
+obj.data_CC<-left_join(data_CC.join,obj,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID","ID.y"))
+
+alt.data_pub<-left_join(data_pub.join,mgmt,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID","ID.y"))
+alt.data_CC<-left_join(data_CC.join,mgmt,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID","ID.y"))
+
+data_pub<-data_pub %>%
+  select(-c("ID"))
+data_CC<-data_CC %>%
+  select(-c("ID"))
+
+# Get columns whose width needs editing
+targets<-match(c("FullCitation","Comments"),names(data))
+
+#####-- Data analysis --#####
+# Get columns for study summary
+summary.col<-c("Citation",
+               "Species",
+               "Location",
+               "System")
+# Get columns for study drivers and problem
+prob.col<-c("Citation",
+            "ProblemDefinition",
+            "Drivers",
+            "ConsequencePrediction",
+            "TradeOffMethod_Exp",
+            "TradeOffMethod_Sub",
+            "Decision")
+# Get columns for frequency analysis
+freq.col<-c("ProcessExplicit",
+            "ProblemDefinitionExplicit",
+            "ObjectivesExplicit",
+            "AlternativesExplicit",
+            "TradeOffsExplicit",
+            "DecisionExplicit",
+            "RolesExplicit",
+            "OpenMeetings",
+            "ResultsAdopted")
+# Get columns for participant analysis
+part.col<-c("Leader",
+            "Participants",
+            "ObjElicitationSource_Exp",
+            "ProcedureElicitation_Exp",
+            "ObjElicitationSource_Sub",
+            "ProcedureElicitation_Sub")
+obj.col<-c("ObjType",
+           "ObjCategory",
+           "ObjDirection",
+           "ObjScale")
+alt.col<-c("ManagementType",
+           "AlternativesEvaluated")
+# Get columns for map
+map.col<-c("Latitude",
+           "Longitude",
+           "Citation",
+           "Drivers")
+
+# Select data summary columns
+summary.data_pub<-data_pub %>%
+    select(summary.col)
+summary.data_CC<-data_CC %>%
+  select(summary.col)
+
+# Frequency of method
+n_mse<-nrow(data)
+n_pub<-nrow(data_pub)
+n_obj_pub<-nrow(obj.data_pub)
+n_CC<-nrow(data_CC)
+n_obj_CC<-nrow(obj.data_CC)
+
+freq.data_pub<-data_pub %>%
+    select(freq.col) %>%
+    rename("Process"="ProcessExplicit",
+           "Problem"="ProblemDefinitionExplicit",
+           "Objectives"="ObjectivesExplicit",
+           "Alternatives"="AlternativesExplicit",
+           "Tradeoffs"="TradeOffsExplicit",
+           "Decision"="DecisionExplicit",
+           "Roles"="RolesExplicit",
+           "Open Meetings"="OpenMeetings",
+           "Adopted"="ResultsAdopted") %>%
+    summarise_all(funs(sum))%>%
+    gather(Explicit) %>%
+    mutate(Percent=value/n_mse*100) %>%
+    mutate(Explicit=factor(Explicit,levels=
+                  c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision",
+                               "Roles","Open Meetings","Adopted"))) %>%
+    rename("Number"="value")
+
+freq.data_CC<-data_CC %>%
+  select(freq.col) %>%
+  rename("Process"="ProcessExplicit",
+         "Problem"="ProblemDefinitionExplicit",
+         "Objectives"="ObjectivesExplicit",
+         "Alternatives"="AlternativesExplicit",
+         "Tradeoffs"="TradeOffsExplicit",
+         "Decision"="DecisionExplicit",
+         "Roles"="RolesExplicit",
+         "Open Meetings"="OpenMeetings",
+         "Adopted"="ResultsAdopted") %>%
+  summarise_all(funs(sum))%>%
+  gather(Explicit) %>%
+  mutate(Percent=value/n_mse*100) %>%
+  mutate(Explicit=factor(Explicit,levels=
+                           c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision",
+                             "Roles","Open Meetings","Adopted"))) %>%
+  rename("Number"="value")
+
+# Who participates
+part.data_pub<-data_pub %>%
+    select(part.col) %>%
+    rename("Process"="Leader",
+           "Doc Objectives"="ObjElicitationSource_Exp",
+           "Doc Alternatives"="ProcedureElicitation_Exp",
+           "Sub Objectives"="ObjElicitationSource_Sub",
+           "Sub Alternatives"="ProcedureElicitation_Sub")
+
+part.data_pub<- part.data_pub %>%
+    purrr::map(~ strsplit(as.character(.),split=",")) %>%
+    purrr::map(unlist) %>%
+    purrr::map(table)
+
+part.data_pub<-plyr::ldply(part.data_pub,data.frame)
+colnames(part.data_pub)<-c("Stage","Participants","Number")
+
+neworder <- c("Process","Participants","Doc Objectives",
+              "Sub Objectives","Doc Alternatives",
+              "Sub Alternatives")
+newlabels <- c("Process","Participants","Explicit Objectives Process",
+               "Subjective Objectives Process","Explicit Alternatives Process",
+               "Subjective Alternatives Process")
+
+part.data_pub <- part.data_pub %>%
+    mutate(Percent=Number/n_pub*100) %>%
+    mutate(Stage=factor(Stage,levels=neworder,labels=newlabels)) %>%
+    ungroup() %>%
+    # 2. Arrange by
+    #   i.  facet group =Stage
+    #   ii. bar height
+    arrange(Stage, Percent) %>%
+    # 3. Add order column of row numbers
+    mutate(order = row_number())
+
+part.data_CC<-data_CC %>%
+  select(part.col) %>%
+  rename("Process"="Leader",
+         "Doc Objectives"="ObjElicitationSource_Exp",
+         "Doc Alternatives"="ProcedureElicitation_Exp",
+         "Sub Objectives"="ObjElicitationSource_Sub",
+         "Sub Alternatives"="ProcedureElicitation_Sub")
+
+part.data_CC<- part.data_CC %>%
+  purrr::map(~ strsplit(as.character(.),split=",")) %>%
+  purrr::map(unlist) %>%
+  purrr::map(table)
+
+part.data_CC<-plyr::ldply(part.data_CC,data.frame)
+colnames(part.data_CC)<-c("Stage","Participants","Number")
+
+part.data_CC <- part.data_CC %>%
+  mutate(Percent=Number/n_CC*100) %>%
+  mutate(Stage=factor(Stage,levels=neworder,labels=newlabels)) %>%
+  ungroup() %>%
+  # 2. Arrange by
+  #   i.  facet group =Stage
+  #   ii. bar height
+  arrange(Stage, Percent) %>%
+  # 3. Add order column of row numbers
+  mutate(order = row_number())
+
+part.dataTable_pub<-part.data_pub %>%
+    as_tibble() %>%
+    select(Stage,Participants,Number) %>%
+    rename("Participant Group"="Participants") %>%
+    spread(Stage,Number,fill=0)
+
+part.dataTable_CC<-part.data_CC %>%
+  as_tibble() %>%
+  select(Stage,Participants,Number) %>%
+  rename("Participant Group"="Participants") %>%
+  spread(Stage,Number,fill=0)
+
+# What drivers are considered
+drive.data_pub<-data_pub %>%
+    select(Drivers) %>%
+    purrr::map(~ strsplit(as.character(.),split=",")) %>%
+    purrr::map(unlist) %>%
+    purrr::map(table) %>%
+    plyr::ldply(data.frame) %>%
+    select(Var1,Freq) %>%
+    rename("Driver"="Var1","Frequency"="Freq") %>%
+    mutate(Percent=Frequency/n_pub*100) %>%
+    arrange(desc(Frequency))
+
+drive.data_CC<-data_CC %>%
+  select(Drivers) %>%
+  purrr::map(~ strsplit(as.character(.),split=",")) %>%
+  purrr::map(unlist) %>%
+  purrr::map(table) %>%
+  plyr::ldply(data.frame) %>%
+  select(Var1,Freq) %>%
+  rename("Driver"="Var1","Frequency"="Freq") %>%
+  mutate(Percent=Frequency/n_CC*100) %>%
+  arrange(desc(Frequency))
+
+# What objectives categories were considered
+objcat.data_pub<-data_pub %>%
+    select(ObjectiveCategories) %>%
+    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+    purrr::map(unlist) %>%
+    purrr::map(table) %>%
+    plyr::ldply(data.frame) %>%
+    select(Var1,Freq) %>%
+    rename("Objective Category"="Var1","Frequency"="Freq") %>%
+    mutate(Percent=Frequency/n_pub*100) %>%
+    arrange(desc(Frequency))
+
+objcat.data_CC<-data_CC %>%
+  select(ObjectiveCategories) %>%
+  purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+  purrr::map(unlist) %>%
+  purrr::map(table) %>%
+  plyr::ldply(data.frame) %>%
+  select(Var1,Freq) %>%
+  rename("Objective Category"="Var1","Frequency"="Freq") %>%
+  mutate(Percent=Frequency/n_CC*100) %>%
+  arrange(desc(Frequency))
+
+# How were objectives defined
+obj.data_pub<-obj.data_pub %>%
+  select(obj.col) %>%
+  purrr::map(table) %>%
+  plyr::ldply(data.frame)
+
+obj.data_CC<-obj.data_CC %>%
+  select(obj.col) %>%
+  purrr::map(table) %>%
+  plyr::ldply(data.frame)
+
+colnames(obj.data_pub)<-colnames(obj.data_CC)<-c("Objective","Type","Number")
+
+neworder <- c("ObjCategory","ObjType","ObjDirection","ObjScale")
+newlabels <- c("Category","Type","Direction","Scale")
+
+# What Objective types are considered
+obj.dataTable_pub <- obj.data_pub %>%
+  mutate(Percent=round(Number/n_obj_pub*100,0)) %>%
+  mutate('Per MSE'=round(Number/n_pub,2)) %>%
+  mutate(Objective=factor(Objective,levels=neworder,labels=newlabels)) %>%
+  arrange(Objective,desc(Number))
+
+obj.dataTable_CC <- obj.data_CC %>%
+  mutate(Percent=round(Number/n_obj_CC*100,0)) %>%
+  mutate('Per MSE'=round(Number/n_CC,2)) %>%
+  mutate(Objective=factor(Objective,levels=neworder,labels=newlabels)) %>%
+  arrange(Objective,desc(Number))
+
+# What Alternative types are considered
+altcat.data_pub<-alt.data_pub %>%
+    select(ManagementType) %>%
+    purrr::map(~ strsplit(as.character(.),split=",")) %>%
+    purrr::map(unlist) %>%
+    purrr::map(table) %>%
+    plyr::ldply(data.frame) %>%
+    select(Var1,Freq) %>%
+    rename("Management Type"="Var1","Number"="Freq") %>%
+    mutate(Percent=round(Number/n_pub*100,0)) %>%
+    mutate('Per MSE'=round(Number/n_pub,2)) %>%
+    arrange(desc(Number))
+
+altcat.data_CC<-alt.data_CC %>%
+  select(ManagementType) %>%
+  purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+  purrr::map(unlist) %>%
+  purrr::map(table) %>%
+  plyr::ldply(data.frame) %>%
+  select(Var1,Freq) %>%
+  rename("Management Type"="Var1","Number"="Freq") %>%
+  mutate(Percent=round(Number/n_CC*100,0)) %>%
+  mutate('Per MSE'=round(Number/n_CC,2)) %>%
+  arrange(desc(Number))
+
+# Get map background for plotting the map
+world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
+
+# Where MSEs have occured
+map.data<-rbind(data_pub,data_CC) %>%
+  select(map.col) %>%
+  mutate(Drivers=str_extract(Drivers, "Climate Change")) %>%
+  mutate(Drivers=replace_na(Drivers, "Other"))
+
+# label MSEs articles by analysis for ploting
+freq.data_pub<-mutate(freq.data_pub,Analysis="Publication")
+freq.data_CC<-mutate(freq.data_CC,Analysis="Climate Change")
+freq.data<-rbind(freq.data_pub,freq.data_CC)
+part.data_pub<-mutate(part.data_pub,Analysis="Publication")
+part.data_CC<-mutate(part.data_CC,Analysis="Climate Change")
+part.data<-rbind(part.data_pub,part.data_CC) %>% 
+  arrange(Stage, Percent) %>%
+  mutate(order = row_number())
+part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,16,17,18,20,21,22,23,24,25,21,23,26,25,26,27,27,28,29,30,31,29,32,30,33,32,33,34,35,36,37,38,39,39)
+
+# ##### Data Analysis Outputs #####
+# # number of climate change MSE articles
+# # n_CC
+# 
+# # plot MSEs on map
+# MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
+#   geom_point(size=2) + scale_color_manual(values=c("#FF0033","#0000FF")) +  theme_void() +
+#   theme(legend.position = c(0.15, 0.25))
+# 
+# # plot explicit documentation of steps or components of MSE processes 
+# # Publication MSEs
+# Freq.plot_pub<-ggplot(freq.data_pub,aes(Explicit,Percent))+
+#     geom_col()+geom_vline(xintercept=3.5,linetype="dashed")+
+#     coord_flip()+scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
+#     scale_x_discrete(
+#       limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+#                "Alternatives","Objectives","Problem","Process"), 
+#       labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+#                "Alternatives","Objectives","Problem","Process"))
+# 
+# #Climate Change MSEs
+# Freq.plot_CC<-ggplot(freq.data_CC,aes(Explicit,Percent))+
+#   geom_col()+geom_vline(xintercept=3.5,linetype="dashed")+
+#   coord_flip()+scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
+#   scale_x_discrete(
+#     limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+#              "Alternatives","Objectives","Problem","Process"), 
+#     labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+#              "Alternatives","Objectives","Problem","Process"))
+# 
+# Freq.plot<-ggplot(freq.data,aes(Explicit,Percent,fill=Analysis))+
+#   geom_col(position="dodge")+geom_vline(xintercept=3.5,linetype="dashed")+
+#   coord_flip()+scale_fill_manual(values=c("#FF0033","#0000FF")) +
+#   scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
+#   scale_x_discrete(
+#     limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+#              "Alternatives","Objectives","Problem","Process"), 
+#     labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+#              "Alternatives","Objectives","Problem","Process")) +
+#   guides(fill = guide_legend(reverse=T)) + theme_bw()
+# 
+# Part.plot_pub<-ggplot(part.data_pub,aes(x=order,y=Percent)) +
+#     facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +  
+#     scale_x_continuous(breaks = part.data_pub$order,
+#       labels = part.data_pub$Participants)+
+#   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
+#     ylab("Percent")+coord_flip()
+# 
+# Part.plot_CC<-ggplot(part.data_CC,aes(x=order,y=Percent)) +
+#   facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +  
+#   scale_x_continuous(breaks = part.data_CC$order,
+#                      labels = part.data_CC$Participants)+
+#   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
+#   ylab("Percent")+coord_flip()
+# 
+# Part.plot<-ggplot(part.data,aes(x=order,y=Percent,fill=Analysis)) +
+#   facet_wrap(~Stage,scale="free",ncol=2) + geom_col(position="dodge") +  
+#   scale_fill_manual(values=c("#FF0033","#0000FF")) +
+#   scale_x_continuous(breaks = part.data$order,
+#                      labels = part.data$Participants)+
+#   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
+#   ylab("Percent")+xlab(NULL)+coord_flip() + 
+#   guides(fill = guide_legend(reverse=T)) + theme_bw() +
+#   theme(legend.position = c(0.875,0.085),legend.title = element_blank())
+# 
+# as_tibble(summary.data_pub)
+# as_tibble(summary.data_CC)
 ```
@@
 
+MSE articles Reviewed
+We conducted our search for management strategy evaluation documents using the Web of Science database.
+We searched in topic for “management strategy evaluation” on January 8th, 2019 in the SCI-EXPANDED index with a timespan of all years.
+This search returned 253 results.
+We sampled this list of 253 potential management strategy evaluations without replacement to produce an ordered list of articles for potential review.
+We reviewed the first 30 articles in this randomly ordered list that document a management strategy evaluation.
+Not all articles in the potential pool of 253 articles were management strategy evaluations.
+For example, once published the article would be returned by our Web of Science search, but as this is a review it is not documenting a management strategy evaluation.
+
+Identify management strategy evaluations that include climate change:
+Web of science  
+details  
+Score the MSE documentation:  
+Steps  
+Problem  
+Objectives  
+Alternatives  
+Consequences  
+Trade-offs  
+Decision  
+Participants  
+Roles
+
+We highlight lessons learned from these exercises and unique aspects of MSEs that model climate change as a driver relative to a sample of MSEs that do not consider climate change.
+
 I will show here how to include poster elements that may be useful, such as an equation using mathjax:
@@
 # Results
+
+Relatively few cases of climate change MSEs  
+Less likely to include stakeholders
 

### 2019-08-23 Poster_ICES.Rmd
Draft Objectives & Methods sections

@@
 
-# ##### Data Analysis Outputs #####
-# # number of climate change MSE articles
-# # n_CC
-# 
-# # plot MSEs on map
-# MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
-#   geom_point(size=2) + scale_color_manual(values=c("#FF0033","#0000FF")) +  theme_void() +
-#   theme(legend.position = c(0.15, 0.25))
-# 
-# # plot explicit documentation of steps or components of MSE processes 
-# # Publication MSEs
-# Freq.plot_pub<-ggplot(freq.data_pub,aes(Explicit,Percent))+
-#     geom_col()+geom_vline(xintercept=3.5,linetype="dashed")+
-#     coord_flip()+scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
-#     scale_x_discrete(
-#       limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-#                "Alternatives","Objectives","Problem","Process"), 
-#       labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-#                "Alternatives","Objectives","Problem","Process"))
-# 
-# #Climate Change MSEs
-# Freq.plot_CC<-ggplot(freq.data_CC,aes(Explicit,Percent))+
-#   geom_col()+geom_vline(xintercept=3.5,linetype="dashed")+
-#   coord_flip()+scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
-#   scale_x_discrete(
-#     limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-#              "Alternatives","Objectives","Problem","Process"), 
-#     labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-#              "Alternatives","Objectives","Problem","Process"))
-# 
-# Freq.plot<-ggplot(freq.data,aes(Explicit,Percent,fill=Analysis))+
-#   geom_col(position="dodge")+geom_vline(xintercept=3.5,linetype="dashed")+
-#   coord_flip()+scale_fill_manual(values=c("#FF0033","#0000FF")) +
-#   scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
-#   scale_x_discrete(
-#     limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-#              "Alternatives","Objectives","Problem","Process"), 
-#     labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-#              "Alternatives","Objectives","Problem","Process")) +
-#   guides(fill = guide_legend(reverse=T)) + theme_bw()
-# 
-# Part.plot_pub<-ggplot(part.data_pub,aes(x=order,y=Percent)) +
-#     facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +  
-#     scale_x_continuous(breaks = part.data_pub$order,
-#       labels = part.data_pub$Participants)+
-#   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
-#     ylab("Percent")+coord_flip()
-# 
-# Part.plot_CC<-ggplot(part.data_CC,aes(x=order,y=Percent)) +
-#   facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +  
-#   scale_x_continuous(breaks = part.data_CC$order,
-#                      labels = part.data_CC$Participants)+
-#   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
-#   ylab("Percent")+coord_flip()
-# 
-# Part.plot<-ggplot(part.data,aes(x=order,y=Percent,fill=Analysis)) +
-#   facet_wrap(~Stage,scale="free",ncol=2) + geom_col(position="dodge") +  
-#   scale_fill_manual(values=c("#FF0033","#0000FF")) +
-#   scale_x_continuous(breaks = part.data$order,
-#                      labels = part.data$Participants)+
-#   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
-#   ylab("Percent")+xlab(NULL)+coord_flip() + 
-#   guides(fill = guide_legend(reverse=T)) + theme_bw() +
-#   theme(legend.position = c(0.875,0.085),legend.title = element_blank())
-# 
-# as_tibble(summary.data_pub)
-# as_tibble(summary.data_CC)
+##### Data Analysis Outputs #####
+# number of climate change MSE articles
+# n_CC
+
+# plot MSEs on map
+MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
+  geom_point(size=2) + scale_color_manual(values=c("#FF0033","#0000FF")) +  theme_void() +
+  theme(legend.position = c(0.15, 0.25))
+
+# plot explicit documentation of steps or components of MSE processes
+# Publication MSEs
+Freq.plot_pub<-ggplot(freq.data_pub,aes(Explicit,Percent))+
+    geom_col()+geom_vline(xintercept=3.5,linetype="dashed")+
+    coord_flip()+scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
+    scale_x_discrete(
+      limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+               "Alternatives","Objectives","Problem","Process"),
+      labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+               "Alternatives","Objectives","Problem","Process"))
+
+#Climate Change MSEs
+Freq.plot_CC<-ggplot(freq.data_CC,aes(Explicit,Percent))+
+  geom_col()+geom_vline(xintercept=3.5,linetype="dashed")+
+  coord_flip()+scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
+  scale_x_discrete(
+    limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+             "Alternatives","Objectives","Problem","Process"),
+    labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+             "Alternatives","Objectives","Problem","Process"))
+
+Freq.plot<-ggplot(freq.data,aes(Explicit,Percent,fill=Analysis))+
+  geom_col(position="dodge")+geom_vline(xintercept=3.5,linetype="dashed")+
+  coord_flip()+scale_fill_manual(values=c("#FF0033","#0000FF")) +
+  scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
+  scale_x_discrete(
+    limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+             "Alternatives","Objectives","Problem","Process"),
+    labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+             "Alternatives","Objectives","Problem","Process")) +
+  guides(fill = guide_legend(reverse=T)) + theme_bw()
+
+Part.plot_pub<-ggplot(part.data_pub,aes(x=order,y=Percent)) +
+    facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +
+    scale_x_continuous(breaks = part.data_pub$order,
+      labels = part.data_pub$Participants)+
+  scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
+    ylab("Percent")+coord_flip()
+
+Part.plot_CC<-ggplot(part.data_CC,aes(x=order,y=Percent)) +
+  facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +
+  scale_x_continuous(breaks = part.data_CC$order,
+                     labels = part.data_CC$Participants)+
+  scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
+  ylab("Percent")+coord_flip()
+
+Part.plot<-ggplot(part.data,aes(x=order,y=Percent,fill=Analysis)) +
+  facet_wrap(~Stage,scale="free",ncol=2) + geom_col(position="dodge") +
+  scale_fill_manual(values=c("#FF0033","#0000FF")) +
+  scale_x_continuous(breaks = part.data$order,
+                     labels = part.data$Participants)+
+  scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
+  ylab("Percent")+xlab(NULL)+coord_flip() +
+  guides(fill = guide_legend(reverse=T)) + theme_bw() +
+  theme(legend.position = c(0.875,0.085),legend.title = element_blank())
+
+as_tibble(summary.data_pub)
+as_tibble(summary.data_CC)
 ```
@@
 # Introduction
+For most of its history fisheries management has prodominately focused on fishing impacts with ecosystem status viewed as a background constant.
+The increasing rate and degree of ecosystem climatic change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
 
-For most of its history fisheries management has taken place against the backdrop of ecosystem status with fishing impacts occupying center stage.
-However, the increased rate and degree of ecosystem climatic change is also bringing ecosystem status to the forefront of fisheries management.
 Management strategy evaluation (MSE) is commonly considered the gold standard of fisheries management approaches, and is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” (citation).
 Thus, MSE is a compelling tool to address the impact of climate change on fisheries management.
-Despite numerous applications of MSE, the application that include climate change are limited, leaving uncertainty about the best approach to applying MSEs that account for a changing environment.
+
+Despite numerous applications of MSE, the applications that include climate change are limited, leaving uncertainty about the best approach to MSE use for fisheries management that accounts for a changing environment.
+
+Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives.
+Using the structured decision making (SDM) process- the decision making framework in which adaptive management occurs - as our framework, how do published MSE projects utilize standard SDM components?
 
@@
 ## Objectives
+We review published management strategy evaluations (MSEs) against the steps of a structured decision making (SDM) framework to note how the literature accelerates or delays learning.
 
-Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives.
-With an eye towards adaptive management, we seek to note the lessons learned from successful MSEs to:
-
+* Identify explicit documentation and use (or disuse) of SDM steps in MSEs that enable or limit adaptation in future MSEs.
+* Identify participants in MSE processes that enable or limit adaptation in future MSEs.
 * Highlight unique aspects of MSEs that model climate change as a driver
-* Identify shortcomings in MSE documentation that delays learning
 * Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate  
-
-Need climate change imagery
-
-Need MSE imagery
-
-Need learning and adaptive management imagery
+* Provide insights about MSE processes and documentation that can further learning to improve future MSEs
 
 # Methods
+#### Published MSE article selection
+We conducted our search for management strategy evaluations (MSEs) in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019.
+This search returned 253 results.
 
-MSE articles Reviewed
-We conducted our search for management strategy evaluation documents using the Web of Science database.
-We searched in topic for “management strategy evaluation” on January 8th, 2019 in the SCI-EXPANDED index with a timespan of all years.
-This search returned 253 results.
-We sampled this list of 253 potential management strategy evaluations without replacement to produce an ordered list of articles for potential review.
-We reviewed the first 30 articles in this randomly ordered list that document a management strategy evaluation.
-Not all articles in the potential pool of 253 articles were management strategy evaluations.
-For example, once published the article would be returned by our Web of Science search, but as this is a review it is not documenting a management strategy evaluation.
+We took a random sample of these 253 articles and reviewed 30 that document a management strategy evaluation.
+Not all articles in the potential pool of 253 articles were management strategy evaluations due to articles that are reviews, meta-analyses, etc., or simply cite other MSE articles.
+Of the 253 articles ### documented a management strategy evaluation, and 11 included climate change as a driver of fishery status.
 
-Identify management strategy evaluations that include climate change:
-Web of science  
-details  
+| MSE type | Count |
+|:--:|:--:|
+| Published | ### |
+| Reviewed | 30 |
+| Climate Change | 11 |
+
 Score the MSE documentation:  
@@
 We highlight lessons learned from these exercises and unique aspects of MSEs that model climate change as a driver relative to a sample of MSEs that do not consider climate change.
-
-I will show here how to include poster elements that may be useful, such as an equation using mathjax:
-
-$$
-E = mc^2
-$$
-
-To reference a citation you can add your `.bib` file to the working directory and name it in the YAML metadata or generate an automated one as done here, then you only need to reference the label value in the `.bib` file.
-For example this package is built on top of the wonderful {pagedown} package and I will cite it at the end of this sentance using this in the rmd `[@R-pagedown]` [@R-pagedown].
-
-To get a better understanding of how to include features like these please refer to the {posterdown} [wiki](https://github.com/posterdown/wiki).
-
-**_Now on to the results!_**
 
@@
 # References
+To reference a citation you can add your `.bib` file to the working directory and name it in the YAML metadata or generate an automated one as done here, then you only need to reference the label value in the `.bib` file.
+For example this package is built on top of the wonderful {pagedown} package and I will cite it at the end of this sentance using this in the rmd `[@R-pagedown]` [@R-pagedown].
+
+How to include an equation using mathjax:
+
+$$ E = mc^2 $$
+
+To get a better understanding of how to include features like these please refer to the {posterdown} [wiki](https://github.com/posterdown/wiki).

### 2019-08-28 Poster_ICES.Rmd
Change 'Publication' label to 'Reviewed'; remove unused plots and tables that separated Pub and CC results; and driver data table

@@
 #ESSENTIALS
-title: Use of management strategy evaluation to inform in fisheries management in a changing climate
 author:
@@
 library(DBI) # database interface
+library(kableExtra) # extra formating of the tables
 
@@
 # label MSEs articles by analysis for ploting
-freq.data_pub<-mutate(freq.data_pub,Analysis="Publication")
+freq.data_pub<-mutate(freq.data_pub,Analysis="Reviewed")
 freq.data_CC<-mutate(freq.data_CC,Analysis="Climate Change")
 freq.data<-rbind(freq.data_pub,freq.data_CC)
-part.data_pub<-mutate(part.data_pub,Analysis="Publication")
+part.data_pub<-mutate(part.data_pub,Analysis="Reviewed")
 part.data_CC<-mutate(part.data_CC,Analysis="Climate Change")
@@
   mutate(order = row_number())
-part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,16,17,18,20,21,22,23,24,25,21,23,26,25,26,27,27,28,29,30,31,29,32,30,33,32,33,34,35,36,37,38,39,39)
+part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,
+                   16,17,18,20,21,22,23,24,25,21,23,26,25,26,27,27,28,29,30,31,29,32,30,33,
+                   32,33,34,35,36,37,38,39,39)
+drive.data_pub<-mutate(drive.data_pub,Analysis="Reviewed") %>% 
+  mutate(Percent=round(Percent,0)) %>% 
+  select(Analysis,Driver,Percent,Frequency)
+drive.data_CC<-mutate(drive.data_CC,Analysis="Climate Change")%>% 
+  mutate(Percent=round(Percent,0)) %>% 
+  select(Analysis,Driver,Percent,Frequency)
+drive.data<-rbind(drive.data_pub,drive.data_CC) %>% 
+  select(Analysis,Driver,Percent,Frequency) %>% 
+  arrange(desc(Percent),Driver)
 
@@
 # plot explicit documentation of steps or components of MSE processes
-# Publication MSEs
-Freq.plot_pub<-ggplot(freq.data_pub,aes(Explicit,Percent))+
-    geom_col()+geom_vline(xintercept=3.5,linetype="dashed")+
-    coord_flip()+scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
-    scale_x_discrete(
-      limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-               "Alternatives","Objectives","Problem","Process"),
-      labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-               "Alternatives","Objectives","Problem","Process"))
-
-#Climate Change MSEs
-Freq.plot_CC<-ggplot(freq.data_CC,aes(Explicit,Percent))+
-  geom_col()+geom_vline(xintercept=3.5,linetype="dashed")+
-  coord_flip()+scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
-  scale_x_discrete(
-    limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-             "Alternatives","Objectives","Problem","Process"),
-    labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-             "Alternatives","Objectives","Problem","Process"))
-
+# Reviewed MSEs
 Freq.plot<-ggplot(freq.data,aes(Explicit,Percent,fill=Analysis))+
   geom_col(position="dodge")+geom_vline(xintercept=3.5,linetype="dashed")+
+  geom_vline(xintercept=1.5,linetype="dashed")+
   coord_flip()+scale_fill_manual(values=c("#FF0033","#0000FF")) +
@@
              "Alternatives","Objectives","Problem","Process")) +
-  guides(fill = guide_legend(reverse=T)) + theme_bw()
-
-Part.plot_pub<-ggplot(part.data_pub,aes(x=order,y=Percent)) +
-    facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +
-    scale_x_continuous(breaks = part.data_pub$order,
-      labels = part.data_pub$Participants)+
-  scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
-    ylab("Percent")+coord_flip()
-
-Part.plot_CC<-ggplot(part.data_CC,aes(x=order,y=Percent)) +
-  facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +
-  scale_x_continuous(breaks = part.data_CC$order,
-                     labels = part.data_CC$Participants)+
-  scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
-  ylab("Percent")+coord_flip()
+  guides(fill = guide_legend(reverse=T)) + theme_bw() +
+  theme(legend.position = c(0.875,0.085),legend.title = element_blank())
 
@@
 ```
+---
+title: Use of management strategy evaluation to inform in fisheries management in a changing climate
+---
 
@@
 | Published | ### |
-| Reviewed | 30 |
+| Reviewed Sample | 30 |
 | Climate Change | 11 |
@@
 
-Here you may have some figures to show off, bellow I have made a scatterplot with the infamous Iris dataset and I can even reference to the figure automatically like this, `Figure \@ref(fig:irisfigure)`, Figure \@ref(fig:irisfigure).
+Map of where reviewed MSEs and climate change focues MSEs occurred: Figure \@ref(fig:mapfigure).
 
-```{r, irisfigure, fig.cap='Here is a caption for the figure.
-This can be added by using the "fig.cap" option in the r code chunk options, see this [link](https://yihui.name/knitr/options/#plots) from the legend himself, [Yihui Xie](https://twitter.com/xieyihui).', out.width="80%"}
-par(mar=c(2,2,0,1))
-plot(x = iris$Sepal.Length, y = iris$Sepal.Width,
-     col = iris$Species, pch = 19, xlab = "Sepal Length",
-     ylab = "Sepal Width")
+```{r, mapfigure, fig.cap='Here is a caption for the figure.
+This can be added by using the "fig.cap" option in the r code chunk options, see this [link](https://yihui.name/knitr/options/#plots) from the legend himself, [Yihui Xie](https://twitter.com/xieyihui).', out.width="100%"}
+MSE.map
 ```
 
-Maybe you want to show off some of that fancy code you spent so much time on to make that figure, well you can do that too!
-Just use the `echo=TRUE` option in the r code chunk options, Figure \@ref(fig:myprettycode)!
+Figure \@ref(fig:explicit)!
 
-```{r myprettycode, echo=TRUE, fig.cap='Boxplots, so hot right now!', fig.height=6, out.width="80%"}
-#trim whitespace
-par(mar=c(2,2,0,0))
-#plot boxplots
-boxplot(iris$Sepal.Width~iris$Species,
-        col = "#008080", 
-        border = "#0b4545",
-        ylab = "Sepal Width (cm)",
-        xlab = "Species")
+```{r explicit, fig.cap='Percentage of MSEs that included explicit documentation of these SDM process steps!', fig.height=6, out.width="100%"}
+text1<-expression(paste("Absent or \nInexplicit"))
+text2<-paste("Absent")
+text3<-paste("Unknown")
+Freq.plot + annotate("text", x = 5, y = 55, label = text1, size=25) + 
+  annotate("text", x = 2.6, y = 50, label = text2, size=20) + 
+  annotate("text", x = 1, y = 50, label = text3, size=12)
 ```
 
-How about a neat table of data? See, Table \@ref(tab:iristable):
+Figure \@ref(fig:participants)!
+
+```{r participants, fig.cap='Who participated during these stages of the SDM process', fig.height=6, out.width="100%"}
+Part.plot
+```
+
+The drivers modeled in the reviewed MSEs Table \@ref(tab:driversPub) and the climate change MSEs Table \@ref(tab:driversCC) [description of results].
+
+```{r, drivers}
+color_Pub <- which(drive.data$Analysis=='Reviewed')
+color_CC <- which(drive.data$Analysis=='Climate Change')
+
+drive.data %>% 
+  kable(align = "c") %>%
+  row_spec(color_CC, color = "#FF0033", background = "white") %>% 
+  row_spec(color_Pub, color = "#0000FF", background = "white")
+
+knitr::kable(
+  as_tibble(drive.data), format = "html",
+  caption = "Drivers modeled in management strategy evaluations",
+  align = "c", escape = FALSE)
+```
 

### 2019-08-28 Poster_ICES.Rmd
Remove footer; remove title for now; wrangle data for tables; add driver plot; add tables

@@
 main_topsize: 0.2 #percent coverage of the poster
-main_bottomsize: 0.1
+main_bottomsize: 0
 #ESSENTIALS
+title:
 author:
@@
 drive.data<-rbind(drive.data_pub,drive.data_CC) %>% 
-  select(Analysis,Driver,Percent,Frequency) %>% 
-  arrange(desc(Percent),Driver)
+  arrange(Percent) %>%
+  mutate(order = row_number())
+drive.data$order<-c(6,5,4,3,2,1,8,7,12,9,3,15,14,13,12,11,10,14,16,15,
+                   16,6)
+# Wrangle objective category data by reviewed, climate change, and combined
+objcat.data_pub<-mutate(objcat.data_pub,Analysis="Reviewed") %>% 
+  mutate(Percent=round(Percent,0)) %>% 
+  select(Analysis,'Objective Category',Percent,Frequency)
+objcat.data_CC<-mutate(objcat.data_CC,Analysis="Climate Change")%>% 
+  mutate(Percent=round(Percent,0)) %>% 
+  select(Analysis,'Objective Category',Percent,Frequency)
+objcat.data<-rbind(objcat.data_pub,objcat.data_CC) %>% 
+  rename(OC='Objective Category') %>% 
+  arrange(OC) %>% 
+  rename('Objective Category'=OC)
+
+# Wrangle objective data by reviewed, climate change, and combined
+obj.dataTable_pub<-mutate(obj.dataTable_pub,Analysis="Reviewed") %>% 
+  mutate(Percent=round(Percent,0))
+obj.dataTable_CC<-mutate(obj.dataTable_CC,Analysis="Climate Change")%>% 
+  mutate(Percent=round(Percent,0))
+obj.dataTable<-rbind(obj.dataTable_pub,obj.dataTable_CC) %>% 
+  filter(Objective=="Category") %>%
+  group_by(Analysis) %>% 
+  summarise('AVG'=sum(Number)) %>% 
+  mutate(count=c(11,30)) %>% 
+  mutate('Per MSE'=round(AVG/count,1)) %>% 
+  mutate('Type'=c("Objectives","Objectives"))
+
+altcat.data_pub<-mutate(altcat.data_pub,Analysis="Reviewed") %>% 
+  mutate(Percent=round(Percent,0))
+altcat.data_CC<-mutate(altcat.data_CC,Analysis="Climate Change")%>% 
+  mutate(Percent=round(Percent,0))
+altcat.data<-rbind(altcat.data_pub,altcat.data_CC) %>%
+  group_by(Analysis) %>% 
+  summarise('AVG'=sum(Number)) %>% 
+  mutate(count=c(11,30)) %>% 
+  mutate('Per MSE'=round(AVG/count,1)) %>% 
+  mutate('Type'=c("Alternatives","Alternatives"))
+
+per.MSE<-rbind(obj.dataTable,altcat.data) %>% 
+  mutate(order=c(2,1,4,3)) %>% 
+  arrange(order) %>% 
+  select(Type,Analysis,'Per MSE')
 
@@
 
-as_tibble(summary.data_pub)
-as_tibble(summary.data_CC)
+Driver.plot<-ggplot(drive.data,aes(x=order,y=Percent,fill=Analysis)) +
+  geom_col(position="dodge") +
+  scale_fill_manual(values=c("#FF0033","#0000FF")) +
+  scale_x_continuous(breaks = drive.data$order,
+                     labels = drive.data$Driver)+
+  scale_y_continuous(limits=c(0,100),expand = c(0,0))+
+  ylab("Percent")+xlab(NULL)+coord_flip() +
+  guides(fill = guide_legend(reverse=T)) + theme_bw() +
+  theme(legend.position = c(0.85,0.085),legend.title = element_blank())
 ```
----
-title: Use of management strategy evaluation to inform in fisheries management in a changing climate
----
 
 # Introduction
-For most of its history fisheries management has prodominately focused on fishing impacts with ecosystem status viewed as a background constant.
-The increasing rate and degree of ecosystem climatic change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
+For most of its history fisheries management has focused on fishing impacts with ecosystem status viewed as a background constant.
+The increasing rate and degree of climate change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
 
@@
 
-Score the MSE documentation:  
-Steps  
-Problem  
-Objectives  
-Alternatives  
-Consequences  
-Trade-offs  
-Decision  
-Participants  
-Roles
-
 We highlight lessons learned from these exercises and unique aspects of MSEs that model climate change as a driver relative to a sample of MSEs that do not consider climate change.
@@
 
-```{r, mapfigure, fig.cap='Here is a caption for the figure.
+```{r, mapfigure, fig.cap='Map of MSE study locations.
+Here is a caption for the figure.
 This can be added by using the "fig.cap" option in the r code chunk options, see this [link](https://yihui.name/knitr/options/#plots) from the legend himself, [Yihui Xie](https://twitter.com/xieyihui).', out.width="100%"}
@@
 
-Figure \@ref(fig:explicit)!
+Figure \@ref(fig:explicit)
 
@@
 
-Figure \@ref(fig:participants)!
+Figure \@ref(fig:participants)
 
@@
 
-The drivers modeled in the reviewed MSEs Table \@ref(tab:driversPub) and the climate change MSEs Table \@ref(tab:driversCC) [description of results].
+Figure \@ref(fig:driversPlot)
 
-```{r, drivers}
-color_Pub <- which(drive.data$Analysis=='Reviewed')
-color_CC <- which(drive.data$Analysis=='Climate Change')
+```{r driversPlot, fig.cap='text', fig.height=6, out.width="100%"}
+Driver.plot
+```
 
-drive.data %>% 
-  kable(align = "c") %>%
+The categories of objectives considered in MSEs, Table \@ref(tab:objectiveCategories)
+
+```{r, objectiveCategories}
+color_Pub <- which(objcat.data$Analysis=='Reviewed')
+color_CC <- which(objcat.data$Analysis=='Climate Change')
+
+objcat.data %>%
+  kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>%
   row_spec(color_CC, color = "#FF0033", background = "white") %>%
   row_spec(color_Pub, color = "#0000FF", background = "white")
-
-knitr::kable(
-  as_tibble(drive.data), format = "html",
-  caption = "Drivers modeled in management strategy evaluations",
-  align = "c", escape = FALSE)
 ```
 
-```{r, iristable}
-knitr::kable(
-  iris[1:15,1:5], format = "html",
-  caption = "A table made with the **knitr::kable** function.",
-  align = "c", col.names = c("Sepal <br> Length",
-                             "Sepal <br> Width",
-                             "Petal <br> Length",
-                             "Petal <br> Width",
-                             "Species"),
-  escape = FALSE)
+A greater number of objectives were considered per MSE in the random sample of reviewed MSEs than in the climate change MSEs (Table \@ref(tab:PerMSE))
+
+```{r, PerMSE}
+color_Pub <- which(per.MSE$Analysis=='Reviewed')
+color_CC <- which(per.MSE$Analysis=='Climate Change')
+
+per.MSE %>%
+  kable(align = "c", caption = "Objective data reviewed") %>%  
+  row_spec(color_CC, color = "#FF0033", background = "white") %>%
+  row_spec(color_Pub, color = "#0000FF", background = "white")
 ```

### 2019-08-29 Poster_ICES.Rmd
Remove footer; edit title; edit main_findings; add custom bibliography; edit per.MSE table; edit text

@@
 ---
-main_topsize: 0.2 #percent coverage of the poster
+main_topsize: 0.25 #percent coverage of the poster
 main_bottomsize: 0
 #ESSENTIALS
-title:
+title: Use of management strategy evaluation to inform fisheries management in a changing climate
 author:
@@
     address: School for Marine Science and Technology, University of Massachusetts - Dartmouth 836 South Rodney French Blvd,New Bedford, MA, USA 02744
-main_findings:
-  - "Make **better posters** with RMarkdown + **posterdown**.
-  Use of management strategy evaluation to inform in fisheries management in a changing climate"
-logoleft_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
-logoright_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
-logocenter_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/qr-code-black.png
+main_findings: "Explicit documentation of MSE processes is limited.
+<br> Climate change MSEs have been less explict, more scientist driven, included fewer objectives -- with less consideration of economic or social objectives -- and fewer drivers."
 output: 
@@
     number_sections: false
-bibliography: packages.bib
+bibliography: MSEreview.bib
 link-citations: true
+nocite: '@*'
+title_textsize: "50pt"
+column_numbers: 4
+#body_textsize: "20pt"
 ---
@@
 
-per.MSE<-rbind(obj.dataTable,altcat.data) %>% 
-  mutate(order=c(2,1,4,3)) %>% 
+per.MSE<-obj.dataTable %>% 
+  mutate(order=c(2,1)) %>% 
   arrange(order) %>% 
-  select(Type,Analysis,'Per MSE')
+  select(Analysis,'Per MSE')
 
@@
 
-Management strategy evaluation (MSE) is commonly considered the gold standard of fisheries management approaches, and is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” (citation).
+Management strategy evaluation (MSE) is commonly considered the gold standard of fisheries management approaches, and is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@R-Punt].
 Thus, MSE is a compelling tool to address the impact of climate change on fisheries management.
@@
 Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives.
-Using the structured decision making (SDM) process- the decision making framework in which adaptive management occurs - as our framework, how do published MSE projects utilize standard SDM components?
+Using the structured decision making (SDM) process -- the decision making framework in which adaptive management occurs -- as our framework, how do published MSE projects utilize standard SDM components and support learning within the MSE practitioner community?
 
@@
 ## Objectives
-We review published management strategy evaluations (MSEs) against the steps of a structured decision making (SDM) framework to note how the literature accelerates or delays learning.
+We review published MSEs against the steps of the SDM framework to:
 
-* Identify explicit documentation and use (or disuse) of SDM steps in MSEs that enable or limit adaptation in future MSEs.
-* Identify participants in MSE processes that enable or limit adaptation in future MSEs.
-* Highlight unique aspects of MSEs that model climate change as a driver
+* Determine the rate of documentation and use of steps in the SDM process.
+* Identify participants in MSE processes.
+* Highlight unique aspects of MSEs that model climate change as a driver.
 * Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate  
@@
 #### Published MSE article selection
-We conducted our search for management strategy evaluations (MSEs) in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019.
+We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019.
 This search returned 253 results.
 
-We took a random sample of these 253 articles and reviewed 30 that document a management strategy evaluation.
-Not all articles in the potential pool of 253 articles were management strategy evaluations due to articles that are reviews, meta-analyses, etc., or simply cite other MSE articles.
-Of the 253 articles ### documented a management strategy evaluation, and 11 included climate change as a driver of fishery status.
+We took a random sample of these 253 articles and reviewed 30 that document a MSE.
+Not all articles in the potential pool of 253 articles were MSEs due to articles that are reviews, meta-analyses, etc., or simply cite other MSE articles.
+Of the 253 articles, ### documented a MSE, and 11 included climate change as a driver of fishery status.
 
@@
 
-We highlight lessons learned from these exercises and unique aspects of MSEs that model climate change as a driver relative to a sample of MSEs that do not consider climate change.
+# Results
+Apart from one each in Europe and the Western Pacific, to date MSEs modeling climate change have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).
 
-# Results
-
-Relatively few cases of climate change MSEs  
-Less likely to include stakeholders
-
-Map of where reviewed MSEs and climate change focues MSEs occurred: Figure \@ref(fig:mapfigure).
-
-```{r, mapfigure, fig.cap='Map of MSE study locations.
-Here is a caption for the figure.
-This can be added by using the "fig.cap" option in the r code chunk options, see this [link](https://yihui.name/knitr/options/#plots) from the legend himself, [Yihui Xie](https://twitter.com/xieyihui).', out.width="100%"}
+```{r, mapfigure, fig.cap='Map of MSE study locations.', out.width="100%"}
 MSE.map
@@
 
-Figure \@ref(fig:explicit)
+The vast majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Figure \@ref(fig:explicit)).
 
@@
 
-Figure \@ref(fig:participants)
+The primary participants and leaders of the MSE processes were scientists, while some processes were lead by governments and management agencies (Figure \@ref(fig:participants)).
+Where explicit documentation occurred objectives and alternatives were elicited from a variety of participants, with scientists seemingly selecting these components in the inexplicit cases.
 
@@
 
-Figure \@ref(fig:driversPlot)
+General uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the reviewed MSEs (Figure \@ref(fig:driversPlot)).
 
@@
 
-The categories of objectives considered in MSEs, Table \@ref(tab:objectiveCategories)
+A greater number of objectives were considered per MSE in the random sample of reviewed MSEs than in the climate change MSEs (Table \@ref(tab:PerMSE)).
+Climate change MSEs were less likely to consider economic and social objectives than the MSEs sample for review (Table \@ref(tab:objectiveCategories)).
+
+```{r, PerMSE}
+color_Pub <- which(per.MSE$Analysis=='Reviewed')
+color_CC <- which(per.MSE$Analysis=='Climate Change')
+
+per.MSE %>%
+  kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%  
+  row_spec(color_CC, color = "#FF0033", background = "white") %>%
+  row_spec(color_Pub, color = "#0000FF", background = "white")
+```
 
@@
 
-A greater number of objectives were considered per MSE in the random sample of reviewed MSEs than in the climate change MSEs (Table \@ref(tab:PerMSE))
+# Conclusions
+The picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes.
+While documentation about the modeling conducted in a MSE was explicit, documentation about the process used to make decision about the MSE model were limited.
+Scientists participated in, or seemingly directly selected the objectives and alternatives to evaluate in most MSEs.
+Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.
+Conservation, Economic, and Yield focused objectives were commonly considered, while broader social objectives were rarely considered.
 
-```{r, PerMSE}
-color_Pub <- which(per.MSE$Analysis=='Reviewed')
-color_CC <- which(per.MSE$Analysis=='Climate Change')
+To date climate change has not been considered in many MSEs.
+MSEs modeling climate change have been less explicitly documented, more likely to be scientist driven, and included fewer system drivers and fewer objectives, with less consideration of economic or social objectives.
 
-per.MSE %>%
-  kable(align = "c", caption = "Objective data reviewed") %>%  
-  row_spec(color_CC, color = "#FF0033", background = "white") %>%
-  row_spec(color_Pub, color = "#0000FF", background = "white")
-```
+# More Details
+Check out our shiny app (<https://jonathancummings.shinyapps.io/MSEreview/>) and keep an eye out for the forthcoming publication.
 
 # References
-To reference a citation you can add your `.bib` file to the working directory and name it in the YAML metadata or generate an automated one as done here, then you only need to reference the label value in the `.bib` file.
-For example this package is built on top of the wonderful {pagedown} package and I will cite it at the end of this sentance using this in the rmd `[@R-pagedown]` [@R-pagedown].
-
-How to include an equation using mathjax:
-
-$$ E = mc^2 $$
-
-To get a better understanding of how to include features like these please refer to the {posterdown} [wiki](https://github.com/posterdown/wiki).

### 2019-08-30 Poster_ICES.Rmd
Edit format for printing

@@
 ---
-main_topsize: 0.25 #percent coverage of the poster
+main_topsize: 0.2 #percent coverage of the poster
 main_bottomsize: 0
@@
 main_findings: "Explicit documentation of MSE processes is limited.
-<br> Climate change MSEs have been less explict, more scientist driven, included fewer objectives -- with less consideration of economic or social objectives -- and fewer drivers."
+<br> <br> Climate change MSEs have been less explict, more scientist driven, included fewer objectives -- with less consideration of economic or social objectives -- and fewer drivers."
 output: 
@@
 title_textsize: "50pt"
-column_numbers: 4
-#body_textsize: "20pt"
+main_textsize: "90pt"
+column_numbers: 3
+body_textsize: "29.5pt"
 ---
@@
 
+</br>
+</br>
+
 # Results

### 2019-09-03 Poster_ICES.Rmd
Edits based on feedback; header spelling; figure captions; sdm citation;

@@
     address: School for Marine Science and Technology, University of Massachusetts - Dartmouth 836 South Rodney French Blvd,New Bedford, MA, USA 02744
-main_findings: "Explicit documentation of MSE processes is limited.
-<br> <br> Climate change MSEs have been less explict, more scientist driven, included fewer objectives -- with less consideration of economic or social objectives -- and fewer drivers."
+main_findings: "Explicit documentation of management strategy evaluation processes is limited.
+<br> <br> Climate change MSEs have been less explicit, more scientist driven, included fewer objectives -- with less consideration of economic or social objectives -- and fewer drivers."
 output: 
@@
 Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives.
-Using the structured decision making (SDM) process -- the decision making framework in which adaptive management occurs -- as our framework, how do published MSE projects utilize standard SDM components and support learning within the MSE practitioner community?
+Using the structured decision making (SDM) process [@R-SDM] -- the decision making framework in which adaptive management occurs -- as our framework, how do published MSE projects utilize standard SDM components and support learning within the MSE practitioner community?
 
@@
 
-```{r explicit, fig.cap='Percentage of MSEs that included explicit documentation of these SDM process steps!', fig.height=6, out.width="100%"}
-text1<-expression(paste("Absent or \nInexplicit"))
+```{r explicit, fig.cap='Percentage of MSEs that included explicit documentation (colored bars) of SDM process steps --- Roles and meetings --- and adoption of results.
+In the remaining MSEs was documentation was inexplicit, or absent, and whether the results were adopted was unknown.', fig.height=6, out.width="100%"}
+text1<-expression(paste("Inexplicit \nor Absent"))
 text2<-paste("Absent")
@@
 
-```{r participants, fig.cap='Who participated during these stages of the SDM process', fig.height=6, out.width="100%"}
+```{r participants, fig.cap='Who guided, participated in, or provided input during the specified steps of the MSE process', fig.height=6, out.width="100%"}
 Part.plot
@@
 
-```{r driversPlot, fig.cap='text', fig.height=6, out.width="100%"}
+```{r driversPlot, fig.cap='Drivers included in MSE models', fig.height=6, out.width="100%"}
 Driver.plot

### 2019-09-03 Poster_ICES.Rmd
Edit header font; edit caption font size; Change name of reviewed to random sample; edit colors; edit text based on feedback

@@
 body_textsize: "29.5pt"
+main_fontfamily: "Rasa"
+caption_textsize: "24pt"
 ---
@@
   mutate(Drivers=str_extract(Drivers, "Climate Change")) %>%
-  mutate(Drivers=replace_na(Drivers, "Other"))
+  mutate(Drivers=replace_na(Drivers, "Random Sample"))
 
 # label MSEs articles by analysis for ploting
-freq.data_pub<-mutate(freq.data_pub,Analysis="Reviewed")
+freq.data_pub<-mutate(freq.data_pub,Analysis="Random Sample")
 freq.data_CC<-mutate(freq.data_CC,Analysis="Climate Change")
 freq.data<-rbind(freq.data_pub,freq.data_CC)
-part.data_pub<-mutate(part.data_pub,Analysis="Reviewed")
+part.data_pub<-mutate(part.data_pub,Analysis="Random Sample")
 part.data_CC<-mutate(part.data_CC,Analysis="Climate Change")
@@
                    32,33,34,35,36,37,38,39,39)
-drive.data_pub<-mutate(drive.data_pub,Analysis="Reviewed") %>% 
+drive.data_pub<-mutate(drive.data_pub,Analysis="Random Sample") %>% 
   mutate(Percent=round(Percent,0)) %>% 
@@
                    16,6)
-# Wrangle objective category data by reviewed, climate change, and combined
-objcat.data_pub<-mutate(objcat.data_pub,Analysis="Reviewed") %>% 
+# Wrangle objective category data by Random Sample, climate change, and combined
+objcat.data_pub<-mutate(objcat.data_pub,Analysis="Random Sample") %>% 
   mutate(Percent=round(Percent,0)) %>% 
@@
 
-# Wrangle objective data by reviewed, climate change, and combined
-obj.dataTable_pub<-mutate(obj.dataTable_pub,Analysis="Reviewed") %>% 
+# Wrangle objective data by Random Sample, climate change, and combined
+obj.dataTable_pub<-mutate(obj.dataTable_pub,Analysis="Random Sample") %>% 
   mutate(Percent=round(Percent,0))
@@
 
-altcat.data_pub<-mutate(altcat.data_pub,Analysis="Reviewed") %>% 
+altcat.data_pub<-mutate(altcat.data_pub,Analysis="Random Sample") %>% 
   mutate(Percent=round(Percent,0))
@@
 MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
-  geom_point(size=2) + scale_color_manual(values=c("#FF0033","#0000FF")) +  theme_void() +
+  geom_point(size=2.5) + scale_color_manual(values=c("#66CCCC","#006666")) +  theme_void() +
   theme(legend.position = c(0.15, 0.25))
@@
 # plot explicit documentation of steps or components of MSE processes
-# Reviewed MSEs
+# Random Sample MSEs
 Freq.plot<-ggplot(freq.data,aes(Explicit,Percent,fill=Analysis))+
@@
   geom_vline(xintercept=1.5,linetype="dashed")+
-  coord_flip()+scale_fill_manual(values=c("#FF0033","#0000FF")) +
+  coord_flip()+scale_fill_manual(values=c("#66CCCC","#006666")) +
   scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
@@
   guides(fill = guide_legend(reverse=T)) + theme_bw() +
-  theme(legend.position = c(0.875,0.085),legend.title = element_blank())
+  theme(legend.position = c(0.85,0.07),legend.title = element_blank())
 
@@
   facet_wrap(~Stage,scale="free",ncol=2) + geom_col(position="dodge") +
-  scale_fill_manual(values=c("#FF0033","#0000FF")) +
+  scale_fill_manual(values=c("#66CCCC","#006666")) +
   scale_x_continuous(breaks = part.data$order,
@@
   guides(fill = guide_legend(reverse=T)) + theme_bw() +
-  theme(legend.position = c(0.875,0.085),legend.title = element_blank())
+  theme(legend.position = c(0.86,0.075),legend.title = element_blank())
 
@@
   geom_col(position="dodge") +
-  scale_fill_manual(values=c("#FF0033","#0000FF")) +
+  scale_fill_manual(values=c("#66CCCC","#006666")) +
   scale_x_continuous(breaks = drive.data$order,
@@
   guides(fill = guide_legend(reverse=T)) + theme_bw() +
-  theme(legend.position = c(0.85,0.085),legend.title = element_blank())
+  theme(legend.position = c(0.825,0.075),legend.title = element_blank())
 ```
 
-# Introduction
-For most of its history fisheries management has focused on fishing impacts with ecosystem status viewed as a background constant.
-The increasing rate and degree of climate change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
+###
+Fisheries management has mostly focused on fishing impacts with ecosystem status viewed as a background constant.
 
-Management strategy evaluation (MSE) is commonly considered the gold standard of fisheries management approaches, and is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@R-Punt].
-Thus, MSE is a compelling tool to address the impact of climate change on fisheries management.
+The increasing rate of climate change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
 
-Despite numerous applications of MSE, the applications that include climate change are limited, leaving uncertainty about the best approach to MSE use for fisheries management that accounts for a changing environment.
+Management strategy evaluation (MSE) is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@R-Punt].
+Thus, MSE is a compelling tool to assess climate change impacts and test climate-ready options for fisheries management decisions.
 
 Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives.
-Using the structured decision making (SDM) process [@R-SDM] -- the decision making framework in which adaptive management occurs -- as our framework, how do published MSE projects utilize standard SDM components and support learning within the MSE practitioner community?
+We used the structured decision making (SDM) process [@R-SDM] -- the decision making framework in which adaptive management occurs -- as our framework, how do published MSE projects utilize standard SDM components and support learning within the MSE practitioner community?
+
+<center>
+![SDM process](SDM process.png)
 
@@
 
-## Objectives
+### Objectives
 We review published MSEs against the steps of the SDM framework to:
 
-* Determine the rate of documentation and use of steps in the SDM process.
+* Determine the rate of documentation, and use of steps, in the MSE process:  
 * Identify participants in MSE processes.
@@
 
-# Methods
-#### Published MSE article selection
+### MSE Literature Review Methods
 We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019.
 This search returned 253 results.
-
-We took a random sample of these 253 articles and reviewed 30 that document a MSE.
-Not all articles in the potential pool of 253 articles were MSEs due to articles that are reviews, meta-analyses, etc., or simply cite other MSE articles.
-Of the 253 articles, ### documented a MSE, and 11 included climate change as a driver of fishery status.
+We reviewed a random sample of 30 articles that document a MSE, removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
+Of the 253 articles, 11 included climate change as a driver of fishery status, and after removing other articles we estimate 110-160 articles document a MSE.
 
@@
 |:--:|:--:|
-| Published | ### |
-| Reviewed Sample | 30 |
+| Published (estimated) | 110-160 |
+| Random Sample | 30 |
 | Climate Change | 11 |  
+<br>
+<br>
 
-</br>
-</br>
+### Where have climate change MSEs occurred?
+To date MSEs modeling climate change have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).
 
-# Results
-Apart from one each in Europe and the Western Pacific, to date MSEs modeling climate change have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).
-
-```{r, mapfigure, fig.cap='Map of MSE study locations.', out.width="100%"}
-MSE.map
+```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=4, out.width="100%"}
+MSE.map +
+  theme(text = element_text(size=14))
 ```
 
-The vast majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Figure \@ref(fig:explicit)).
+### Are structured decision making steps explicit in MSEs?
+The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Figure \@ref(fig:explicit)).
 
 ```{r explicit, fig.cap='Percentage of MSEs that included explicit documentation (colored bars) of SDM process steps --- Roles and meetings --- and adoption of results.
-In the remaining MSEs was documentation was inexplicit, or absent, and whether the results were adopted was unknown.', fig.height=6, out.width="100%"}
+In the remaining MSEs was documentation was inexplicit, or absent, and whether the results were adopted was unknown.', fig.height=5.5, out.width="100%"}
 text1<-expression(paste("Inexplicit \nor Absent"))
@@
 text3<-paste("Unknown")
-Freq.plot + annotate("text", x = 5, y = 55, label = text1, size=25) + 
+Freq.plot + annotate("text", x = 8.75, y = 50, label = 'underline("Else")', size=20, parse=TRUE) +
+  annotate("text", x = 5, y = 52, label = text1, size=20) + 
   annotate("text", x = 2.6, y = 50, label = text2, size=20) + 
-  annotate("text", x = 1, y = 50, label = text3, size=12)
+  annotate("text", x = 1, y = 50, label = text3, size=12) +
+  theme(text = element_text(size=14))
 ```
 
-The primary participants and leaders of the MSE processes were scientists, while some processes were lead by governments and management agencies (Figure \@ref(fig:participants)).
-Where explicit documentation occurred objectives and alternatives were elicited from a variety of participants, with scientists seemingly selecting these components in the inexplicit cases.
+### Who is involved in MSEs?
+The primary participants and leaders of the MSE processes were scientists, with some lead by governments and management agencies (Figure \@ref(fig:participants)).
+Objectives and alternatives were elicited from a variety of participants, with scientists seemingly selecting these components in cases where documentation was inexplicit.
 
 ```{r participants, fig.cap='Who guided, participated in, or provided input during the specified steps of the MSE process', fig.height=6, out.width="100%"}
-Part.plot
+Part.plot +
+  theme(text = element_text(size=14))
 ```
 
-General uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the reviewed MSEs (Figure \@ref(fig:driversPlot)).
+### What questions and management objectives do MSEs address?
+Uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the random sample of MSEs (Figure \@ref(fig:driversPlot)).
 
 ```{r driversPlot, fig.cap='Drivers included in MSE models', fig.height=6, out.width="100%"}
-Driver.plot
+Driver.plot +
+  theme(text = element_text(size=14))
 ```
 
-A greater number of objectives were considered per MSE in the random sample of reviewed MSEs than in the climate change MSEs (Table \@ref(tab:PerMSE)).
-Climate change MSEs were less likely to consider economic and social objectives than the MSEs sample for review (Table \@ref(tab:objectiveCategories)).
+More objectives were considered per MSE in the random sample than in the climate change MSEs (Table \@ref(tab:PerMSE)).
+Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:objectiveCategories)).
 
 ```{r, PerMSE}
-color_Pub <- which(per.MSE$Analysis=='Reviewed')
+color_Pub <- which(per.MSE$Analysis=='Random Sample')
 color_CC <- which(per.MSE$Analysis=='Climate Change')
@@
   kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%  
-  row_spec(color_CC, color = "#FF0033", background = "white") %>%
-  row_spec(color_Pub, color = "#0000FF", background = "white")
+  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
+  row_spec(color_Pub, color = "#006666", background = "white")
 ```
@@
 ```{r, objectiveCategories}
-color_Pub <- which(objcat.data$Analysis=='Reviewed')
+color_Pub <- which(objcat.data$Analysis=='Random Sample')
 color_CC <- which(objcat.data$Analysis=='Climate Change')
@@
   kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>%
-  row_spec(color_CC, color = "#FF0033", background = "white") %>%
-  row_spec(color_Pub, color = "#0000FF", background = "white")
+  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
+  row_spec(color_Pub, color = "#006666", background = "white")
 ```
 
-# Conclusions
+### Main Conclusions
 The picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes. 
-While documentation about the modeling conducted in a MSE was explicit, documentation about the process used to make decision about the MSE model were limited.
+While documentation about the **modeling** conducted in a MSE was explicit, documentation about the **process used to make decisions** about the MSE model **were limited**.
 Scientists participated in, or seemingly directly selected the objectives and alternatives to evaluate in most MSEs.
 Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.
-Conservation, Economic, and Yield focused objectives were commonly considered, while broader social objectives were rarely considered.
+Conservation, Economic, and Yield focused objectives were commonly considered, while **broader social objectives were rarely considered**.
 
-To date climate change has not been considered in many MSEs.
-MSEs modeling climate change have been less explicitly documented, more likely to be scientist driven, and included fewer system drivers and fewer objectives, with less consideration of economic or social objectives.
+Despite numerous applications of MSE, the applications that include climate change are limited, leaving uncertainty about the best approach to MSE use for fishereis management that accounts for a changing environment.
 
-# More Details
-Check out our shiny app (<https://jonathancummings.shinyapps.io/MSEreview/>) and keep an eye out for the forthcoming publication.
+### Explore our review on the web!
+Check out our shiny app (<https://jonathancummings.shinyapps.io/MSEreview/>).
 
-# References
+### References

### 2019-09-09 app.R
Us saved workspace for data

@@ -25,22 +25,24 @@ library(DBI) # database interface
 ## cloud_sql_proxy -instances=alpine-tracker-230222:us-east1:mse-review=tcp:3306
 
 # Create connection function
-getSqlConnection <- function(){
-  cn <-
-    dbConnect(
-      RMySQL::MySQL(),
-      dbname = 'MSEreview',
-      host = '127.0.0.1',
-      port=3306,
-      username = 'root',
-      password = 'tLenKdQCs9M2pqK')
-  return(cn)
-}
-# Open connection
-con <- getSqlConnection()
+# getSqlConnection <- function(){
+#   cn <-
+#     dbConnect(
+#       RMySQL::MySQL(),
+#       dbname = 'MSEreview',
+#       host = '127.0.0.1',
+#       port=3306,
+#       username = 'root',
+#       password = 'tLenKdQCs9M2pqK')
+#   return(cn)
+# }
+# # Open connection
+# con <- getSqlConnection()
 
 # To close the connection
 # dbDisconnect(con)
+
+load("MSEreview.RData")
 
 ##### 
 # Script
@@ -52,20 +54,20 @@ world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a
 #####
 ###-- Data Processing --###
 
-# Obtain data tables
-study<-dbReadTable(con,"tblStudy")
-# study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
-mgmt<-dbReadTable(con,"tblStudyManagement")
-# mgmt<-read_xlsx("DB files - Excel/tblStudyManagementTools.xlsx")
-# mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
-obj<-dbReadTable(con,"tblStudyObjectives")
-# obj<-read_xlsx("DB files - Excel/tblStudyObjectives.xlsx")
-# obj$fkStudyID<-parse_integer(obj$fkStudyID)
-fields<-dbReadTable(con,"tblStudyFields")
-# fields<-read_xlsx("DB files - Excel/tblStudyFields.xlsx")
-
-# To close the connection
-dbDisconnect(con)
+# # Obtain data tables
+# study<-dbReadTable(con,"tblStudy")
+# # study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
+# mgmt<-dbReadTable(con,"tblStudyManagement")
[…]

### 2019-09-09 app.R
edit split from ", " to ","

@@ -303,7 +303,7 @@ server <- function(input, output, session) {   # code to create output using ren
              "Sub Alternatives"="ProcedureElicitation_Sub")
   })
   part.data_reviewed2<- reactive({part.data_reviewed() %>%
-    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+    purrr::map(~ strsplit(as.character(.),split=",")) %>%
     purrr::map(unlist) %>%
     purrr::map(table)
   })
@@ -345,7 +345,7 @@ server <- function(input, output, session) {   # code to create output using ren
   # What drivers are considered
   drive.data<-reactive({data_reviewed() %>%
     select(Drivers) %>%
-    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+    purrr::map(~ strsplit(as.character(.),split=",")) %>%
     purrr::map(unlist) %>%
     purrr::map(table) %>%
     plyr::ldply(data.frame) %>%
@@ -358,7 +358,7 @@ server <- function(input, output, session) {   # code to create output using ren
   # What Objective types are considered
   objcat.data<-reactive({data_reviewed() %>%
     select(ObjectiveCategories) %>%
-    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+    purrr::map(~ strsplit(as.character(.),split=",")) %>%
     purrr::map(unlist) %>%
     purrr::map(table) %>%
     plyr::ldply(data.frame) %>%
@@ -401,7 +401,7 @@ server <- function(input, output, session) {   # code to create output using ren
   # What Alternative types are considered
   altcat.data<-reactive({data_reviewed() %>%
     select(ManagementType) %>%
-    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+    purrr::map(~ strsplit(as.character(.),split=",")) %>%
     purrr::map(unlist) %>%
     purrr::map(table) %>%
     plyr::ldply(data.frame) %>%

### 2019-09-10 app.R
edit table wrangling and column select

@@ -88,22 +88,45 @@ obj.join<-obj %>%
 
 # Join tables
 data<-full_join(study,mgmt.join,by=c("ID"="fkStudyID"))
-data<-full_join(data,obj.join,by=c("ID"="fkStudyID")) %>% 
-  select(-c("ID"))
-obj.data<-right_join(study.join,obj,by=c("ID"="fkStudyID"))%>% 
-  select(-c("ID","ID.y"))
+data<-full_join(data,obj.join,by=c("ID"="fkStudyID"))
+obj.data<-right_join(study.join,obj,by=c("ID"="fkStudyID"))
 
 # Move comment column to the end
 data<-data %>% 
   select(-'Comments', 'Comments')
+
+# Filter data to create separate pub and climate change objects
+data_pub<-filter(data,IncludeInPublication==TRUE)
+data_CC<-filter(data,str_detect(Drivers,"Climate Change"))
+
+data_pub.join<-data_pub %>%
+  select(ID,Citation)
+data_CC.join<-data_CC %>%
+  select(ID,Citation)
+
+obj.data_pub<-left_join(data_pub.join,obj,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID","ID.y"))
+obj.data_CC<-left_join(data_CC.join,obj,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID","ID.y"))
+
+alt.data_pub<-left_join(data_pub.join,mgmt,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID","ID.y"))
+alt.data_CC<-left_join(data_CC.join,mgmt,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID","ID.y"))
+
+data_pub<-data_pub %>%
+  select(-c("ID"))
+data_CC<-data_CC %>%
+  select(-c("ID"))
+
+data<-select(data,-c("ID"))
+obj.data<-select(obj.data,-c("ID","ID.y"))
 
 # Get columns whose width needs editing
 targets<-match(c("FullCitation","Comments"),names(data))
 
 # Get columns for study summary
 summary.col<-c("Citation",
-               "Authors",
-               "YearPub",
                "Species",
                "Location",
                "System")
@@ -141,7 +164,8 @@ alt.col<-c("ManagementType",
 # Get columns for map
 map.col<-c("Latitude",
            "Longitude",
-           "Citation")
+           "Citation",
+           "Drivers")
 
 #####
 ###-- Shiny App --###
[…]

### 2019-09-10 Poster_ICES.Rmd
fix typo

@@
 ```{r explicit, fig.cap='Percentage of MSEs that included explicit documentation (colored bars) of SDM process steps --- Roles and meetings --- and adoption of results.
-In the remaining MSEs was documentation was inexplicit, or absent, and whether the results were adopted was unknown.', fig.height=5.5, out.width="100%"}
+In the remaining MSEs documentation was inexplicit, or absent, and whether the results were adopted was unknown.', fig.height=5.5, out.width="100%"}
 text1<-expression(paste("Inexplicit \nor Absent"))

### 2019-09-10 app.R
Add background tab with introductory text and sdm image

@@ -261,6 +261,28 @@ ui <- fluidPage(
              hr(),
              h2("Problem Components"),
              DT::dataTableOutput("MSE.problem")
+             ),
+    tabPanel("Background and methods",
+             h2("Background"),
+             p("Fisheries management has mostly focused on fishing impacts
+  with ecosystem status viewed as a background constant. The increasing rate
+  of climate change is changing this dynamic, bringing ecosystem status to
+  the forefront of fisheries management. Management strategy evaluation (MSE)
+  is 'widely considered to be the most appropriate way to evaluate the
+  trade‐offs achieved by alternative management strategies and to assess the
+  consequences of uncertainty for achieving management goals' (Punt et al.
+  2014). Thus, MSE is a compelling tool to assess climate change impacts and
+  test climate-ready options for fisheries management decisions Adaptive
+  management arose to address uncertainties and accelerate progress towards
+  meeting management objectives. We used the structured decision making (SDM)
+  process - the decision making framework in which adaptive management
+  occurs - as our framework, how do published MSE projects utilize standard
+  SDM components and support learning within the MSE practitioner community?"),
+          imageOutput("imageSDM"),
+          p(em("Structured Decision Making Process")),
+             h2("Study Summaries"),
+             hr(),
+             h2("Problem Components")
     )
   )
 )
@@ -530,6 +552,12 @@ server <- function(input, output, session) {   # code to create output using ren
   output$MSE.problem <- DT::renderDataTable({
     arrange(prob.data(),Citation)
   })
+  # Tab 6
+  output$imageSDM<-renderImage({
+    filename<-normalizePath(file.path('./www',paste("SDMprocess.png")))
+    list(src=filename,
+         width=500,
+         height=350)},deleteFile = FALSE)
 }
 
 shinyApp(ui, server)

### 2019-09-10 app.R
improve background tab format

@@ -265,20 +265,25 @@ ui <- fluidPage(
     tabPanel("Background and methods",
              h2("Background"),
              p("Fisheries management has mostly focused on fishing impacts
-  with ecosystem status viewed as a background constant. The increasing rate
-  of climate change is changing this dynamic, bringing ecosystem status to
-  the forefront of fisheries management. Management strategy evaluation (MSE)
-  is 'widely considered to be the most appropriate way to evaluate the
-  trade‐offs achieved by alternative management strategies and to assess the
-  consequences of uncertainty for achieving management goals' (Punt et al.
-  2014). Thus, MSE is a compelling tool to assess climate change impacts and
-  test climate-ready options for fisheries management decisions Adaptive
-  management arose to address uncertainties and accelerate progress towards
-  meeting management objectives. We used the structured decision making (SDM)
-  process - the decision making framework in which adaptive management
-  occurs - as our framework, how do published MSE projects utilize standard
-  SDM components and support learning within the MSE practitioner community?"),
-          imageOutput("imageSDM"),
+                with ecosystem status viewed as a background constant."),
+             p("The increasing rate of climate change is changing this
+     dynamic, bringing ecosystem status to the forefront of fisheries
+               management."),
+             p("Management strategy evaluation (MSE) is 'widely considered to
+     be the most appropriate way to evaluate the trade‐offs achieved by
+     alternative management strategies and to assess the consequences of
+     uncertainty for achieving management goals' (Punt et al. 2014). Thus,
+     MSE is a compelling tool to assess climate change impacts and test
+     climate-ready options for fisheries management decisions Adaptive
+     management arose to address uncertainties and accelerate progress
+     towards meeting management objectives. We used the structured decision
+     making (SDM) process - the decision making framework in which adaptive
+     management occurs - as our framework, how do published MSE projects
+     utilize standard SDM components and support learning within the MSE
+     practitioner community?"),
+          imageOutput("imageSDM",
+                      width=400,
+                      height=275),
           p(em("Structured Decision Making Process")),
              h2("Study Summaries"),
              hr(),
@@ -556,8 +561,8 @@ server <- function(input, output, session) {   # code to create output using ren
   output$imageSDM<-renderImage({
     filename<-normalizePath(file.path('./www',paste("SDMprocess.png")))
     list(src=filename,
-         width=500,
-         height=350)},deleteFile = FALSE)
+         width=400,
+         height=275)},deleteFile = FALSE)
 }
 
 shinyApp(ui, server)

### 2019-09-10 app.R
Add methods to background tab

@@ -173,6 +173,7 @@ map.col<-c("Latitude",
 ui <- fluidPage(
   theme = shinytheme("readable"),
   tabsetPanel(
+    #### Tab 1: Filter Results ####
     tabPanel("Filter Results", 
              headerPanel("An Assessment of Management Strategy Evaluations"),# title of the page
              h2("MSE literature review results"),
@@ -205,6 +206,7 @@ ui <- fluidPage(
                )
              )
              ),
+    #### Tab 2: Results - Plots ####
     tabPanel("Results - Plots", 
              headerPanel("An Assessment of Management Strategy Evaluations"),# title of the page
              h2("MSE literature review results"),
@@ -221,6 +223,7 @@ ui <- fluidPage(
              plotOutput("part.plot"),
              p("Figure 3. Participation at different stages of the MSE process")
              ),
+    #### Tab 3: Results - Tables ####
     tabPanel("Results - Tables", 
              h2("Explict Process Documentation"),
              p("Table 1. The number and percentage of MSE processes that explicitly included each step in the process"),
@@ -243,6 +246,7 @@ ui <- fluidPage(
              p("Table 5. Number of, percentage of objectives, and Frequency per MSE of objective usage"),
              tableOutput("MSE.alt")
              ),
+    #### Tab 4: Data - All ####
     tabPanel("Data - All",
              h1("All of the MSE literature review data"),
              h2("Study data"),
@@ -254,6 +258,7 @@ ui <- fluidPage(
              h2("Field Descriptions"),
              tableOutput("MSE.Fields")
              ),
+    #### Tab 5: Data Tables - Subsets ####
     tabPanel("Data Tables - Subsets",
              h1("Selected columns of MSE literature review data"),
              h2("Study Summaries"),
@@ -262,32 +267,40 @@ ui <- fluidPage(
              h2("Problem Components"),
              DT::dataTableOutput("MSE.problem")
              ),
+    #### Tab 6: Background and methods ####
     tabPanel("Background and methods",
-             h2("Background"),
+             h4("Background"),
              p("Fisheries management has mostly focused on fishing impacts
                 with ecosystem status viewed as a background constant."),
              p("The increasing rate of climate change is changing this
-     dynamic, bringing ecosystem status to the forefront of fisheries
+      dynamic, bringing ecosystem status to the forefront of fisheries
                management."),
              p("Management strategy evaluation (MSE) is 'widely considered to
[…]

### 2019-09-12 app.R
edit section breaks

@@ -2,8 +2,7 @@
 # Literature Review Data Analysis
 # 3/8/2019, JWC
 
-##### 
-# Meta Data and setup
+##### Meta Data and setup #####
 # Description: Import and analyze MSE documentation from database
 # Value: TBD
 
@@ -44,15 +43,12 @@ library(DBI) # database interface
 
 load("MSEreview.RData")
 
-##### 
-# Script
 
 # Get map background for plotting the map
 world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
 
 
-#####
-###-- Data Processing --###
+##### Data Processing #####
 
 # # Obtain data tables
 # study<-dbReadTable(con,"tblStudy")
@@ -69,7 +65,7 @@ world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a
 # # To close the connection
 # dbDisconnect(con)
 
-#####
+
 # Edit tables for joining
 study<-study %>% 
   unite(Citation,c(Authors,YearPub),sep=" ",remove=F)
@@ -167,8 +163,7 @@ map.col<-c("Latitude",
            "Citation",
            "Drivers")
 
-#####
-###-- Shiny App --###
+##### Shiny App #####
 
 ui <- fluidPage(
   theme = shinytheme("readable"),

### 2019-09-12 app.R
Move the items displayed in the "background and methods" tab into the first tab now called "About" and rearrange the data filter radio buttons.

@@ -168,18 +168,56 @@ map.col<-c("Latitude",
 ui <- fluidPage(
   theme = shinytheme("readable"),
   tabsetPanel(
-    #### Tab 1: Filter Results ####
-    tabPanel("Filter Results", 
-             headerPanel("An Assessment of Management Strategy Evaluations"),# title of the page
-             h2("MSE literature review results"),
+    #### Tab 1: About ####
+    tabPanel("About", 
+             h1("An Assessment of Management Strategy Evaluations"),# title of the page
              hr(),
-             radioButtons("data_filter",
-                          "Show results from:",
-                          choices = c(
-                                      "MSEs reviewed for Cummings et. al. Publication"="pub",
-                                      "MSE included climate change as a driver"="CC",
-                                      "All MSEs"="all")),
-             textOutput("radio"),
+             h4("Background"),
+             p("Fisheries management has mostly focused on fishing impacts
+               with ecosystem status viewed as a background constant."),
+             p("The increasing rate of climate change is changing this
+               dynamic, bringing ecosystem status to the forefront of fisheries
+               management."),
+             p("Management strategy evaluation (MSE) is 'widely considered to
+               be the most appropriate way to evaluate the trade‐offs achieved by
+               alternative management strategies and to assess the consequences of
+               uncertainty for achieving management goals' (Punt et al. 2014). Thus,
+               MSE is a compelling tool to assess climate change impacts and test
+               climate-ready options for fisheries management decisions Adaptive
+               management arose to address uncertainties and accelerate progress
+               towards meeting management objectives. We used the structured decision
+               making (SDM) process - the decision making framework in which adaptive
+               management occurs - as our framework, how do published MSE projects
+               utilize standard SDM components and support learning within the MSE
+               practitioner community?"),
+             imageOutput("imageSDM",
+                         width=400,
+                         height=275),
+             p(em("Structured Decision Making Process")),
+             h4("MSE Review Methods"),
+             p("We conducted our search for MSEs in the SCI-EXPANDED index from
+               Web of Science, searching for “management strategy evaluation” by topic
+               across all years on January 8th, 2019. This search returned 253
+               results. We reviewed a random sample of 30 articles that document a
+               MSE, removing articles that were reviews, meta-analyses, or simply
+               cited other MSE articles from our sample. Of the 253 articles, 11
+               included climate change as a driver of fishery status, and after
[…]

### 2019-09-19 MSE manuscript.Rmd
setup of document

--- 2019-09-10 Poster_ICES.Rmd
@@
 ---
-main_topsize: 0.2 #percent coverage of the poster
-main_bottomsize: 0
-#ESSENTIALS
-title: Use of management strategy evaluation to inform fisheries management in a changing climate
+title: 'No Replication Without Process Documentation:  An Assessment of Management
+  Strategy '
 author:
-  - name: '**Jonathan W. Cummings**'
-    affil: 1
-    main: true
-    orcid: '0000-0001-8028-5787'
-    twitter: jonwcummings 
-    email: 'jcummings@umassd.edu'
-  - name: Amanda Hart
-    affil: 1
-    main: true
-    orcid: '0000-0002-6449-9328'
-    email: 'ahart2@umassd.edu'
-  - name: Gavin Fay
-    main: true
-    affil: 1
-    email: 'gfay@umassd.edu'
-affiliation:
-  - num: 1
-    address: School for Marine Science and Technology, University of Massachusetts - Dartmouth 836 South Rodney French Blvd,New Bedford, MA, USA 02744
-main_findings: "Explicit documentation of management strategy evaluation processes is limited.
-<br> <br> Climate change MSEs have been less explicit, more scientist driven, included fewer objectives -- with less consideration of economic or social objectives -- and fewer drivers."
+- affiliation: School for Marine Science and Technology, University of Massachusetts - Dartmouth
+  name: Jonathan W. Cummings
+date: "`r format(Sys.time(), '%B %d, %Y')`"
 output:
-  posterdown::posterdown_betterport:
-    self_contained: false
-    pandoc_args: --mathjax
-    number_sections: false
+  html_document:
+    df_print: paged
+  pdf_document:
+    citation_package: natbib
+    fig_caption: yes
+    keep_tex: yes
+    latex_engine: pdflatex
+    template: svm-latex-ms.tex
+biblio-style: apsr
+endnote: no
+fontfamily: mathpazo
+fontsize: 11pt
+geometry: margin=1in
+header-includes: \usepackage{hyperref}
+keywords: pandoc, r markdown, knitr
 bibliography: MSEreview.bib
-link-citations: true
-nocite: '@*'
-title_textsize: "50pt"
-main_textsize: "90pt"
-column_numbers: 3
-body_textsize: "29.5pt"
-main_fontfamily: "Rasa"
-caption_textsize: "24pt"
+abstract: Enter abstract text here
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(cache=TRUE,
+                      message=FALSE, warning=FALSE,
+                      fig.path='figs/',
+                      cache.path = '_cache/',
+                      fig.process = function(x) {
+                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
+                      if (file.rename(x, x2)) x2 else x
+                      })
+```
+
 ```{r, include=FALSE}
@@
 
-###
+# Introduction
 Fisheries management has mostly focused on fishing impacts with ecosystem status viewed as a background constant.
@@
 
-<center>
-![SDM process](SDM process.png)
+# Methods
 
-```{r, include=FALSE}
-knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
+# Results
+
+# Discussion
+
+Academic workflow, certainly in political science, is at a crossroads.
+The *American Journal of Political Science* (*AJPS*) announced a (my words) ["show your work" initiative](http://ajps.org/2015/03/26/the-ajps-replication-policy-innovations-and-revisions/) in which authors who are tentatively accepted for publication at the journal must hand over the raw code and data that produced the results shown in the manuscript.
+The editorial team at *AJPS* then reproduces the code from the manuscript.
+Pending successful replication, the manuscript moves toward publication.
+The *AJPS* might be at the fore of this movement, and it could be the most aggressive among political science journals, but other journals in our field have signed the joint [Data Access & Research Transparency](http://www.dartstatement.org/) (DART) initiative.
+This, at a bare minimum, requires uploading code from quantitatively-oriented published articles to in-house directories hosted by the journal or to services like [Dataverse](http://dataverse.org/).
+
+There are workflow implications to the Lacour controversy as well.
+Political science, for the foreseeable future, will struggle with the extent of [the data fraud perpetrated by Michael Lacour](http://stanford.edu/~dbroock/broockman_kalla_aronow_lg_irregularities.pdf) in an article co-authored with Donald P.
+Green in *Science*, the general scientific journal of record in the United States.
+A failure to reproduce LaCour's results with different samples uncovered a comprehensive effort by LaCour to "fake" data that provided results to what we felt or believed to be true [(i.e.
+"truthiness")](http://chronicle.com/article/LAffaire-LaCour/230905/).
+However, [fake data can have real consequences](http://kieranhealy.org/blog/archives/2015/05/20/fake-science-real-consequences/) for both the researcher and those who want to learn from it and use it for various purposes.
+Even research done honestly may suffer the same fate if researchers are not diligent in their workflow.
+
+These recent events underscore the DART push and cast a shadow over our workflow.
+However, good workflow has always been an issue in our discipline.
+Cloud storage services like [Dropbox](http://www.dropbox.com) are still relatively new among political scientists.
+Without cloud storage, previous workflow left open the possibility that work between a home computer and an office computer was lost as a function of a corrupted thumb drive, an overheated power supply, or, among other things, the wave of viruses that [would particularly affect Microsoft users every summer](http://money.cnn.com/2003/11/05/technology/microsoftbounty/).
+Social sciences, [unlike engineering](http://kieranhealy.org/blog/archives/2014/01/23/plain-text/), have traditionally relied on software like Microsoft Word for manuscript preparation though any word processor reduces workflow to a series of clicks and strokes on a keyboard.
+This is [a terrible way to track changes](http://www.nytimes.com/2013/04/19/opinion/krugman-the-excel-depression.html) or maintain version control.
+The addition of collaborators only compounds all the aforementioned issues.
+The proverbial left hand may not know what the right hand is doing.
+
+I think there is reason for optimism.
+We only struggle with it now because we have tools like [R Markdown](http://rmarkdown.rstudio.com/) and [Pandoc](http://pandoc.org/), more generally, that make significant strides in workflow.
+LaTeX resolved earlier issues of corrupted binary files by reducing documents to raw markup that was little more than raw text and revisions that could be easily kept as ["commented" text](http://tex.stackexchange.com/questions/11177/how-to-write-hidden-notes-in-a-latex-file).
+However, for all its benefits (including pretty PDFs), [LaTeX is *ugly* code](http://www-rohan.sdsu.edu/~aty/bibliog/latex/gripe.html) and does not provide means of seamlessly working with the actual data analysis itself.
+R Markdown both eliminates markup and allows the author and her collaborators to write and reproduce the manuscript in one fell swoop.
+
+# Getting Started with YAML
+
+The lion's share of a R Markdown document will be raw text, though the front matter may be the most important part of the document.
+R Markdown uses [YAML](http://www.yaml.org/) for its metadata and the fields differ from [what an author would use for a Beamer presentation](http://svmiller.com/blog/2015/02/moving-from-beamer-to-r-markdown/).
+I provide a sample YAML metadata largely taken from this exact document and explain it below.
+
+```{r eval=FALSE}
+---
+output: 
+  pdf_document:
+    citation_package: natbib
+    keep_tex: true
+    fig_caption: true
+    latex_engine: pdflatex
+    template: ~/Dropbox/miscelanea/svm-r-markdown-templates/svm-latex-ms.tex
+title: "A Pandoc Markdown Article Starter and Template"
+thanks: "Replication files are available on the author's Github account..."
+author:
+- name: Steven V. Miller
+  affiliation: Clemson University
+- name: Mary Margaret Albright
+  affiliation: Pendelton State University
+- name: Rembrandt Q. Einstein
+  affiliation: Springfield University
+abstract: "This document provides an introduction to R Markdown, argues for its..."
+keywords: "pandoc, r markdown, knitr"
+date: "`r format(Sys.time(), '%B %d, %Y')`"
+geometry: margin=1in
+fontfamily: mathpazo
+fontsize: 11pt
+# spacing: double
+bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`"
+biblio-style: apsr
+---
 ```
 
-### Objectives
-We review published MSEs against the steps of the SDM framework to:
+`output:` will tell R Markdown we want a PDF document rendered with LaTeX.
+Since we are adding a fair bit of custom options to this call, we specify `pdf_document:` on the next line (with, importantly, a two-space indent).
+We specify additional output-level options underneath it, each are indented with four spaces.
+`citation_package: natbib` tells R Markdown to use `natbib` to handle bibliographic citations.[^natbib] Thereafter, the next line (`keep_tex: true`) tells R Markdown to render a raw `.tex` file along with the PDF document.
+This is useful for both debugging and the publication stage, when the editorial team will ask for the raw `.tex` so that they could render it and later provide page proofs.
+The next line `fig_caption: true` tells R Markdown to make sure that whatever images are included in the document are treated as figures in which our caption in brackets in a Markdown call is treated as the caption in the figure.
+The next line (`latex_engine: pdflatex`) tells R Markdown to use pdflatex and not some other option like `lualatex`.
+For my template, I'm pretty sure this is mandatory.[^pdflatex]
 
-* Determine the rate of documentation, and use of steps, in the MSE process:  
-* Identify participants in MSE processes.
-* Highlight unique aspects of MSEs that model climate change as a driver.
-* Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate  
-* Provide insights about MSE processes and documentation that can further learning to improve future MSEs
+[^natbib]: R Markdown can use Pandoc's native bibliography management system or even `biblatex`, but I've found that it chokes with some of the more advanced stuff I've done with my .bib file over the years.
+For example, I've been diligent about special characters (e.g.
+umlauts and acute accents) in author names in my .bib file, but Pandoc's native citation system will choke on these characters in a .bib file.
+I effectively need `natbib` for my own projects.
+[^pdflatex]: The main reason I still use `pdflatex` (and most readers probably do as well) is because of LaTeX fonts.
+[Unlike others](http://www-rohan.sdsu.edu/~aty/bibliog/latex/gripe.html), I find standard LaTeX fonts to be appealing.
 
-### MSE Literature Review Methods
-We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019.
-This search returned 253 results.
-We reviewed a random sample of 30 articles that document a MSE, removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
-Of the 253 articles, 11 included climate change as a driver of fishery status, and after removing other articles we estimate 110-160 articles document a MSE.
+The next line (`template: ...`) tells R Markdown to use my custom LaTeX template.[^path] While I will own any errors in the code, I confess to "Frankensteining" this template from [the default LaTeX template](https://github.com/jgm/pandoc-templates) from Pandoc, [Kieran Healy's LaTeX template](https://github.com/kjhealy/pandoc-templates/tree/master/templates), and liberally using raw TeX from the [Association for Computing Machinery's (ACM) LaTeX template](https://www.acm.org/publications/article-templates/acm-latex-style-guide).
+I rather like that template since it resembles standard manuscripts when they are published in some of our more prominent journals.
+I will continue with a description of the YAML metadata in the next paragraph, though invite the curious reader to scroll to the end of the accompanying post to see the PDF this template produces.
 
-| MSE type | Count |
-|:--:|:--:|
-| Published (estimated) | 110-160 |
-| Random Sample | 30 |
-| Climate Change | 11 |  
-<br>
-<br>
+[^path]: Notice that the path is relative.
+The user can, if she wishes, install this in the default Pandoc directory.
+I don't think this is necessary.
+Just be mindful of wherever the template is placed.
+Importantly, `~` is used in R to find the home directory (not necessarily the working directory).
+It is equivalent to saying `/home/steve` in Linux, or `/Users/steve` on a Mac, in my case.
 
-### Where have climate change MSEs occurred?
-To date MSEs modeling climate change have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).
+The next fields get to the heart of the document itself.
+`title:` is, intuitively, the title of the manuscript.
+Do note that fields like `title:` do not have to be in quotation marks, but must be in quotation marks if the title of the document includes a colon.
+That said, the only reason to use a colon in an article title is if it is followed by a subtitle, hence the optional field (`subtitle:`).
+Notice I "comment out" the subtitle in the above example with a pound sign since this particular document does not have a subtitle.
+If `thanks:` is included and has an accompanying entry, the ensuing title of the document gets an asterisk and a footnote.
+This field is typically used to advise readers that the document is a working paper or is forthcoming in a journal.
 
-```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=4, out.width="100%"}
-MSE.map +
-  theme(text = element_text(size=14))
+The next field (`author:`) is a divergence from standard YAML, but I think it is useful.
+I will also confess to pilfering this idea from Kieran Healy's template.
+Typically, multiple authors for a given document are separated by an `\and` in this field.
+However, standard LaTeX then creates a tabular field separating multiple authors that is somewhat restrictive and not easy to override.
+As a result, I use this setup (again, taken from Kieran Healy) to sidestep the restrictive rendering of authors in the standard `\maketitle` tag.
+After `author:`, enter `- name:` (no space before the dash) and fill in the field with the first author.
+On the next line, enter two spaces, followed by `affiliation:` and the institute or university affiliation of the first author.
+
+Do notice this can be repeated for however many co-authors there are to a manuscript.
+The rendered PDF will enter each co-author in a new line in a manner similar to journals like *American Journal of Political Science*, *American Political Science Review*, or *Journal of Politics*.
+
+The next two fields pertain to the frontmatter of a manuscript.
+They should also be intuitive for the reader.
+`abstract` should contain the abstract and `keywords` should contain some keywords that describe the research project.
+Both fields are optional, though are practically mandatory.
+Every manuscript requires an abstract and some journals---especially those published by Sage---request them with submitted manuscripts.
+My template also includes these keywords in the PDF's metadata.
+
+`date` comes standard with R Markdown and you can use it to enter the date of the most recent compile.
+I typically include the date of the last compile for a working paper in the `thanks:` field, so this field currently does not do anything in my Markdown-LaTeX manuscript template.
+I include it in my YAML as a legacy, basically.
+
+The next items are optional and cosmetic.
+`geometry:` is a standard option in LaTeX.
+I set the margins at one inch, and you probably should too.
+`fontfamily:` is optional, but I use it to specify the Palatino font.
+The default option is Computer Modern Roman.
+`fontsize:` sets, intuitively, the font size.
+The default is 10-point, but I prefer 11-point.
+`spacing:` is an optional field.
+If it is set as "double", the ensuing document is double-spaced.
+"single" is the only other valid entry for this field, though not including the entry in the YAML metadata amounts to singlespacing the document by default.
+Notice I have this "commented out" in the example code.
+
+The final two options pertain to the bibliography.
+`bibliography:` specifies the location of the .bib file, so the author could make citations in the manuscript.
+`biblio-style` specifies the type of bibliography to use.
+You'll typically set this as APSR.
+You could also specify the relative path of [my *Journal of Peace Research* .bst file](http://svmiller.com/miscellany/journal-of-peace-research-bst-file/) if you are submitting to that journal.
+
+# Getting Started with Markdown Syntax
+
+There are a lot of cheatsheets and reference guides for Markdown (e.g.
+[Adam Prichard](https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet), [Assemble](http://assemble.io/docs/Cheatsheet-Markdown.html), [Rstudio](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf), [Rstudio again](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf), [Scott Boms](http://scottboms.com/downloads/documentation/markdown_cheatsheet.pdf), [Daring Fireball](https://daringfireball.net/projects/markdown/syntax), among, I'm sure, several others).
+I encourage the reader to look at those, though I will retread these references here with a minimal working example below.
+
+```markdown
+
+# Introduction
+
+**Lorem ipsum** dolor *sit amet*.
+
+- Single asterisks italicize text *like this*. 
+- Double asterisks embolden text **like this**.
+
+Start a new paragraph with a blank line separating paragraphs.
+
+- This will start an unordered list environment, and this will be the first item.
+- This will be a second item.
+- A third item.
+    - Four spaces and a dash create a sublist and this item in it.
+- The fourth item.
+
+1. This starts a numerical list.
+2. This is no. 2 in the numerical list.
+
+# This Starts A New Section
+## This is a Subsection
+### This is a Subsubsection
+#### This starts a Paragraph Block.
+
+> This will create a block quote, if you want one.
+Want a table? This will create one.
+
+Table Header  | Second Header
+------------- | -------------
+Table Cell    | Cell 2
+Cell 3        | Cell 4
+
+Note that the separators *do not* have to be aligned.
+
+Want an image? This will do it.
+
+![caption for my image](path/to/image.jpg)
+
+`fig_caption: yes` will provide a caption. Put that in the YAML metadata.
+
+Almost forgot about creating a footnote.[^1] This will do it again.[^2]
+
+[^1]: The first footnote
+[^2]: The second footnote
+
+Want to cite something?
+
+- Find your biblatexkey in your bib file.
+- Put an @ before it, like @smith1984, or whatever it is.
+- @smith1984 creates an in-text citation (e.g. Smith (1984) says...)
+- [@smith1984] creates a parenthetical citation (Smith, 1984)
+
+That'll also automatically create a reference list at the end of the document.
+
+[In-text link to Google](http://google.com) as well.
 ```
 
-### Are structured decision making steps explicit in MSEs?
-The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Figure \@ref(fig:explicit)).
+That's honestly it. Markdown takes the chore of markup from your manuscript (hence: "Markdown").
 
-```{r explicit, fig.cap='Percentage of MSEs that included explicit documentation (colored bars) of SDM process steps --- Roles and meetings --- and adoption of results.
-In the remaining MSEs documentation was inexplicit, or absent, and whether the results were adopted was unknown.', fig.height=5.5, out.width="100%"}
-text1<-expression(paste("Inexplicit \nor Absent"))
-text2<-paste("Absent")
-text3<-paste("Unknown")
-Freq.plot + annotate("text", x = 8.75, y = 50, label = 'underline("Else")', size=20, parse=TRUE) +
-  annotate("text", x = 5, y = 52, label = text1, size=20) + 
-  annotate("text", x = 2.6, y = 50, label = text2, size=20) + 
-  annotate("text", x = 1, y = 50, label = text3, size=12) +
-  theme(text = element_text(size=14))
+On that note, you could easily pass most LaTeX code through Markdown if you're writing a LaTeX document.
+However, you don't need to do this (unless you're using the math environment) and probably shouldn't anyway if you intend to share your document in HTML as well.
+
+# Using R Markdown with Knitr
+
+Perhaps the greatest intrigue of R Markdown comes with the [`knitr` package](http://yihui.name/knitr/) provided by @xie2013ddrk.
+In other words, the author can, if she chooses, do the analysis in the Markdown document itself and compile/execute it in R.
+
+Take, for example, this simple exercise using the `voteincome` data from the `Zelig` package.
+Suppose I want to explain the decision to vote using data from this package.
+I load in the data, clean the data, run the analyses, and present the results as a coefficient plot.
+
+Here's what this code looks like.
+All I did was create a code display, which starts with three *backticks* (i.e.
+those ticks next to the number 1 key on your keyboard) and ends with three backticks on another line.
+On the first line of backticks (i.e.
+to start the code display) enter `{r, eval=FALSE, tidy=TRUE}`.
+The `eval=FALSE` option just displays the R code (and does not run it), `tidy=TRUE` wraps long code so it does not run off the page.
+
+Within that code display, I enter my R code like this.
+
+```{r, eval=FALSE, tidy = TRUE}
+library(stevemisc)
+data(uniondensity)
+M1 <- lm(union ~ left + size + concen, data=uniondensity)
+library(arm)
+coefplot(M1)
 ```
 
-### Who is involved in MSEs?
-The primary participants and leaders of the MSE processes were scientists, with some lead by governments and management agencies (Figure \@ref(fig:participants)).
-Objectives and alternatives were elicited from a variety of participants, with scientists seemingly selecting these components in cases where documentation was inexplicit.
+The implications for workflow are faily substantial.
+Authors can rather quickly display the code they used to run the analyses in the document itself (likely in the appendix).
+As such, there's little guesswork for reviewers and editors in understanding what the author did in the analyses reported in the manuscript.
 
-```{r participants, fig.cap='Who guided, participated in, or provided input during the specified steps of the MSE process', fig.height=6, out.width="100%"}
-Part.plot +
-  theme(text = element_text(size=14))
+It doesn't end there. In fact, here's what happens when `eval=FALSE` is omitted or changed to `eval=TRUE`. Now, the code runs within R. Observe.
+
+```{r, eval=TRUE, tidy = TRUE, cache=FALSE, fig.cap="A Coefficient Plot", message=F, warning=F}
+library(stevemisc)
+data(uniondensity)
+M1 <- lm(union ~ left + size + concen, data=uniondensity)
+library(arm)
+coefplot(M1)
 ```
 
-### What questions and management objectives do MSEs address?
-Uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the random sample of MSEs (Figure \@ref(fig:driversPlot)).
+To get `knitr` to present the results of a table, add `results="asis"` to the brackets to start the R code chunk. The ensuing output will look like this (though the table may come on the next page).
 
-```{r driversPlot, fig.cap='Drivers included in MSE models', fig.height=6, out.width="100%"}
-Driver.plot +
-  theme(text = element_text(size=14))
+```{r, eval=TRUE, tidy = TRUE, size="small", cache=FALSE, results="asis", message=F, warning=F}
+library(stevemisc)
+data(uniondensity)
+library(stargazer)
+M1 <- lm(union ~ left + size + concen, data=uniondensity)
+stargazer(M1, title="A Handsome Table", header=FALSE)
 ```
 
-More objectives were considered per MSE in the random sample than in the climate change MSEs (Table \@ref(tab:PerMSE)).
-Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:objectiveCategories)).
+Adding `echo="FALSE"` inside the brackets to start the R chunk will omit the presentation of the R commands.
+It will just present the table.
+This provides substantial opportunity for authors in doing their analyses.
+Now, the analysis and presentation in the form of a polished manuscript can be effectively simultaneous.[^4]
 
-```{r, PerMSE}
-color_Pub <- which(per.MSE$Analysis=='Random Sample')
-color_CC <- which(per.MSE$Analysis=='Climate Change')
+[^4]: I'm not sure if I'm ready to commit to this myself since my workflow is still largely derived from [Rob J.
+Hyndman's example](http://robjhyndman.com/hyndsight/workflow-in-r/).
+However, *knitr* has endless potential, especially when analyses can stored in cache, saved as chunks, or loaded in the preamble of a document to reference later in the manuscript.
 
-per.MSE %>%
-  kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%  
-  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
-  row_spec(color_Pub, color = "#006666", background = "white")
-```
-
-```{r, objectiveCategories}
-color_Pub <- which(objcat.data$Analysis=='Random Sample')
-color_CC <- which(objcat.data$Analysis=='Climate Change')
-
-objcat.data %>%
-  kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>%
-  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
-  row_spec(color_Pub, color = "#006666", background = "white")
-```
-
-### Main Conclusions
-The picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes. 
-While documentation about the **modeling** conducted in a MSE was explicit, documentation about the **process used to make decisions** about the MSE model **were limited**.
-Scientists participated in, or seemingly directly selected the objectives and alternatives to evaluate in most MSEs.
-Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.
-Conservation, Economic, and Yield focused objectives were commonly considered, while **broader social objectives were rarely considered**.
-
-Despite numerous applications of MSE, the applications that include climate change are limited, leaving uncertainty about the best approach to MSE use for fishereis management that accounts for a changing environment.
-
-### Explore our review on the web!
-Check out our shiny app (<https://jonathancummings.shinyapps.io/MSEreview/>).
-
-### References
+<!--
+# References
+\setlength{\parindent}{-0.2in}
+\setlength{\leftskip}{0.2in}
+\setlength{\parskip}{8pt}
+\vspace*{-0.2in}
+\noindent
+-->

### 2019-09-19 MSE manuscript.Rmd
Edit r code snippet names and break points; Remove roles, open meetings, and adopted from the freq.data analysis; create random sample only freq table and part.plot

@@
 
-```{r, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE,
-                      warning = FALSE,
-                      tidy = FALSE,
-                      message = FALSE,
-                      fig.align = 'center',
-                      out.width = "100%")
+```{r load data, include=FALSE}
 options(knitr.table.format = "html")
@@
   select(-'Comments', 'Comments')
+```
 
-# Filter data
+```{r filter data, include=FALSE}
+
+# Filter data by study type
 data_pub<-filter(data,IncludeInPublication==TRUE)
@@
 
-#####-- Data analysis --#####
+#####-- Filter Data by analysis --#####
 # Get columns for study summary
@@
             "TradeOffsExplicit",
-            "DecisionExplicit",
-            "RolesExplicit",
+            "DecisionExplicit")
+freq2.col<-c("RolesExplicit",
             "OpenMeetings",
@@
   select(summary.col)
+```
+
+```{r analyze data, include=FALSE}
+options(knitr.table.format = "html")
+knitr::opts_chunk$set(echo = FALSE,
+                      warning = FALSE,
+                      tidy = FALSE,
+                      message = FALSE,
+                      fig.align = 'center',
+                      out.width = "100%")
 
@@
            "Tradeoffs"="TradeOffsExplicit",
-           "Decision"="DecisionExplicit",
-           "Roles"="RolesExplicit",
-           "Open Meetings"="OpenMeetings",
-           "Adopted"="ResultsAdopted") %>%
+           "Decision"="DecisionExplicit") %>%
     summarise_all(funs(sum))%>%
@@
     mutate(Explicit=factor(Explicit,levels=
-                  c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision",
-                               "Roles","Open Meetings","Adopted"))) %>%
+                  c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision"))) %>%
     rename("Number"="value")
@@
          "Tradeoffs"="TradeOffsExplicit",
-         "Decision"="DecisionExplicit",
-         "Roles"="RolesExplicit",
-         "Open Meetings"="OpenMeetings",
-         "Adopted"="ResultsAdopted") %>%
+         "Decision"="DecisionExplicit") %>%
   summarise_all(funs(sum))%>%
@@
   mutate(Explicit=factor(Explicit,levels=
-                           c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision",
-                             "Roles","Open Meetings","Adopted"))) %>%
+                           c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision"))) %>%
   rename("Number"="value")
@@
 # Random Sample MSEs
-Freq.plot<-ggplot(freq.data,aes(Explicit,Percent,fill=Analysis))+
-  geom_col(position="dodge")+geom_vline(xintercept=3.5,linetype="dashed")+
-  geom_vline(xintercept=1.5,linetype="dashed")+
-  coord_flip()+scale_fill_manual(values=c("#66CCCC","#006666")) +
+Freq.plot<-ggplot(freq.data_pub,aes(Explicit,Percent))+
+  geom_col()+
+  coord_flip() +
   scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
   scale_x_discrete(
-    limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
+    limits=c("Decision","Tradeoffs",
              "Alternatives","Objectives","Problem","Process"),
-    labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
-             "Alternatives","Objectives","Problem","Process")) +
-  guides(fill = guide_legend(reverse=T)) + theme_bw() +
-  theme(legend.position = c(0.85,0.07),legend.title = element_blank())
+    labels=c("Decision","Tradeoffs",
+             "Alternatives","Objectives","Problem","Process")) + theme_bw()
 
-Part.plot<-ggplot(part.data,aes(x=order,y=Percent,fill=Analysis)) +
-  facet_wrap(~Stage,scale="free",ncol=2) + geom_col(position="dodge") +
+Part.plot<-ggplot(part.data_pub,aes(x=order,y=Percent)) +
+  facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +
   scale_fill_manual(values=c("#66CCCC","#006666")) +
@@
   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
-  ylab("Percent")+xlab(NULL)+coord_flip() +
-  guides(fill = guide_legend(reverse=T)) + theme_bw() +
-  theme(legend.position = c(0.86,0.075),legend.title = element_blank())
+  ylab("Percent")+xlab(NULL)+coord_flip() + theme_bw()
 
@@
 
+### Objectives
+We review published MSEs against the steps of the SDM framework to:
+
+* Determine the rate of documentation, and use of steps, in the MSE process:  
+* Identify participants in MSE processes.
+* Highlight unique aspects of MSEs that model climate change as a driver.
+* Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate  
+* Provide insights about MSE processes and documentation that can further learning to improve future MSEs
+
 # Methods
+## Finding and Sampling the MSE Literature
+We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019.
+This search returned 253 results.
+We reviewed a random sample of 30 articles that document a MSE, removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
+Of the 253 articles, 11 included climate change as a driver of fishery status, and after removing other articles we estimate 110-160 articles document a MSE.
+
+| MSE type | Count |
+|:--:|:--:|
+| Published (estimated) | 110-160 |
+| Random Sample | 30 |
+| Climate Change | 11 |
+
+## Reviewing MSE documentation 
+Methods for scoring the documentation.
 
 # Results
+## Are structured decision making steps explicit in MSEs?
+The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Table \@ref(tab:explicit)).
+The most documented stage was the alternative stage, while 3 of the 30 studies explicitly documented a tradeoff analysis and 2 of the 30 documented a management decision.
+
+```{r explicit}
+freq.data_pub %>%
+  mutate("Percent"=round(Percent,0)) %>% 
+  kable(align = "c", caption = "Prevalence of Explicit Documentation of stages in the MSE process")
+```
+
+### Who is involved in MSEs?
+The primary participants and leaders of the MSE processes were scientists, with some lead by governments and management agencies (Figure \@ref(fig:participants)).
+Objectives and alternatives were elicited from a variety of participants, with scientists seemingly selecting these components in cases where documentation was inexplicit.
+
+```{r participants, fig.cap='Who guided, participated in, or provided input during the specified steps of the MSE process', fig.height=6, out.width="100%"}
+Part.plot
+```
+
+<!-- ### Where have climate change MSEs occurred? -->
+<!-- To date MSEs modeling climate change have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)). -->
+
+<!-- ```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=4, out.width="100%"} -->
+<!-- MSE.map + -->
+<!--   theme(text = element_text(size=14)) -->
+<!-- ``` -->
+
+<!-- ### What questions and management objectives do MSEs address? -->
+<!-- Uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the random sample of MSEs (Figure \@ref(fig:driversPlot)). -->
+
+<!-- ```{r driversPlot, fig.cap='Drivers included in MSE models', fig.height=6, out.width="100%"} -->
+<!-- Driver.plot + -->
+<!--   theme(text = element_text(size=14)) -->
+<!-- ``` -->
+
+<!-- More objectives were considered per MSE in the random sample than in the climate change MSEs (Table \@ref(tab:PerMSE)).
+Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:objectiveCategories)).
+-->
+
+<!-- ```{r, PerMSE} -->
+<!-- color_Pub <- which(per.MSE$Analysis=='Random Sample') -->
+<!-- color_CC <- which(per.MSE$Analysis=='Climate Change') -->
+
+<!-- per.MSE %>% -->
+<!--   kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%   -->
+<!--   row_spec(color_CC, color = "#66CCCC", background = "white") %>% -->
+<!--   row_spec(color_Pub, color = "#006666", background = "white") -->
+<!-- ``` -->
+
+<!-- ```{r, objectiveCategories} -->
+<!-- color_Pub <- which(objcat.data$Analysis=='Random Sample') -->
+<!-- color_CC <- which(objcat.data$Analysis=='Climate Change') -->
+
+<!-- objcat.data %>% -->
+<!--   kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>% -->
+<!--   row_spec(color_CC, color = "#66CCCC", background = "white") %>% -->
+<!--   row_spec(color_Pub, color = "#006666", background = "white") -->
+<!-- ``` -->
+
+### Main Conclusions
+The picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes. 
+While documentation about the **modeling** conducted in a MSE was explicit, documentation about the **process used to make decisions** about the MSE model **were limited**.
+Scientists participated in, or seemingly directly selected the objectives and alternatives to evaluate in most MSEs.
+Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.
+Conservation, Economic, and Yield focused objectives were commonly considered, while **broader social objectives were rarely considered**.
+
+Despite numerous applications of MSE, the applications that include climate change are limited, leaving uncertainty about the best approach to MSE use for fishereis management that accounts for a changing environment.
+
+### Explore our review on the web!
+Check out our shiny app (<https://jonathancummings.shinyapps.io/MSEreview/>).
+
+### References
 

### 2019-10-03 Figures.R
Load data from file; edit participant order; turn off database SQL examples

@@ -40,6 +40,8 @@ getSqlConnection <- function(){
 # Open connection
 con <- getSqlConnection()
 
+# load("C:/Users/jcummings/OneDrive - UMASS Dartmouth/SMAST/Research/MSE Problem Framing/MSEreview/MSEreview.RData")
+
 # Obtain data tables
 study<-dbReadTable(con,"tblStudy")
 # study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
@@ -392,8 +394,26 @@ part.data_CC<-mutate(part.data_CC,Analysis="Climate Change")
 part.data<-rbind(part.data_pub,part.data_CC) %>% 
   arrange(Stage, Percent) %>% 
   mutate(order = row_number())
-part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,16,17,18,20,21,22,23,24,25,21,23,
-                   26,25,26,27,27,28,29,30,31,29,32,30,33,32,33,34,35,36,37,38,39,39)
+part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,
+                   16,17,18,20,21,22,23,24,25,21,23,26,25,26,27,27,28,29,30,31,29,32,30,33,
+                   32,33,34,35,36,37,38,39,39)
+
+# part.data<-rbind(part.data,c('Process','Government',1,1,1,'Climate Change'),
+#                  c('Process','Management',1,1,2,'Climate Change'),
+#                  c('Participants','Analysts',1,1,5,'Climate Change'),
+#                  c('Participants','Decision Makers',1,1,11,'Climate Change'),
+#                  c('Participants','Public',1,1,7,'Climate Change'),
+#                  c('Explicit Objectives Process','Analysts',1,1,15,'Climate Change'),
+#                  c('Explicit Objectives Process','Independent',1,1,19,'Publication'),
+#                  c('Subjective Objectives Process','Experts',1,1,24,'Publication'),
+#                  c('Subjective Objectives Process','Independent',1,1,22,'Climate Change'),
+#                  c('Explicit Alternatives Process','Analysts',1,1,28,'Climate Change'),
[…]

### 2019-10-03 MSE manuscript.Rmd
Edit results section

@@
 world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
+```
+text
 
+```{r plots}
 # Where MSEs have occured
@@
 
-Driver.plot<-ggplot(drive.data,aes(x=order,y=Percent,fill=Analysis)) +
-  geom_col(position="dodge") +
-  scale_fill_manual(values=c("#66CCCC","#006666")) +
-  scale_x_continuous(breaks = drive.data$order,
-                     labels = drive.data$Driver)+
+Driver.plot<-ggplot(drive.data_pub,aes(x=order,y=Percent)) +
+  geom_col() +
+  scale_x_continuous(breaks = drive.data_pub$order,
+                     labels = drive.data_pub$Driver)+
   scale_y_continuous(limits=c(0,100),expand = c(0,0))+
-  ylab("Percent")+xlab(NULL)+coord_flip() +
-  guides(fill = guide_legend(reverse=T)) + theme_bw() +
-  theme(legend.position = c(0.825,0.075),legend.title = element_blank())
+  ylab("Percent")+xlab(NULL)+coord_flip()+ theme_bw()
 ```
@@
 The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Table \@ref(tab:explicit)).
-The most documented stage was the alternative stage, while 3 of the 30 studies explicitly documented a tradeoff analysis and 2 of the 30 documented a management decision.
+The most documented stage was the alternative stage, while 3 of the 30 MSEs explicitly documented a tradeoff analysis and 2 of the 30 documented a management decision.
+One of the 30 documented either the roles of participants in the MSE process or that the results of the MSE were adopted.
 
@@
 
+### What questions and management objectives do MSEs address?
+Uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the random sample of MSEs (Figure \@ref(fig:driversPlot)).
+
+```{r driversPlot}
+drive.data_pub %>%
+  mutate("Percent"=round(Percent,0)) %>% 
+  select(Driver,Percent,Frequency) %>% 
+  kable(align = "c", caption = "Prevalence of Explicit Documentation of stages in the MSE process")
+```
+
+More objectives were considered per MSE in the random sample than in the climate change MSEs (Table \@ref(tab:PerMSE)).
+Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:objectiveCategories)).
+
+```{r, PerMSE}
+color_Pub <- which(per.MSE$Analysis=='Random Sample')
+color_CC <- which(per.MSE$Analysis=='Climate Change')
+
+per.MSE %>%
+  kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%
+  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
+  row_spec(color_Pub, color = "#006666", background = "white")
+```
+
+```{r, objectiveCategories}
+color_Pub <- which(objcat.data$Analysis=='Random Sample')
+color_CC <- which(objcat.data$Analysis=='Climate Change')
+
+objcat.data %>%
+  kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>%
+  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
+  row_spec(color_Pub, color = "#006666", background = "white")
+```
+
 <!-- ### Where have climate change MSEs occurred? -->
@@
 <!--   theme(text = element_text(size=14)) -->
-<!-- ``` -->
-
-<!-- ### What questions and management objectives do MSEs address? -->
-<!-- Uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the random sample of MSEs (Figure \@ref(fig:driversPlot)). -->
-
-<!-- ```{r driversPlot, fig.cap='Drivers included in MSE models', fig.height=6, out.width="100%"} -->
-<!-- Driver.plot + -->
-<!--   theme(text = element_text(size=14)) -->
-<!-- ``` -->
-
-<!-- More objectives were considered per MSE in the random sample than in the climate change MSEs (Table \@ref(tab:PerMSE)).
-Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:objectiveCategories)).
--->
-
-<!-- ```{r, PerMSE} -->
-<!-- color_Pub <- which(per.MSE$Analysis=='Random Sample') -->
-<!-- color_CC <- which(per.MSE$Analysis=='Climate Change') -->
-
-<!-- per.MSE %>% -->
-<!--   kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%   -->
-<!--   row_spec(color_CC, color = "#66CCCC", background = "white") %>% -->
-<!--   row_spec(color_Pub, color = "#006666", background = "white") -->
-<!-- ``` -->
-
-<!-- ```{r, objectiveCategories} -->
-<!-- color_Pub <- which(objcat.data$Analysis=='Random Sample') -->
-<!-- color_CC <- which(objcat.data$Analysis=='Climate Change') -->
-
-<!-- objcat.data %>% -->
-<!--   kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>% -->
-<!--   row_spec(color_CC, color = "#66CCCC", background = "white") %>% -->
-<!--   row_spec(color_Pub, color = "#006666", background = "white") -->
 <!-- ``` -->
@@
 
-# Introduction
-
-**Lorem ipsum** dolor *sit amet*.
-
-- Single asterisks italicize text *like this*. 
-- Double asterisks embolden text **like this**.
-
-Start a new paragraph with a blank line separating paragraphs.
-
-- This will start an unordered list environment, and this will be the first item.
-- This will be a second item.
-- A third item.
-    - Four spaces and a dash create a sublist and this item in it.
-- The fourth item.
-
-1. This starts a numerical list.
-2. This is no. 2 in the numerical list.
-
-# This Starts A New Section
-## This is a Subsection
-### This is a Subsubsection
-#### This starts a Paragraph Block.
-
-> This will create a block quote, if you want one.
-Want a table? This will create one.
-
-Table Header  | Second Header
-------------- | -------------
-Table Cell    | Cell 2
-Cell 3        | Cell 4
-
-Note that the separators *do not* have to be aligned.
-
 Want an image? This will do it.
@@
 `fig_caption: yes` will provide a caption. Put that in the YAML metadata.
-
-Almost forgot about creating a footnote.[^1] This will do it again.[^2]
-
-[^1]: The first footnote
-[^2]: The second footnote
-
-Want to cite something?
 

### 2019-10-03 app.R
Add about tab; Edit column width for all column data table & problem data table; Edit plot formats; round and edit number of digits displayed in tables; fix stringsplits

@@ -119,7 +119,11 @@ data<-select(data,-c("ID"))
 obj.data<-select(obj.data,-c("ID","ID.y"))
 
 # Get columns whose width needs editing
-targets<-match(c("FullCitation","Comments"),names(data))
+targetsC<-match(c("Comments","ProblemDefinition"),names(data))
+targetsAE<-match(c("AlternativesEvaluated","FullCitation"),names(data))
+targetsSp<-match(c("Species","ObjElicitationMethod"),names(data))
+targetsSy<-match(c("System"),names(data))
+targetsL<-match(c("Location"),names(data))
 
 # Get columns for study summary
 summary.col<-c("Citation",
@@ -134,6 +138,7 @@ prob.col<-c("Citation",
             "TradeOffMethod_Exp",
             "TradeOffMethod_Sub",
             "Decision")
+targetsPD<-match(c("ProblemDefinition"),prob.col)
 # Get columns for frequency analysis
 freq.col<-c("ProcessExplicit",
             "ProblemDefinitionExplicit",
@@ -170,90 +175,93 @@ ui <- fluidPage(
   tabsetPanel(
     #### Tab 1: About ####
     tabPanel("About", 
+      h1("An Assessment of Management Strategy Evaluations"),# title of the page
+      hr(),
+      h4("Background"),
+      p("Fisheries management has mostly focused on fishing impacts
+      with ecosystem status viewed as a background constant."),
+      p("The increasing rate of climate change is changing this
+      dynamic, bringing ecosystem status to the forefront of fisheries
+      management."),
+      p("Management strategy evaluation (MSE) is 'widely considered to
+      be the most appropriate way to evaluate the trade‐offs achieved by
+      alternative management strategies and to assess the consequences of
+      uncertainty for achieving management goals' (Punt et al. 2014). Thus,
+      MSE is a compelling tool to assess climate change impacts and test
+      climate-ready options for fisheries management decisions Adaptive
+      management arose to address uncertainties and accelerate progress
+      towards meeting management objectives. We used the structured decision
+      making (SDM) process - the decision making framework in which adaptive
+      management occurs - as our framework, how do published MSE projects
+      utilize standard SDM components and support learning within the MSE
+      practitioner community?"),
+      imageOutput("imageSDM",
+                   width=400,
+                   height=275),
+      p(em("Structured Decision Making Process")),
+      h4("MSE Review Methods"),
+      p("We conducted our search for MSEs in the SCI-EXPANDED index from
[…]

### 2019-10-23 MSE manuscript.Rmd
Edits to create draft MS

@@
 title: 'No Replication Without Process Documentation:  An Assessment of Management
-  Strategy '
+  Strategy Evaluations'
 author:
-- affiliation: School for Marine Science and Technology, University of Massachusetts - Dartmouth
-  name: Jonathan W. Cummings
-date: "`r format(Sys.time(), '%B %d, %Y')`"
+- affiliation: School for Marine Science and Technology, University of Massachusetts
+    - Dartmouth
+  name: Jonathan W. Cummings, Amanda Hart, Gavin Fay
 output:
-  html_document:
-    df_print: paged
-  pdf_document:
-    citation_package: natbib
-    fig_caption: yes
-    keep_tex: yes
-    latex_engine: pdflatex
-    template: svm-latex-ms.tex
+  bookdown::html_document2:
+    number_sections: false
+  bookdown::pdf_document2:
+    number_sections: false
 biblio-style: apsr
@@
 bibliography: MSEreview.bib
-abstract: Enter abstract text here
+abstract: The aim and vision for fisheries management that comes from management strategy evaluation and adaptive management is a nobel and aspirational vision.
+We review MSE documentation in the published literature to appraise the achievement of this vision, finding that reality doesn't fully reflect this vision.
+The steps taken to complete the MSE process often are not documented.
+Scientists often fulled the role of fishery stakeholders, seemingly selecting the objectives and alternatives to evaluate in most MSEs.
+Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.
+Conservation, Economic, and Yield focused objectives were commonly considered in MSE processes, while broader social objectives were rarely considered.
+The picture painted by MSE documentation is primarily one of scientist driven simulation studies rather than management driven decision processes, with limited process documentation.
 ---
@@
 
-load("C:/Users/jcummings/OneDrive - UMASS Dartmouth/SMAST/Research/MSE Problem Framing/MSEreview/MSEreview.RData")
+load("MSEreview.RData")
 
@@
             "TradeOffsExplicit",
-            "DecisionExplicit")
+            "DecisionExplicit",
+            "ResultsAdopted")
 freq2.col<-c("RolesExplicit",
-            "OpenMeetings",
-            "ResultsAdopted")
+             "OpenMeetings")
 # Get columns for participant analysis
@@
            "Tradeoffs"="TradeOffsExplicit",
-           "Decision"="DecisionExplicit") %>%
+           "Decision"="DecisionExplicit",
+           "Results Adopted"="ResultsAdopted") %>%
     summarise_all(funs(sum))%>%
     gather(Explicit) %>%
-    mutate(Percent=value/n_mse*100) %>%
+    mutate(Percent=value/n_pub*100) %>%
     mutate(Explicit=factor(Explicit,levels=
-                  c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision"))) %>%
+                  c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision","Results Adopted"))) %>%
     rename("Number"="value")
@@
          "Tradeoffs"="TradeOffsExplicit",
-         "Decision"="DecisionExplicit") %>%
+         "Decision"="DecisionExplicit",
+         "Results Adopted"="ResultsAdopted") %>%
   summarise_all(funs(sum))%>%
   gather(Explicit) %>%
-  mutate(Percent=value/n_mse*100) %>%
+  mutate(Percent=value/n_CC*100) %>%
   mutate(Explicit=factor(Explicit,levels=
-                           c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision"))) %>%
+                           c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision","Results Adopted"))) %>%
+  rename("Number"="value")
+
+freq2.data_pub<-data_pub %>%
+    select(freq2.col) %>%
+    rename("Roles"="RolesExplicit",
+           "Open Meetings"="OpenMeetings") %>%
+    summarise_all(funs(sum)) %>%
+    gather(Explicit) %>%
+    mutate(Percent=value/n_pub*100) %>%
+    mutate(Explicit=factor(Explicit,levels=
+                  c("Roles","Open Meetings"))) %>%
     rename("Number"="value")
@@
 ```
-text
 
@@
   scale_fill_manual(values=c("#66CCCC","#006666")) +
-  scale_x_continuous(breaks = part.data$order,
-                     labels = part.data$Participants)+
+  scale_x_continuous(breaks = part.data_pub$order,
+                     labels = part.data_pub$Participants)+
   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
@@
 # Introduction 
-Fisheries management has mostly focused on fishing impacts with ecosystem status viewed as a background constant.
+The calls for fisheries management to utilize management strategy evaluation and adaptive management have been both aspirational and estimable in their aim and vision.
+For example, management strategy evaluation (MSE) is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@Punt].
+The positive aspects of MSE that have been listed include that it is “A tool that scientists and managers can use to simulate the workings of a fisheries system and allow them to test whether potential harvest strategies—or management procedures— can achieve pre-agreed management objectives.”[@Pew], that it “can be used to identify a ‘best’ management strategy among a set of candidate strategies" [@PuntLab] and “MSE has the advantage of being able to reveal the trade-offs among a range of possible management decisions, and should assist our Commissioners in determining whether the objectives should be weighted differently.
+Specifically, to provide the information on which to base a rational decision, given objectives, preferences, and attitudes to risk.”  – International Pacific Halibut Commission  Based on these descriptions MSE is a compelling tool to support and improve fisheries management decisions.
 
-The increasing rate of climate change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
+In terms of the MSE process, it is described as being "at the interface between science and policy.” [@PuntLab] and being an “Undertaking [that] requires scientists, managers, and stakeholders to be involved throughout the process." [@IPHC] The International Pacific Halibut Commission (2017) also states, "While the scientists do the modelling, managers must offer extensive input.
+Because of the many steps and the iterative process, communication among parties is critical for achieving buy-in on the results of the management strategy evaluation.”
 
-Management strategy evaluation (MSE) is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@R-Punt].
-Thus, MSE is a compelling tool to assess climate change impacts and test climate-ready options for fisheries management decisions.
-
-Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives.
-We used the structured decision making (SDM) process [@R-SDM] -- the decision making framework in which adaptive management occurs -- as our framework, how do published MSE projects utilize standard SDM components and support learning within the MSE practitioner community?
+Adaptive management developed in fisheries management as a means to improve the achievement of objectives through time, when learning can aid achieving those objectives.
+For example, the Millennium Ecosystem Assessment (2005) defines adaptive management as “A systematic process for continually improving management policies and practices by learning from the outcomes of previously employed policies and practices.
+In active adaptive management, management is treated as a deliberate experiment for purposes of learning.” We contend that scientific publication is essentially a process intended to support adaptive management, in that researchers gain knowledge from the decisions made and outcomes produced by previous work to inform their decisions.
 
 ### Objectives 
-We review published MSEs against the steps of the SDM framework to:
+Our overall objectives is to understanding of how MSEs utilize standard components of decision making processes and to what extent their documentation can support learning within the MSE practitioner community?
+Given that scientific publication is the primary means of communicating the results of MSE, we review the scientific literature to assess the extent to which the process and aims of MSE, and adaptive management are being achieved and supported by the scientific literature.
+We use the structured decision making (SDM) process [@SDM] -- the decision making framework in which adaptive management occurs -- as our framework for evaluation.
+Using this framework we review published MSEs to:
 
-* Determine the rate of documentation, and use of steps, in the MSE process:  
+* Determine what questions MSE processes are addressing
+* Identify the objectives and alternatives considered in MSE
+* Determine the prevalence of explicit documentation and use of decision making steps in the MSE process, and 
 * Identify participants in MSE processes.
-* Highlight unique aspects of MSEs that model climate change as a driver.
-* Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate  
-* Provide insights about MSE processes and documentation that can further learning to improve future MSEs
+
+We also conduct a focused analysis of MSE processes that address the influence of climate change on fisheries management, highlighting unique aspects of MSE processes that model climate change as a driver of system status.
 
@@
 |:--:|:--:|
-| Published (estimated) | 110-160 |
+| Published (estimated*) | ~142 |
 | Random Sample | 30 |
@@
 ## Reviewing MSE documentation 
-Methods for scoring the documentation.
+We used the structured decision making framework as our framework for evaluating management strategy evaluation documents.
+[insert sdm figure].
+For ease of presentation we present the review methodology and results in four groups: consequences, objectives, stages, and process.
+
+In the consequence group we report the drivers included as the primary factors influencing the predictions produced by the MSE model and report how the consequence predictions were made.
+The information reported here addresses what question the MSE was designed to address, and how it was addressed.
+
+In the objectives group we report the category of objectives considered in the MSE, and what type of objectives were considered.
+We noted four types of objectives: strategic objectives regarding the overall mission of the system in which the decision evaluated by the MSE falls, process objectives addressing how the management decision addressed by the MSE would be made, fundamental objectives addressing the goals of the management decision, and means objectives addressing means to achieving fundamental objectives of the management decision.
+
+In the stages group we report whether the stages (Problem, Objectives, Alternatives, Consequences, Trade-offs, and Decision) of the structured decision making framework were explicitly completed and documented.
+We also report whether the MSE results were adopted in subsequent management.
+In many cases the process used to elicit objectives and alternatives was not explicit, but some information is provided or can be gleaned about who participated in the production of those components of the MSE.
+In those cases we record and report who those participants were, referring to them as subjective participants in the objectives or alternatives process.
+
+In the process group we report who lead the MSE, who participated in it, if the roles of participants were established and documented.
+
+We collected additional information about the MSEs we reviewed.
+For a full description of the information recorded in database columns for this MSE review see the supplemental information (supplement), and for the full results see our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
 
 # Results 
+### What questions and management objectives do MSEs address? 
+We found that uncertainty was the main driver of the system state included in management strategy evaluations, with uncertainty included in 21 of the 30 evaluations we reviewed (Table \@ref(tab:drivers)).
+In addition to uncertainty about the system model itself, monitoring methodology was the most prevalent driver in MSE models, but a variety of drivers were considered across the random sample of MSEs.
+
+```{r drivers}
+drive.data_pub %>%
+  mutate("Percent"=round(Percent,0)) %>% 
+  select(Driver,Percent,Frequency) %>% 
+  kable(caption = "Drivers included in management strategy evaluations") %>% 
+  kable_styling()
+```
+<br>
+Conservation objectives were present in nearly all of the MSEs we reviewed (Table \@ref(tab:objectiveCategories)).
+Yield and economic objectives were present in about three quarters of the MSEs, while social objectives were only included in 2 of the MSEs.
+
+```{r, objectiveCategories}
+objcat.data_pub %>% 
+  select(-Analysis) %>% 
+  kable(caption = "Consideration of objective categories in management strategy evaluation") %>% 
+  kable_styling()
+```
+<br>
+
 ## Are structured decision making steps explicit in MSEs? 
 The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Table \@ref(tab:explicit)).
-The most documented stage was the alternative stage, while 3 of the 30 MSEs explicitly documented a tradeoff analysis and 2 of the 30 documented a management decision.
+The most documented stage was the alternative stage with 6 MSEs documenting this stage of the process.
+Three of the 30 MSEs explicitly documented a trade-off analysis and 2 of the 30 documented a management decision.
 One of the 30 documented either the roles of participants in the MSE process or that the results of the MSE were adopted.
@@
 freq.data_pub %>%
+  select(-Analysis) %>% 
   mutate("Percent"=round(Percent,0)) %>% 
-  kable(align = "c", caption = "Prevalence of Explicit Documentation of stages in the MSE process")
+  kable(caption = "Prevalence of explicit documentation of stages in the MSE process") %>% 
+  kable_styling()
 ```
@@
 ### Who is involved in MSEs? 
+MSEs were rarely explicit about the roles participants played in MSE processes, as only one MSE documented participant roles (Table \@ref(tab:roles)).
+None of the MSEs noted whether or not public meetings were held in which the public could participate in the MSE process.
+
+```{r roles}
+freq2.data_pub %>% 
+  mutate("Percent"=round(Percent,0)) %>% 
+  kable(caption = "Prevalence of explicit roles for MSE participants and meetings that were open to the public during the MSE process") %>% 
+  kable_styling()
+```
+
 The primary participants and leaders of the MSE processes were scientists, with some lead by governments and management agencies (Figure \@ref(fig:participants)).
-Objectives and alternatives were elicited from a variety of participants, with scientists seemingly selecting these components in cases where documentation was inexplicit.
+Scientist lead 73 percent of the MSE processes, with another 23 percent lacking documentation that made the leader of the process clear.
+
+Scientists were participants in nearly all of the reviewed MSE processes.
+Other stakeholders in MSE processes were less frequent participants.
+Members of the fishery participated in 30 percent, management and government representatives in 23 percent, representatives of independent institutions in 7 percent, and the members of the public in 3 percent of reviewed MSEs.
+Decision makers, those responsible for selecting the management plan to implement, were participants in 13 percent of MSEs.
+Experts, a possible source of data or predictions, participated in 3 percent of the MSEs reviewed.
+Participants involved to the assist the MSE process itself, facilitators and analysts participated in 10 and 3 percent of MSE processes respectively.
+
+Objectives and alternatives were elicited from a variety of participants.
+About 10 percent of the objectives and alternatives explicitly were produced by stakeholders in the fisheries, coming from management, government, or the fishery.
+In most cases the source of the objectives and alternatives was not explicit.
+In those cases scientists selected these components most frequently.
 
@@
 
-### What questions and management objectives do MSEs address?
-Uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the random sample of MSEs (Figure \@ref(fig:driversPlot)).
+# Climate Change in MSE (As *Box Insert*) 
+Fisheries management has mostly focused on fishing impacts with ecosystem status viewed as a background constant.
+The increasing rate of climate change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
 
-```{r driversPlot}
-drive.data_pub %>%
+We review published MSEs including climate change as a driver against the steps of the SDM framework to:  Highlight unique aspects of MSEs that model climate change as a driver to inform fisheries management in a changing climate.
+
+We found that as of the end of December 2018 11 MSEs included climate change as a driver. These MSEs have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).
+
+```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=4, out.width="100%"}
+MSE.map +
+  theme(text = element_text(size=14))
+```
+
+Uncertainty and monitoring methodology were also included in some of the climate change MSEs (Table \@ref(tab:driversCC)), but not as often.
+Other biological drivers were also included in some of the climate change MSEs.
+Overall climate change MSEs considered fewer drivers on average, 1.64, than the random sample of MSEs, which included 1.87.
+
+```{r driversCC}
+drive.data_CC %>%
   mutate("Percent"=round(Percent,0)) %>% 
   select(Driver,Percent,Frequency) %>% 
-  kable(align = "c", caption = "Prevalence of Explicit Documentation of stages in the MSE process")
+  kable(caption = "Drivers included in climate change management strategy evaluations") %>% 
+  kable_styling()
 ```
 
-More objectives were considered per MSE in the random sample than in the climate change MSEs (Table \@ref(tab:PerMSE)).
-Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:objectiveCategories)).
-
-```{r, PerMSE}
+More objectives were considered per MSE in the random sample of MSEs than in the climate change MSEs (Table \@ref(tab:CCPerMSE)).
+Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:CCobjectiveCategories)).
+```{r, CCPerMSE}
 color_Pub <- which(per.MSE$Analysis=='Random Sample')
@@
 per.MSE %>%
-  kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%
+  kable(caption = "Number of objectives considered per management strategy evaluation") %>%
   row_spec(color_CC, color = "#66CCCC", background = "white") %>%
-  row_spec(color_Pub, color = "#006666", background = "white")
+  row_spec(color_Pub, color = "#006666", background = "white")%>% 
+  kable_styling()
 ```
 
-```{r, objectiveCategories}
+```{r, CCobjectiveCategories}
 color_Pub <- which(objcat.data$Analysis=='Random Sample')
@@
 objcat.data %>%
-  kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>%
+  kable(caption = "Objective categories considered in management strategy evaluations") %>%
   row_spec(color_CC, color = "#66CCCC", background = "white") %>%
-  row_spec(color_Pub, color = "#006666", background = "white")
+  row_spec(color_Pub, color = "#006666", background = "white")%>% 
+  kable_styling()
 ```
 
-<!-- ### Where have climate change MSEs occurred? -->
-<!-- To date MSEs modeling climate change have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)). -->
+Climate change MSEs were more likely to explicitly document the alternative elicitation and trade-off analysis steps, but otherwise were less likely to document steps in the decision process.
+```{r, CCstages}
+freq.data_CC %>%
+  select(-Analysis) %>% 
+  mutate("Percent"=round(Percent,0)) %>% 
+  kable(caption = "Prevalence of explicit documentation of stages in climate change MSE processes") %>%
+  kable_styling()
+```
 
-<!-- ```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=4, out.width="100%"} -->
-<!-- MSE.map + -->
-<!--   theme(text = element_text(size=14)) -->
-<!-- ``` -->
+Including climate change in a MSE appears to reduce the complexity of other stages in the MSE process, as on average climate change MSEs consider fewer drivers, fewer and less diverse objectives, and are less explicitly documented at most stages of the process.
+If the MSE process is time or effort limited, the inclusion of climate change may be reducing the time and effort that otherwise is invested in other stages of the MSE process.
 
-### Main Conclusions
-The picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes. 
-While documentation about the **modeling** conducted in a MSE was explicit, documentation about the **process used to make decisions** about the MSE model **were limited**.
-Scientists participated in, or seemingly directly selected the objectives and alternatives to evaluate in most MSEs.
-Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.
-Conservation, Economic, and Yield focused objectives were commonly considered, while **broader social objectives were rarely considered**.
+# Discussion 
+The primary picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes.
+While documentation about the modeling conducted in a MSE was explicit, documentation about the MSE process itself was often limited.
 
-Despite numerous applications of MSE, the applications that include climate change are limited, leaving uncertainty about the best approach to MSE use for fishereis management that accounts for a changing environment.
+Management strategy evaluations typically do assess the consequences of uncertainty for achieving management goals, but other factors influencing the consequences of management strategies get more sporadic treatment.
+[more about including uncertainty].
+Data limitation is a frequent concern in fisheries management.
+Next to accounting for uncertainty, accessing the influence of monitoring methodology was the most prevalent driver included in MSEs.
+Further inclusion of this driver, and additional use of value of information analyses could expand the use of MSE as a tool to design cost-effective fisheries monitoring plans.
+While discussions of fisheries research and management seem to be move toward ecosystem effects, spatial models, and multiple species models, these characteristics were rarely considered in the set of MSEs we reviewed.
 
-### Explore our review on the web!
-Check out our shiny app (<https://jonathancummings.shinyapps.io/MSEreview/>).
+Managing natural resources entails achieving the objectives of people interacting with a natural resource, so decision processes are likely to produce more desirable results when they include peoples objectives.
+These objectives are likely to include conservation, yield, economic, and social concerns, however the MSE we reviewed don't cover all categories of objectives, as yield and economic objectives were not included in about one quarter of MSEs, and social objectives were absent from all but 2 of the 30 evaluations.
+While we are unable to know the cause for this, our results may provide some clues.
+
+Scientists led, and participated in most of the MSEs, while decision-makers, decision making institutions, and stakeholders were relatively infrequent leaders and participants in the process.
+The lack of participation by stakeholders may result in less frequent inclusion of social objectives.
+Another factor may be that MSE models are likely constructed mainly by ecosystem modelers, who are likely to be more comfortable modeling the consequences of conservation and yield objectives than economic or social objectives.
+Given that scientists seem to be providing the objectives in many MSE processes, they may be selecting objectives that they are more comfortable evaluating.
+There also appears to be some degree of borrowing occurring in which scientists use a set of generic fishery objectives when conducting a MSE.
+Perhaps because their were few social objectives included in these MSE, there was little inclusion of experts to aid in the prediction of consequences, or consequence prediction techniques other than simulation modeling.
+
+Participants that would be likely to expand the set of objectives considered in a MSE rarely participated in the MSEs we reviewed.
+That is, facilitators and decision analysts were rarely participants in MSE processes.
+The inclusion of facilitators in decision processes aid information gain and conflict resolution, increase trust between participants, and lead to more beneficial environmental and social outcomes[@Vente].
+Facilitators and decision analysts also aid decision processes by helping to diagnose the problem, which can result in selecting more appropriate analysis techniques for the problem type.
+Finally, the additional structure and focus on the decision process would likely result in more explicit documentation when facilitators and decision analysts participate in MSE, as resulted in the MSE reviewed here (cite Williams et al 2011, Smith et al 2013, and Kolody et al 2008).
+
+The relative dearth of explicit documentation of the MSE processes may hinder learning by MSE practitioners.
+While the modeling was clearly documented, the MSE process steps were often inexplicit.
+Without explicit documentation it is unclear why the analysis was structured as it was, why a particular set of drivers was selected, and why a set of objectives or alternatives were chosen.
+Perhaps the most crucial stage to enable understanding a replication of a decision resulting from a MSE is a clear presentation of how trade-offs between the multiple objectives were evaluated, and what the optimal or chosen alternative was that resulted from the MSE process.
+Different decision makers can rationally look at the same prediction of consequences and select a different preferred alternative if they treat the importance of the objectives differently.
+Therefore, MSE documentation often is not supporting replication or learning due to the frequent absence of trade-off and decision documentation.
+
+### Explore and expand upon this review
+To aid future learning from the MSEs that have been conducted to date we produced a shiny application [@R-Shiny] that enables interactive results to accompany this article and entry of additional MSE information (<https://jonathancummings.shinyapps.io/MSEreview/>).
+We hope that as MSE are conducted in the future they are entered here so that MSE practitioners can learn from each other.
 
@@
 
-# Discussion
-
-Academic workflow, certainly in political science, is at a crossroads.
-The *American Journal of Political Science* (*AJPS*) announced a (my words) ["show your work" initiative](http://ajps.org/2015/03/26/the-ajps-replication-policy-innovations-and-revisions/) in which authors who are tentatively accepted for publication at the journal must hand over the raw code and data that produced the results shown in the manuscript.
-The editorial team at *AJPS* then reproduces the code from the manuscript.
-Pending successful replication, the manuscript moves toward publication.
-The *AJPS* might be at the fore of this movement, and it could be the most aggressive among political science journals, but other journals in our field have signed the joint [Data Access & Research Transparency](http://www.dartstatement.org/) (DART) initiative.
-This, at a bare minimum, requires uploading code from quantitatively-oriented published articles to in-house directories hosted by the journal or to services like [Dataverse](http://dataverse.org/).
-
-There are workflow implications to the Lacour controversy as well.
-Political science, for the foreseeable future, will struggle with the extent of [the data fraud perpetrated by Michael Lacour](http://stanford.edu/~dbroock/broockman_kalla_aronow_lg_irregularities.pdf) in an article co-authored with Donald P.
-Green in *Science*, the general scientific journal of record in the United States.
-A failure to reproduce LaCour's results with different samples uncovered a comprehensive effort by LaCour to "fake" data that provided results to what we felt or believed to be true [(i.e.
-"truthiness")](http://chronicle.com/article/LAffaire-LaCour/230905/).
-However, [fake data can have real consequences](http://kieranhealy.org/blog/archives/2015/05/20/fake-science-real-consequences/) for both the researcher and those who want to learn from it and use it for various purposes.
-Even research done honestly may suffer the same fate if researchers are not diligent in their workflow.
-
-These recent events underscore the DART push and cast a shadow over our workflow.
-However, good workflow has always been an issue in our discipline.
-Cloud storage services like [Dropbox](http://www.dropbox.com) are still relatively new among political scientists.
-Without cloud storage, previous workflow left open the possibility that work between a home computer and an office computer was lost as a function of a corrupted thumb drive, an overheated power supply, or, among other things, the wave of viruses that [would particularly affect Microsoft users every summer](http://money.cnn.com/2003/11/05/technology/microsoftbounty/).
-Social sciences, [unlike engineering](http://kieranhealy.org/blog/archives/2014/01/23/plain-text/), have traditionally relied on software like Microsoft Word for manuscript preparation though any word processor reduces workflow to a series of clicks and strokes on a keyboard.
-This is [a terrible way to track changes](http://www.nytimes.com/2013/04/19/opinion/krugman-the-excel-depression.html) or maintain version control.
-The addition of collaborators only compounds all the aforementioned issues.
-The proverbial left hand may not know what the right hand is doing.
-
-I think there is reason for optimism.
-We only struggle with it now because we have tools like [R Markdown](http://rmarkdown.rstudio.com/) and [Pandoc](http://pandoc.org/), more generally, that make significant strides in workflow.
-LaTeX resolved earlier issues of corrupted binary files by reducing documents to raw markup that was little more than raw text and revisions that could be easily kept as ["commented" text](http://tex.stackexchange.com/questions/11177/how-to-write-hidden-notes-in-a-latex-file).
-However, for all its benefits (including pretty PDFs), [LaTeX is *ugly* code](http://www-rohan.sdsu.edu/~aty/bibliog/latex/gripe.html) and does not provide means of seamlessly working with the actual data analysis itself.
-R Markdown both eliminates markup and allows the author and her collaborators to write and reproduce the manuscript in one fell swoop.
-
-# Getting Started with YAML
-
-The lion's share of a R Markdown document will be raw text, though the front matter may be the most important part of the document.
-R Markdown uses [YAML](http://www.yaml.org/) for its metadata and the fields differ from [what an author would use for a Beamer presentation](http://svmiller.com/blog/2015/02/moving-from-beamer-to-r-markdown/).
-I provide a sample YAML metadata largely taken from this exact document and explain it below.
-
-```{r eval=FALSE}
----
-output: 
-  pdf_document:
-    citation_package: natbib
-    keep_tex: true
-    fig_caption: true
-    latex_engine: pdflatex
-    template: ~/Dropbox/miscelanea/svm-r-markdown-templates/svm-latex-ms.tex
-title: "A Pandoc Markdown Article Starter and Template"
-thanks: "Replication files are available on the author's Github account..."
-author:
-- name: Steven V. Miller
-  affiliation: Clemson University
-- name: Mary Margaret Albright
-  affiliation: Pendelton State University
-- name: Rembrandt Q. Einstein
-  affiliation: Springfield University
-abstract: "This document provides an introduction to R Markdown, argues for its..."
-keywords: "pandoc, r markdown, knitr"
-date: "`r format(Sys.time(), '%B %d, %Y')`"
-geometry: margin=1in
-fontfamily: mathpazo
-fontsize: 11pt
-# spacing: double
-bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`"
-biblio-style: apsr
----
-```
-
-`output:` will tell R Markdown we want a PDF document rendered with LaTeX.
-Since we are adding a fair bit of custom options to this call, we specify `pdf_document:` on the next line (with, importantly, a two-space indent).
-We specify additional output-level options underneath it, each are indented with four spaces.
-`citation_package: natbib` tells R Markdown to use `natbib` to handle bibliographic citations.[^natbib] Thereafter, the next line (`keep_tex: true`) tells R Markdown to render a raw `.tex` file along with the PDF document.
-This is useful for both debugging and the publication stage, when the editorial team will ask for the raw `.tex` so that they could render it and later provide page proofs.
-The next line `fig_caption: true` tells R Markdown to make sure that whatever images are included in the document are treated as figures in which our caption in brackets in a Markdown call is treated as the caption in the figure.
-The next line (`latex_engine: pdflatex`) tells R Markdown to use pdflatex and not some other option like `lualatex`.
-For my template, I'm pretty sure this is mandatory.[^pdflatex]
-
-[^natbib]: R Markdown can use Pandoc's native bibliography management system or even `biblatex`, but I've found that it chokes with some of the more advanced stuff I've done with my .bib file over the years.
-For example, I've been diligent about special characters (e.g.
-umlauts and acute accents) in author names in my .bib file, but Pandoc's native citation system will choke on these characters in a .bib file.
-I effectively need `natbib` for my own projects.
-[^pdflatex]: The main reason I still use `pdflatex` (and most readers probably do as well) is because of LaTeX fonts.
-[Unlike others](http://www-rohan.sdsu.edu/~aty/bibliog/latex/gripe.html), I find standard LaTeX fonts to be appealing.
-
-The next line (`template: ...`) tells R Markdown to use my custom LaTeX template.[^path] While I will own any errors in the code, I confess to "Frankensteining" this template from [the default LaTeX template](https://github.com/jgm/pandoc-templates) from Pandoc, [Kieran Healy's LaTeX template](https://github.com/kjhealy/pandoc-templates/tree/master/templates), and liberally using raw TeX from the [Association for Computing Machinery's (ACM) LaTeX template](https://www.acm.org/publications/article-templates/acm-latex-style-guide).
-I rather like that template since it resembles standard manuscripts when they are published in some of our more prominent journals.
-I will continue with a description of the YAML metadata in the next paragraph, though invite the curious reader to scroll to the end of the accompanying post to see the PDF this template produces.
-
-[^path]: Notice that the path is relative.
-The user can, if she wishes, install this in the default Pandoc directory.
-I don't think this is necessary.
-Just be mindful of wherever the template is placed.
-Importantly, `~` is used in R to find the home directory (not necessarily the working directory).
-It is equivalent to saying `/home/steve` in Linux, or `/Users/steve` on a Mac, in my case.
-
-The next fields get to the heart of the document itself.
-`title:` is, intuitively, the title of the manuscript.
-Do note that fields like `title:` do not have to be in quotation marks, but must be in quotation marks if the title of the document includes a colon.
-That said, the only reason to use a colon in an article title is if it is followed by a subtitle, hence the optional field (`subtitle:`).
-Notice I "comment out" the subtitle in the above example with a pound sign since this particular document does not have a subtitle.
-If `thanks:` is included and has an accompanying entry, the ensuing title of the document gets an asterisk and a footnote.
-This field is typically used to advise readers that the document is a working paper or is forthcoming in a journal.
-
-The next field (`author:`) is a divergence from standard YAML, but I think it is useful.
-I will also confess to pilfering this idea from Kieran Healy's template.
-Typically, multiple authors for a given document are separated by an `\and` in this field.
-However, standard LaTeX then creates a tabular field separating multiple authors that is somewhat restrictive and not easy to override.
-As a result, I use this setup (again, taken from Kieran Healy) to sidestep the restrictive rendering of authors in the standard `\maketitle` tag.
-After `author:`, enter `- name:` (no space before the dash) and fill in the field with the first author.
-On the next line, enter two spaces, followed by `affiliation:` and the institute or university affiliation of the first author.
-
-Do notice this can be repeated for however many co-authors there are to a manuscript.
-The rendered PDF will enter each co-author in a new line in a manner similar to journals like *American Journal of Political Science*, *American Political Science Review*, or *Journal of Politics*.
-
-The next two fields pertain to the frontmatter of a manuscript.
-They should also be intuitive for the reader.
-`abstract` should contain the abstract and `keywords` should contain some keywords that describe the research project.
-Both fields are optional, though are practically mandatory.
-Every manuscript requires an abstract and some journals---especially those published by Sage---request them with submitted manuscripts.
-My template also includes these keywords in the PDF's metadata.
-
-`date` comes standard with R Markdown and you can use it to enter the date of the most recent compile.
-I typically include the date of the last compile for a working paper in the `thanks:` field, so this field currently does not do anything in my Markdown-LaTeX manuscript template.
-I include it in my YAML as a legacy, basically.
-
-The next items are optional and cosmetic.
-`geometry:` is a standard option in LaTeX.
-I set the margins at one inch, and you probably should too.
-`fontfamily:` is optional, but I use it to specify the Palatino font.
-The default option is Computer Modern Roman.
-`fontsize:` sets, intuitively, the font size.
-The default is 10-point, but I prefer 11-point.
-`spacing:` is an optional field.
-If it is set as "double", the ensuing document is double-spaced.
-"single" is the only other valid entry for this field, though not including the entry in the YAML metadata amounts to singlespacing the document by default.
-Notice I have this "commented out" in the example code.
-
-The final two options pertain to the bibliography.
-`bibliography:` specifies the location of the .bib file, so the author could make citations in the manuscript.
-`biblio-style` specifies the type of bibliography to use.
-You'll typically set this as APSR.
-You could also specify the relative path of [my *Journal of Peace Research* .bst file](http://svmiller.com/miscellany/journal-of-peace-research-bst-file/) if you are submitting to that journal.
-
-# Getting Started with Markdown Syntax
-
-There are a lot of cheatsheets and reference guides for Markdown (e.g.
-[Adam Prichard](https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet), [Assemble](http://assemble.io/docs/Cheatsheet-Markdown.html), [Rstudio](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf), [Rstudio again](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf), [Scott Boms](http://scottboms.com/downloads/documentation/markdown_cheatsheet.pdf), [Daring Fireball](https://daringfireball.net/projects/markdown/syntax), among, I'm sure, several others).
-I encourage the reader to look at those, though I will retread these references here with a minimal working example below.
-
-```markdown
-
-Want an image? This will do it.
-
-![caption for my image](path/to/image.jpg)
-
-`fig_caption: yes` will provide a caption. Put that in the YAML metadata.
-
-- Find your biblatexkey in your bib file.
-- Put an @ before it, like @smith1984, or whatever it is.
-- @smith1984 creates an in-text citation (e.g. Smith (1984) says...)
-- [@smith1984] creates a parenthetical citation (Smith, 1984)
-
-That'll also automatically create a reference list at the end of the document.
-
-[In-text link to Google](http://google.com) as well.
-```
-
-That's honestly it. Markdown takes the chore of markup from your manuscript (hence: "Markdown").
-
-On that note, you could easily pass most LaTeX code through Markdown if you're writing a LaTeX document.
-However, you don't need to do this (unless you're using the math environment) and probably shouldn't anyway if you intend to share your document in HTML as well.
-
-# Using R Markdown with Knitr
-
-Perhaps the greatest intrigue of R Markdown comes with the [`knitr` package](http://yihui.name/knitr/) provided by @xie2013ddrk.
-In other words, the author can, if she chooses, do the analysis in the Markdown document itself and compile/execute it in R.
-
-Take, for example, this simple exercise using the `voteincome` data from the `Zelig` package.
-Suppose I want to explain the decision to vote using data from this package.
-I load in the data, clean the data, run the analyses, and present the results as a coefficient plot.
-
-Here's what this code looks like.
-All I did was create a code display, which starts with three *backticks* (i.e.
-those ticks next to the number 1 key on your keyboard) and ends with three backticks on another line.
-On the first line of backticks (i.e.
-to start the code display) enter `{r, eval=FALSE, tidy=TRUE}`.
-The `eval=FALSE` option just displays the R code (and does not run it), `tidy=TRUE` wraps long code so it does not run off the page.
-
-Within that code display, I enter my R code like this.
-
-```{r, eval=FALSE, tidy = TRUE}
-library(stevemisc)
-data(uniondensity)
-M1 <- lm(union ~ left + size + concen, data=uniondensity)
-library(arm)
-coefplot(M1)
-```
-
-The implications for workflow are faily substantial.
-Authors can rather quickly display the code they used to run the analyses in the document itself (likely in the appendix).
-As such, there's little guesswork for reviewers and editors in understanding what the author did in the analyses reported in the manuscript.
-
-It doesn't end there. In fact, here's what happens when `eval=FALSE` is omitted or changed to `eval=TRUE`. Now, the code runs within R. Observe.
-
-```{r, eval=TRUE, tidy = TRUE, cache=FALSE, fig.cap="A Coefficient Plot", message=F, warning=F}
-library(stevemisc)
-data(uniondensity)
-M1 <- lm(union ~ left + size + concen, data=uniondensity)
-library(arm)
-coefplot(M1)
-```
-
-To get `knitr` to present the results of a table, add `results="asis"` to the brackets to start the R code chunk. The ensuing output will look like this (though the table may come on the next page).
-
-```{r, eval=TRUE, tidy = TRUE, size="small", cache=FALSE, results="asis", message=F, warning=F}
-library(stevemisc)
-data(uniondensity)
-library(stargazer)
-M1 <- lm(union ~ left + size + concen, data=uniondensity)
-stargazer(M1, title="A Handsome Table", header=FALSE)
-```
-
-Adding `echo="FALSE"` inside the brackets to start the R chunk will omit the presentation of the R commands.
-It will just present the table.
-This provides substantial opportunity for authors in doing their analyses.
-Now, the analysis and presentation in the form of a polished manuscript can be effectively simultaneous.[^4]
-
-[^4]: I'm not sure if I'm ready to commit to this myself since my workflow is still largely derived from [Rob J.
-Hyndman's example](http://robjhyndman.com/hyndsight/workflow-in-r/).
-However, *knitr* has endless potential, especially when analyses can stored in cache, saved as chunks, or loaded in the preamble of a document to reference later in the manuscript.
-
-<!--
 # References 
@@
 \vspace*{-0.2in}
-\noindent
--->

### 2019-10-23 app.R
Use updated SDM process image; edit caption

@@ -198,7 +198,7 @@ ui <- fluidPage(
       imageOutput("imageSDM",
                    width=400,
                    height=275),
-      p(em("Structured Decision Making Process")),
+      p(em("Structured Decision Making process used as review framework")),
       h4("MSE Review Methods"),
       p("We conducted our search for MSEs in the SCI-EXPANDED index from
        Web of Science, searching for “management strategy evaluation” by topic
@@ -587,7 +587,7 @@ server <- function(input, output, session) {   # code to create output using ren
   )
   # Tab 6
   output$imageSDM<-renderImage({
-    filename<-normalizePath(file.path('./www',paste("SDMprocess.png")))
+    filename<-normalizePath(file.path('./www',paste("SDMProcessFramework.png")))
     list(src=filename,
          width=400,
          height=275)},deleteFile = FALSE)

### 2019-11-08 MSE manuscript.Rmd
draft manuscript

@@
     number_sections: false
+  bookdown::word_document2:
   bookdown::pdf_document2:
     number_sections: false
-biblio-style: apsr
+always_allow_html: yes
+biblio-style: apalike
 endnote: no
@@
 The positive aspects of MSE that have been listed include that it is “A tool that scientists and managers can use to simulate the workings of a fisheries system and allow them to test whether potential harvest strategies—or management procedures— can achieve pre-agreed management objectives.”[@Pew], that it “can be used to identify a ‘best’ management strategy among a set of candidate strategies" [@PuntLab] and “MSE has the advantage of being able to reveal the trade-offs among a range of possible management decisions, and should assist our Commissioners in determining whether the objectives should be weighted differently.
-Specifically, to provide the information on which to base a rational decision, given objectives, preferences, and attitudes to risk.”  – International Pacific Halibut Commission  Based on these descriptions MSE is a compelling tool to support and improve fisheries management decisions.
+Specifically, to provide the information on which to base a rational decision, given objectives, preferences, and attitudes to risk.”  [@IPHC].
+Based on these descriptions MSE is a compelling tool to support and improve fisheries management decisions.
 
@@
 
-Adaptive management developed in fisheries management as a means to improve the achievement of objectives through time, when learning can aid achieving those objectives.
-For example, the Millennium Ecosystem Assessment (2005) defines adaptive management as “A systematic process for continually improving management policies and practices by learning from the outcomes of previously employed policies and practices.
+Adaptive management developed in fisheries management as a means to improve the achievement of objectives through time, when learning can aid achieving those objectives [@ADM].
+For example, the Millennium Ecosystem Assessment [-@MEA] defines adaptive management as “A systematic process for continually improving management policies and practices by learning from the outcomes of previously employed policies and practices.
 In active adaptive management, management is treated as a deliberate experiment for purposes of learning.” We contend that scientific publication is essentially a process intended to support adaptive management, in that researchers gain knowledge from the decisions made and outcomes produced by previous work to inform their decisions.
+In the case of fisheries management, are scientific publications of management strategy evaluations supporting adaptive management?
 
 ### Objectives 
-Our overall objectives is to understanding of how MSEs utilize standard components of decision making processes and to what extent their documentation can support learning within the MSE practitioner community?
+Our overall objectives is to determine the extent that MSE documentation can support learning within the MSE practitioner community?
 Given that scientific publication is the primary means of communicating the results of MSE, we review the scientific literature to assess the extent to which the process and aims of MSE, and adaptive management are being achieved and supported by the scientific literature.
+To assess this we determine how MSEs utilize standard components of decision making processes.
 We use the structured decision making (SDM) process [@SDM] -- the decision making framework in which adaptive management occurs -- as our framework for evaluation.
@@
 |:--:|:--:|
-| Published (estimated*) | ~142 |
+| Published | 150 |
 | Random Sample | 30 |
 | Climate Change | 11 |  
+*Estimated by extrapolating from the assessed MSEs to the full set for now, the submission will contain an final count.
 
 ## Reviewing MSE documentation 
-We used the structured decision making framework as our framework for evaluating management strategy evaluation documents.
-[insert sdm figure].
-For ease of presentation we present the review methodology and results in four groups: consequences, objectives, stages, and process.
+We used the structured decision making framework [@conroy2013decision; @gregory2012structured] as our framework for evaluating management strategy evaluation documents (Figure \@ref(fig:SDMfigure)).
+The structured decision making processes is composed of the six process stages: problem definition, objective elicitation, alternative development, consequence prediction, trade-off analysis, and decision implementation.
 
-In the consequence group we report the drivers included as the primary factors influencing the predictions produced by the MSE model and report how the consequence predictions were made.
-The information reported here addresses what question the MSE was designed to address, and how it was addressed.
+```{r, SDMfigure, fig.cap='Review framework - structured decision making process'}
+knitr::include_graphics('figs/SDMProcessFramework.png')
+```
 
-In the objectives group we report the category of objectives considered in the MSE, and what type of objectives were considered.
-We noted four types of objectives: strategic objectives regarding the overall mission of the system in which the decision evaluated by the MSE falls, process objectives addressing how the management decision addressed by the MSE would be made, fundamental objectives addressing the goals of the management decision, and means objectives addressing means to achieving fundamental objectives of the management decision.
+We present the review methodology and results in four groups: consequences - which focuses on what was considered during the prediction stage, objectives - which focuses on what categories of objectives were considered, stages - which focuses on whether and how the stages of the process were completed, and participants - which focuses on how participation in the process was structured.
+These results were entered into a relational database for storage and evaluation, and is available via our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
 
-In the stages group we report whether the stages (Problem, Objectives, Alternatives, Consequences, Trade-offs, and Decision) of the structured decision making framework were explicitly completed and documented.
+The consequence group addresses what question the MSE was designed to address, and how it was addressed.
+We report the drivers included as the primary factors influencing the predictions produced by the MSE model, listing each driver a MSE included.
+We also report how the consequence predictions were made, with possible values of simulation modeling, dynamic programming, expert elicitation, mental models, and unknown.
+
+In the objectives group we address what the concerns, or goals of the MSE were.
+We record each objective considered, and report the type and category of objectives included in each MSE.
+Natural resource management problems typically include four types of objectives: conservation, resource use, i.e., yield in the fisheries context, economic, and social objectives [@runge2013].
+We sort objectives into four types: strategic objectives - which are concerned with the overall mission of the system in which the decision evaluated by the MSE falls, process objectives - which are concerned with how the management decision addressed by the MSE would be made, fundamental objectives - which address the goals of the management decision, and means objectives - which address means to achieving the fundamental objectives of the management decision.
+
+In the stages group we address how the MSE stages were completed.
+We report whether the stages of the structured decision making framework were explicitly completed and documented.
 We also report whether the MSE results were adopted in subsequent management.
-In many cases the process used to elicit objectives and alternatives was not explicit, but some information is provided or can be gleaned about who participated in the production of those components of the MSE.
-In those cases we record and report who those participants were, referring to them as subjective participants in the objectives or alternatives process.
 
-In the process group we report who lead the MSE, who participated in it, if the roles of participants were established and documented.
+In the participant group we address how participation in the MSE process was structured and who participated.
+We report who lead the MSE, who participated in it, and if the roles of participants were established and documented.
+In many cases the process used to elicit objectives and alternatives was not explicit.
+In these cases we used what information was provided to glean who participated in the production of those components of the MSE.
+We record and report who those participants were, referring to them as subjective participants in the objectives or alternatives process.
+In other cases where there was not enough information to glean who participated we report the participants as unknown.
 
-We collected additional information about the MSEs we reviewed.
-For a full description of the information recorded in database columns for this MSE review see the supplemental information (supplement), and for the full results see our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
+We collected additional information about the MSEs we reviewed beyond what is reported here.
+For a full description of the information recorded in the relational database see the supplemental information (supplement), and for the full data set and results see our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
 
@@
 <br>
+No strategic or process objectives were documented in the MSEs we reviewed.  There were roughly the same number of fundamental and means included in the MSEs (Table \@ref(tab:objectiveTypes)).
+```{r, objectiveTypes}
+obj.dataTable_pub %>%
+  filter(Objective=="Type") %>%
+  select(-Analysis, -Objective) %>% 
+  kable(caption = "Objective Types") %>% 
+  kable_styling()
+```
 
@@
 The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Table \@ref(tab:explicit)).
-The most documented stage was the alternative stage with 6 MSEs documenting this stage of the process.
+The consequence prediction stage -- which is the primary focus of MSE publication -- was always addressed.
+Apart from the consequence stage, the most documented stage was the alternative stage with 6 MSEs documenting this stage of the process.
 Three of the 30 MSEs explicitly documented a trade-off analysis and 2 of the 30 documented a management decision.
-One of the 30 documented either the roles of participants in the MSE process or that the results of the MSE were adopted.
+One MSE documented that the results of the MSE were adopted.
 
@@
 
-The primary participants and leaders of the MSE processes were scientists, with some lead by governments and management agencies (Figure \@ref(fig:participants)).
-Scientist lead 73 percent of the MSE processes, with another 23 percent lacking documentation that made the leader of the process clear.
+The primary participants and leaders of the MSE processes were scientists, while some MSE process were led by governments and management agencies (Figure \@ref(fig:participants)).
+Scientist led 73 percent of the MSE processes, with another 23 percent lacking documentation that made the leader of the process clear.
 
@@
 Members of the fishery participated in 30 percent, management and government representatives in 23 percent, representatives of independent institutions in 7 percent, and the members of the public in 3 percent of reviewed MSEs.
-Decision makers, those responsible for selecting the management plan to implement, were participants in 13 percent of MSEs.
+Decision makers -- those responsible for selecting the management plan to implement -- were participants in 13 percent of MSEs.
 Experts, a possible source of data or predictions, participated in 3 percent of the MSEs reviewed.
-Participants involved to the assist the MSE process itself, facilitators and analysts participated in 10 and 3 percent of MSE processes respectively.
+Participants involved to the assist the MSE process itself, i.e., facilitators and analysts, participated in 10 and 3 percent of MSE processes respectively.
 
 Objectives and alternatives were elicited from a variety of participants.
-About 10 percent of the objectives and alternatives explicitly were produced by stakeholders in the fisheries, coming from management, government, or the fishery.
+About 10 percent of the objectives and alternatives explicitly were produced by stakeholders in the fisheries, coming from management, government, or fishery sources.
 In most cases the source of the objectives and alternatives was not explicit.
-In those cases scientists selected these components most frequently.
+In those cases scientists selected these components most frequently, while in some cases the source was unknown.
 
@@
 
-We review published MSEs including climate change as a driver against the steps of the SDM framework to:  Highlight unique aspects of MSEs that model climate change as a driver to inform fisheries management in a changing climate.
+We review published MSEs including climate change as a driver against the stages in a SDM process to:  Highlight unique aspects of MSEs that model climate change as a driver to help inform fisheries management
 
-We found that as of the end of December 2018 11 MSEs included climate change as a driver. These MSEs have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).
+We filtered the original set of MSE by searching for the term "climate change" within the set.
+We found that, as of the end of December 2018, 11 MSEs included climate change as a driver.
+We reviewed the climate change MSEs using the same methodology described above.
+These MSEs have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).
 
@@
 
-Climate change MSEs were more likely to explicitly document the alternative elicitation and trade-off analysis steps, but otherwise were less likely to document steps in the decision process.
+Climate change MSEs were more likely to explicitly document the alternative elicitation and trade-off analysis stages, but otherwise were less likely to document stages in the decision process.
 ```{r, CCstages}
@@
 
-Including climate change in a MSE appears to reduce the complexity of other stages in the MSE process, as on average climate change MSEs consider fewer drivers, fewer and less diverse objectives, and are less explicitly documented at most stages of the process.
-If the MSE process is time or effort limited, the inclusion of climate change may be reducing the time and effort that otherwise is invested in other stages of the MSE process.
+Including climate change in a MSE appears to reduce the complexity of other stages in the MSE process.
+On average climate change MSEs consider fewer drivers, consider fewer and less diverse objectives, and explicitly document fewer stages of the MSE process.
+If the MSE process is time or effort limited, the inclusion of climate change may utlizing time and effort that in other case was invested elsewhere in the MSE process.
 
@@
 Management strategy evaluations typically do assess the consequences of uncertainty for achieving management goals, but other factors influencing the consequences of management strategies get more sporadic treatment.
-[more about including uncertainty].
 Data limitation is a frequent concern in fisheries management.
@@
 Further inclusion of this driver, and additional use of value of information analyses could expand the use of MSE as a tool to design cost-effective fisheries monitoring plans.
-While discussions of fisheries research and management seem to be move toward ecosystem effects, spatial models, and multiple species models, these characteristics were rarely considered in the set of MSEs we reviewed.
+While discussions of fisheries research and management seem to be moving toward ecosystem effects, spatial models, and multiple species models, these characteristics were rarely considered in the set of MSEs we reviewed.
 
 Managing natural resources entails achieving the objectives of people interacting with a natural resource, so decision processes are likely to produce more desirable results when they include peoples objectives.
-These objectives are likely to include conservation, yield, economic, and social concerns, however the MSE we reviewed don't cover all categories of objectives, as yield and economic objectives were not included in about one quarter of MSEs, and social objectives were absent from all but 2 of the 30 evaluations.
-While we are unable to know the cause for this, our results may provide some clues.
+These objectives are likely to include conservation, yield, economic, and social concerns, however the MSEs we reviewed don't cover all categories of objectives, as yield and economic objectives were not included in about one quarter of MSEs, and social objectives were absent from all but 2 of the 30 evaluations.
+While we are unable to know the cause for this with more explicit documentatin of MSE processes, our results may provide some clues.
 
@@
 That is, facilitators and decision analysts were rarely participants in MSE processes.
-The inclusion of facilitators in decision processes aid information gain and conflict resolution, increase trust between participants, and lead to more beneficial environmental and social outcomes[@Vente].
+The inclusion of facilitators in decision processes aid information gain and conflict resolution, increase trust between participants, and lead to more beneficial environmental and social outcomes [@Vente].
 Facilitators and decision analysts also aid decision processes by helping to diagnose the problem, which can result in selecting more appropriate analysis techniques for the problem type.
-Finally, the additional structure and focus on the decision process would likely result in more explicit documentation when facilitators and decision analysts participate in MSE, as resulted in the MSE reviewed here (cite Williams et al 2011, Smith et al 2013, and Kolody et al 2008).
+Finally, the additional structure and focus on the decision process would likely result in more explicit documentation when facilitators and decision analysts participate in MSE, as resulted in some of the MSEs reviewed here [@williams2011; @smith2013; @kolody2008].
 
-The relative dearth of explicit documentation of the MSE processes may hinder learning by MSE practitioners.
+The relative dearth of explicit documentation we found in our review of MSE publications may hinder learning by MSE practitioners.
 While the modeling was clearly documented, the MSE process steps were often inexplicit.
@@
 
+The MSE process may be aided by utilizing the techniques from the structured decision making approach to decision analysis (Cummings, in prep).
+Utilizing the structured decision making approach eases the documentation process by breaking decision making into more discrete stages.
+There may be practical considerations, such as time or budgetary constraints that limit how complex the stages of a MSE are, and the degree of documentation that occurs.
+Where the expertise and budget is available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.
+
 ### Explore and expand upon this review
@@
 We hope that as MSE are conducted in the future they are entered here so that MSE practitioners can learn from each other.
-
-### References
 

### 2019-11-08 MSE manuscript.Rmd
Fix number of published articles to final count of 151

@@
 |:--:|:--:|
-| Published | 150 |
+| Published | 151 |
 | Random Sample | 30 |

### 2019-11-08 app.R

@@ -178,18 +178,17 @@ ui <- fluidPage(
       h1("An Assessment of Management Strategy Evaluations"),# title of the page
       hr(),
       h4("Background"),
-      p("Fisheries management has mostly focused on fishing impacts
-      with ecosystem status viewed as a background constant."),
-      p("The increasing rate of climate change is changing this
-      dynamic, bringing ecosystem status to the forefront of fisheries
-      management."),
-      p("Management strategy evaluation (MSE) is 'widely considered to
+      p("The calls for fisheries management to utilize management strategy
+      evaluation are both aspirational and estimable in their aim and vision.
+      Management strategy evaluation (MSE) is 'widely considered to
       be the most appropriate way to evaluate the trade‐offs achieved by
       alternative management strategies and to assess the consequences of
       uncertainty for achieving management goals' (Punt et al. 2014). Thus,
-      MSE is a compelling tool to assess climate change impacts and test
-      climate-ready options for fisheries management decisions Adaptive
-      management arose to address uncertainties and accelerate progress
+      MSE is a compelling tool to assess fisheries management procedures.  For
+      example, in our changing oceans and increasing climate change impacts MSE
+      can evaluate climate-ready options for fisheries management decisions.
+
+      Adaptive management arose to address uncertainties and accelerate progress
       towards meeting management objectives. We used the structured decision
       making (SDM) process - the decision making framework in which adaptive
       management occurs - as our framework, how do published MSE projects
@@ -592,8 +591,8 @@ server <- function(input, output, session) {   # code to create output using ren
          width=400,
          height=275)},deleteFile = FALSE)
   output$MSEcounts <- renderTable({
-    tibble("MSE type"=c("Published (estimated)","Random Sample","Climate Change"),
-               "Count"=c(142,30,11))
+    tibble("MSE type"=c("Published","Random Sample","Climate Change"),
+               "Count"=c(151,30,11))
 })
 }
 

### 2020-03-24 app.R
move data load to later in script

@@ -38,18 +38,6 @@ library(DBI) # database interface
 # # Open connection
 # con <- getSqlConnection()
 # 
-# # To close the connection
-# dbDisconnect(con)
-
-load("MSEreview.RData")
-
-
-# Get map background for plotting the map
-world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
-
-
-##### Data Processing #####
-
 # # Obtain data tables
 # study<-dbReadTable(con,"tblStudy")
 # # study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
@@ -65,6 +53,14 @@ world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a
 # # To close the connection
 # dbDisconnect(con)
 
+load("MSEreview.RData")
+
+
+# Get map background for plotting the map
+world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
+
+
+##### Data Processing #####
 
 # Edit tables for joining
 study<-study %>% 

### 2020-03-24 R/MSE manuscript.Rmd
Rearrange folders to Wilson et al 2017 structure

--- 2019-11-08 MSE manuscript.Rmd
@@
 We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019.
-This search returned 253 results.
+This search returned 264 results.
 We reviewed a random sample of 30 articles that document a MSE, removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
-Of the 253 articles, 11 included climate change as a driver of fishery status, and after removing other articles we estimate 110-160 articles document a MSE.
+Of the 264 articles, 11 included climate change as a driver of fishery status, and after removing other articles 154 articles document a MSE.
 
@@
 |:--:|:--:|
-| Published | 151 |
+| Published | 154 |
 | Random Sample | 30 |

### 2020-03-24 data/svm-latex-ms.tex
Rearrange folders to Wilson et al 2017 structure

\documentclass[11pt,]{article}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}

  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother

\setcounter{secnumdepth}{0}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

\title{A Pandoc Markdown Article Starter and Template \thanks{Replication files are available on the author's Github account
(\url{http://github.com/svmiller}). \textbf{Current version}: July 26,
2019; \textbf{Corresponding author}:
\href{mailto:svmille@clemson.edu}{\nolinkurl{svmille@clemson.edu}}.}  }

\author{\Large Steven V. Miller\vspace{0.05in} \newline\normalsize\emph{Clemson University}  }

\date{}

\usepackage{titlesec}

\titleformat*{\section}{\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\itshape}
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

\usepackage{natbib}
\bibliographystyle{apsr}
\usepackage[strings]{underscore} % protect underscores in most circumstances

\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{hyperref}

% move the hyperref stuff down here, after header-includes, to allow for - \usepackage{hyperref}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Steven V. Miller (Clemson University)},
             pdfkeywords = {pandoc, r markdown, knitr},  
            pdftitle={A Pandoc Markdown Article Starter and Template},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

% add tightlist ----------
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

\begin{document}

% \pagenumbering{arabic}% resets `page` counter to 1 
%
% \maketitle

{% \usefont{T1}{pnc}{m}{n}
\setlength{\parindent}{0pt}
\thispagestyle{plain}
{\fontsize{18}{20}\selectfont\raggedright 
\maketitle  % title \par

}

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Steven V. Miller} \hskip 15pt \emph{\small Clemson University}

}

}

\begin{abstract}

    \hbox{\vrule height .2pt width 39.14pc}

    \vskip 8.5pt % \small

\noindent This document provides an introduction to R Markdown, argues for its
benefits, and presents a sample manuscript template intended for an
academic audience. I include basic syntax to R Markdown and a minimal
working example of how the analysis itself can be conducted within R
with the \texttt{knitr} package.

\vskip 8.5pt \noindent \emph{Keywords}: pandoc, r markdown, knitr \par

    \hbox{\vrule height .2pt width 39.14pc}

\end{abstract}

\vskip -8.5pt

 % removetitleabstract

\noindent

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Academic workflow, certainly in political science, is at a crossroads.
The \emph{American Journal of Political Science} (\emph{AJPS}) announced
a (my words)
\href{http://ajps.org/2015/03/26/the-ajps-replication-policy-innovations-and-revisions/}{``show
your work'' initiative} in which authors who are tentatively accepted
for publication at the journal must hand over the raw code and data that
produced the results shown in the manuscript. The editorial team at
\emph{AJPS} then reproduces the code from the manuscript. Pending
successful replication, the manuscript moves toward publication. The
\emph{AJPS} might be at the fore of this movement, and it could be the
most aggressive among political science journals, but other journals in
our field have signed the joint
\href{http://www.dartstatement.org/}{Data Access \& Research
Transparency} (DART) initiative. This, at a bare minimum, requires
uploading code from quantitatively-oriented published articles to
in-house directories hosted by the journal or to services like
\href{http://dataverse.org/}{Dataverse}.

There are workflow implications to the Lacour controversy as well.
Political science, for the foreseeable future, will struggle with the
extent of
\href{http://stanford.edu/~dbroock/broockman_kalla_aronow_lg_irregularities.pdf}{the
data fraud perpetrated by Michael Lacour} in an article co-authored with
Donald P. Green in \emph{Science}, the general scientific journal of
record in the United States. A failure to reproduce LaCour's results
with different samples uncovered a comprehensive effort by LaCour to
``fake'' data that provided results to what we felt or believed to be
true \href{http://chronicle.com/article/LAffaire-LaCour/230905/}{(i.e.
``truthiness'')}. However,
\href{http://kieranhealy.org/blog/archives/2015/05/20/fake-science-real-consequences/}{fake
data can have real consequences} for both the researcher and those who
want to learn from it and use it for various purposes. Even research
done honestly may suffer the same fate if researchers are not diligent
in their workflow.

These recent events underscore the DART push and cast a shadow over our
workflow. However, good workflow has always been an issue in our
discipline. Cloud storage services like
\href{http://www.dropbox.com}{Dropbox} are still relatively new among
political scientists. Without cloud storage, previous workflow left open
the possibility that work between a home computer and an office computer
was lost as a function of a corrupted thumb drive, an overheated power
supply, or, among other things, the wave of viruses that
\href{http://money.cnn.com/2003/11/05/technology/microsoftbounty/}{would
particularly affect Microsoft users every summer}. Social sciences,
\href{http://kieranhealy.org/blog/archives/2014/01/23/plain-text/}{unlike
engineering}, have traditionally relied on software like Microsoft Word
for manuscript preparation though any word processor reduces workflow to
a series of clicks and strokes on a keyboard. This is
\href{http://www.nytimes.com/2013/04/19/opinion/krugman-the-excel-depression.html}{a
terrible way to track changes} or maintain version control. The addition
of collaborators only compounds all the aforementioned issues. The
proverbial left hand may not know what the right hand is doing.

I think there is reason for optimism. We only struggle with it now
because we have tools like \href{http://rmarkdown.rstudio.com/}{R
Markdown} and \href{http://pandoc.org/}{Pandoc}, more generally, that
make significant strides in workflow. LaTeX resolved earlier issues of
corrupted binary files by reducing documents to raw markup that was
little more than raw text and revisions that could be easily kept as
\href{http://tex.stackexchange.com/questions/11177/how-to-write-hidden-notes-in-a-latex-file}{``commented''
text}. However, for all its benefits (including pretty PDFs),
\href{http://www-rohan.sdsu.edu/~aty/bibliog/latex/gripe.html}{LaTeX is
\emph{ugly} code} and does not provide means of seamlessly working with
the actual data analysis itself. R Markdown both eliminates markup and
allows the author and her collaborators to write and reproduce the
manuscript in one fell swoop.

\hypertarget{getting-started-with-yaml}{%
\section{Getting Started with YAML}\label{getting-started-with-yaml}}

The lion's share of a R Markdown document will be raw text, though the
front matter may be the most important part of the document. R Markdown
uses \href{http://www.yaml.org/}{YAML} for its metadata and the fields
differ from
\href{http://svmiller.com/blog/2015/02/moving-from-beamer-to-r-markdown/}{what
an author would use for a Beamer presentation}. I provide a sample YAML
metadata largely taken from this exact document and explain it below.

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{---}
\NormalTok{output}\OperatorTok{:}\StringTok{ }
\StringTok{  }\NormalTok{pdf_document}\OperatorTok{:}
\StringTok{    }\NormalTok{citation_package}\OperatorTok{:}\StringTok{ }\NormalTok{natbib}
\NormalTok{    keep_tex}\OperatorTok{:}\StringTok{ }\NormalTok{true}
\NormalTok{    fig_caption}\OperatorTok{:}\StringTok{ }\NormalTok{true}
\NormalTok{    latex_engine}\OperatorTok{:}\StringTok{ }\NormalTok{pdflatex}
\NormalTok{    template}\OperatorTok{:}\StringTok{ }\ErrorTok{~/}\NormalTok{Dropbox}\OperatorTok{/}\NormalTok{miscelanea}\OperatorTok{/}\NormalTok{svm}\OperatorTok{-}\NormalTok{r}\OperatorTok{-}\NormalTok{markdown}\OperatorTok{-}\NormalTok{templates}\OperatorTok{/}\NormalTok{svm}\OperatorTok{-}\NormalTok{latex}\OperatorTok{-}\NormalTok{ms.tex}
\NormalTok{title}\OperatorTok{:}\StringTok{ "A Pandoc Markdown Article Starter and Template"}
\NormalTok{thanks}\OperatorTok{:}\StringTok{ "Replication files are available on the author's Github account..."}
\NormalTok{author}\OperatorTok{:}
\OperatorTok{-}\StringTok{ }\NormalTok{name}\OperatorTok{:}\StringTok{ }\NormalTok{Steven V. Miller}
\NormalTok{  affiliation}\OperatorTok{:}\StringTok{ }\NormalTok{Clemson University}
\OperatorTok{-}\StringTok{ }\NormalTok{name}\OperatorTok{:}\StringTok{ }\NormalTok{Mary Margaret Albright}
\NormalTok{  affiliation}\OperatorTok{:}\StringTok{ }\NormalTok{Pendelton State University}
\OperatorTok{-}\StringTok{ }\NormalTok{name}\OperatorTok{:}\StringTok{ }\NormalTok{Rembrandt Q. Einstein}
\NormalTok{  affiliation}\OperatorTok{:}\StringTok{ }\NormalTok{Springfield University}
\NormalTok{abstract}\OperatorTok{:}\StringTok{ "This document provides an introduction to R Markdown, argues for its..."}
\NormalTok{keywords}\OperatorTok{:}\StringTok{ "pandoc, r markdown, knitr"}
\NormalTok{date}\OperatorTok{:}\StringTok{ "`r format(Sys.time(), '%B %d, %Y')`"}
\NormalTok{geometry}\OperatorTok{:}\StringTok{ }\NormalTok{margin=1in}
\NormalTok{fontfamily}\OperatorTok{:}\StringTok{ }\NormalTok{mathpazo}
\NormalTok{fontsize}\OperatorTok{:}\StringTok{ }\NormalTok{11pt}
\CommentTok{# spacing: double}
\NormalTok{bibliography}\OperatorTok{:}\StringTok{ "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`"}
\NormalTok{biblio}\OperatorTok{-}\NormalTok{style}\OperatorTok{:}\StringTok{ }\NormalTok{apsr}
\OperatorTok{---}
\end{Highlighting}
\end{Shaded}
\texttt{output:} will tell R Markdown we want a PDF document rendered
with LaTeX. Since we are adding a fair bit of custom options to this
call, we specify \texttt{pdf\_document:} on the next line (with,
importantly, a two-space indent). We specify additional output-level
options underneath it, each are indented with four spaces.
\texttt{citation\_package:\ natbib} tells R Markdown to use
\texttt{natbib} to handle bibliographic citations.\footnote{R Markdown
  can use Pandoc's native bibliography management system or even
  \texttt{biblatex}, but I've found that it chokes with some of the more
  advanced stuff I've done with my .bib file over the years. For
  example, I've been diligent about special characters (e.g.~umlauts and
  acute accents) in author names in my .bib file, but Pandoc's native
  citation system will choke on these characters in a .bib file. I
  effectively need \texttt{natbib} for my own projects.} Thereafter, the
next line (\texttt{keep\_tex:\ true}) tells R Markdown to render a raw
\texttt{.tex} file along with the PDF document. This is useful for both
debugging and the publication stage, when the editorial team will ask
for the raw \texttt{.tex} so that they could render it and later provide
page proofs. The next line \texttt{fig\_caption:\ true} tells R Markdown
to make sure that whatever images are included in the document are
treated as figures in which our caption in brackets in a Markdown call
is treated as the caption in the figure. The next line
(\texttt{latex\_engine:\ pdflatex}) tells R Markdown to use pdflatex and
not some other option like \texttt{lualatex}. For my template, I'm
pretty sure this is mandatory.\footnote{The main reason I still use
  \texttt{pdflatex} (and most readers probably do as well) is because of
  LaTeX fonts.
  \href{http://www-rohan.sdsu.edu/~aty/bibliog/latex/gripe.html}{Unlike
  others}, I find standard LaTeX fonts to be appealing.}
The next line (\texttt{template:\ ...}) tells R Markdown to use my
custom LaTeX template.\footnote{Notice that the path is relative. The
  user can, if she wishes, install this in the default Pandoc directory.
  I don't think this is necessary. Just be mindful of wherever the
  template is placed. Importantly, \texttt{\textasciitilde{}} is used in
  R to find the home directory (not necessarily the working directory).
  It is equivalent to saying \texttt{/home/steve} in Linux, or
  \texttt{/Users/steve} on a Mac, in my case.} While I will own any
errors in the code, I confess to ``Frankensteining'' this template from
\href{https://github.com/jgm/pandoc-templates}{the default LaTeX
template} from Pandoc,
\href{https://github.com/kjhealy/pandoc-templates/tree/master/templates}{Kieran
Healy's LaTeX template}, and liberally using raw TeX from the
\href{https://www.acm.org/publications/article-templates/acm-latex-style-guide}{Association
for Computing Machinery's (ACM) LaTeX template}. I rather like that
template since it resembles standard manuscripts when they are published
in some of our more prominent journals. I will continue with a
description of the YAML metadata in the next paragraph, though invite
the curious reader to scroll to the end of the accompanying post to see
the PDF this template produces.
The next fields get to the heart of the document itself. \texttt{title:}
is, intuitively, the title of the manuscript. Do note that fields like
\texttt{title:} do not have to be in quotation marks, but must be in
quotation marks if the title of the document includes a colon. That
said, the only reason to use a colon in an article title is if it is
followed by a subtitle, hence the optional field (\texttt{subtitle:}).
Notice I ``comment out'' the subtitle in the above example with a pound
sign since this particular document does not have a subtitle. If
\texttt{thanks:} is included and has an accompanying entry, the ensuing
title of the document gets an asterisk and a footnote. This field is
typically used to advise readers that the document is a working paper or
is forthcoming in a journal.
The next field (\texttt{author:}) is a divergence from standard YAML,
but I think it is useful. I will also confess to pilfering this idea
from Kieran Healy's template. Typically, multiple authors for a given
document are separated by an \texttt{\textbackslash{}and} in this field.
However, standard LaTeX then creates a tabular field separating multiple
authors that is somewhat restrictive and not easy to override. As a
result, I use this setup (again, taken from Kieran Healy) to sidestep
the restrictive rendering of authors in the standard
\texttt{\textbackslash{}maketitle} tag. After \texttt{author:}, enter
\texttt{-\ name:} (no space before the dash) and fill in the field with
the first author. On the next line, enter two spaces, followed by
\texttt{affiliation:} and the institute or university affiliation of the
first author.
Do notice this can be repeated for however many co-authors there are to
a manuscript. The rendered PDF will enter each co-author in a new line
in a manner similar to journals like \emph{American Journal of Political
Science}, \emph{American Political Science Review}, or \emph{Journal of
Politics}.
The next two fields pertain to the frontmatter of a manuscript. They
should also be intuitive for the reader. \texttt{abstract} should
contain the abstract and \texttt{keywords} should contain some keywords
that describe the research project. Both fields are optional, though are
practically mandatory. Every manuscript requires an abstract and some
journals---especially those published by Sage---request them with
submitted manuscripts. My template also includes these keywords in the
PDF's metadata.
\texttt{date} comes standard with R Markdown and you can use it to enter
the date of the most recent compile. I typically include the date of the
last compile for a working paper in the \texttt{thanks:} field, so this
field currently does not do anything in my Markdown-LaTeX manuscript
template. I include it in my YAML as a legacy, basically.
The next items are optional and cosmetic. \texttt{geometry:} is a
standard option in LaTeX. I set the margins at one inch, and you
probably should too. \texttt{fontfamily:} is optional, but I use it to
specify the Palatino font. The default option is Computer Modern Roman.
\texttt{fontsize:} sets, intuitively, the font size. The default is
10-point, but I prefer 11-point. \texttt{spacing:} is an optional field.
If it is set as ``double'', the ensuing document is double-spaced.
``single'' is the only other valid entry for this field, though not
including the entry in the YAML metadata amounts to singlespacing the
document by default. Notice I have this ``commented out'' in the example
code.
The final two options pertain to the bibliography.
\texttt{bibliography:} specifies the location of the .bib file, so the
author could make citations in the manuscript. \texttt{biblio-style}
specifies the type of bibliography to use. You'll typically set this as
APSR. You could also specify the relative path of
\href{http://svmiller.com/miscellany/journal-of-peace-research-bst-file/}{my
\emph{Journal of Peace Research} .bst file} if you are submitting to
that journal.
\hypertarget{getting-started-with-markdown-syntax}{%
\section{Getting Started with Markdown
Syntax}\label{getting-started-with-markdown-syntax}}
There are a lot of cheatsheets and reference guides for Markdown (e.g.
\href{https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet}{Adam
Prichard},
\href{http://assemble.io/docs/Cheatsheet-Markdown.html}{Assemble},
\href{https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf}{Rstudio},
\href{https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf}{Rstudio
again},
\href{http://scottboms.com/downloads/documentation/markdown_cheatsheet.pdf}{Scott
Boms}, \href{https://daringfireball.net/projects/markdown/syntax}{Daring
Fireball}, among, I'm sure, several others). I encourage the reader to
look at those, though I will retread these references here with a
minimal working example below.
\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{# Introduction}
\NormalTok{**Lorem ipsum** dolor *sit amet*. }
\NormalTok{- }\FloatTok{Single asterisks italicize text *like this*. }
\FloatTok{- Double asterisks embolden text **like this**.}
\NormalTok{Start a new paragraph with a blank line separating paragraphs.}
\NormalTok{- }\FloatTok{This will start an unordered list environment, and this will be the first item.}
\FloatTok{- This will be a second item.}
\FloatTok{- A third item.}
\FloatTok{    - Four spaces and a dash create a sublist and this item in it.}
\FloatTok{- The fourth item.}
\FloatTok{    }
\FloatTok{1. This starts a numerical list.}
\FloatTok{2. This is no. 2 in the numerical list.}
\FloatTok{    }
\FloatTok{# This Starts A New Section}
\FloatTok{## This is a Subsection}
\FloatTok{### This is a Subsubsection}
\FloatTok{#### This starts a Paragraph Block.}
\NormalTok{>}\DataTypeTok{ This will create a block quote, if you want one.}
\NormalTok{Want a table? This will create one.}
\NormalTok{Table Header  | Second Header}
\NormalTok{------------- | -------------}
\NormalTok{Table Cell    | Cell 2}
\NormalTok{Cell 3        | Cell 4 }
\NormalTok{Note that the separators *do not* have to be aligned.}
\NormalTok{Want an image? This will do it.}
\AlertTok{![caption for my image](path/to/image.jpg)}
\BaseNTok{`fig_caption: yes`}\NormalTok{ will provide a caption. Put that in the YAML metadata.}
\NormalTok{Almost forgot about creating a footnote.}\OtherTok{[^1]}\NormalTok{ This will do it again.}\OtherTok{[^2]}
\OtherTok{[^1]}\NormalTok{: The first footnote}
\OtherTok{[^2]}\NormalTok{: The second footnote}
\NormalTok{Want to cite something? }
\NormalTok{- }\FloatTok{Find your biblatexkey in your bib file.}
\FloatTok{- Put an @ before it, like @smith1984, or whatever it is.}
\FloatTok{- @smith1984 creates an in-text citation (e.g. Smith (1984) says...)}
\FloatTok{- [@smith1984] creates a parenthetical citation (Smith, 1984)}
\NormalTok{That'll also automatically create a reference list at the end of the document.}
\OtherTok{[In-text link to Google](http://google.com)}\NormalTok{ as well.}
\end{Highlighting}
\end{Shaded}
That's honestly it. Markdown takes the chore of markup from your
manuscript (hence: ``Markdown'').
On that note, you could easily pass most LaTeX code through Markdown if
you're writing a LaTeX document. However, you don't need to do this
(unless you're using the math environment) and probably shouldn't anyway
if you intend to share your document in HTML as well.
\hypertarget{using-r-markdown-with-knitr}{%
\section{Using R Markdown with
Knitr}\label{using-r-markdown-with-knitr}}
Perhaps the greatest intrigue of R Markdown comes with the
\href{http://yihui.name/knitr/}{\texttt{knitr} package} provided by
\citet{xie2013ddrk}. In other words, the author can, if she chooses, do
the analysis in the Markdown document itself and compile/execute it in
R.
Take, for example, this simple exercise using the \texttt{voteincome}
data from the \texttt{Zelig} package. Suppose I want to explain the
decision to vote using data from this package. I load in the data, clean
the data, run the analyses, and present the results as a coefficient
plot.
Here's what this code looks like. All I did was create a code display,
which starts with three \emph{backticks} (i.e.~those ticks next to the
number 1 key on your keyboard) and ends with three backticks on another
line. On the first line of backticks (i.e.~to start the code display)
enter \texttt{\{r,\ eval=FALSE,\ tidy=TRUE\}}. The \texttt{eval=FALSE}
option just displays the R code (and does not run it),
\texttt{tidy=TRUE} wraps long code so it does not run off the page.
Within that code display, I enter my R code like this.
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(stevemisc)}
\KeywordTok{data}\NormalTok{(uniondensity)}
\NormalTok{M1 <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(union }\OperatorTok{~}\StringTok{ }\NormalTok{left }\OperatorTok{+}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{concen, }\DataTypeTok{data =}\NormalTok{ uniondensity)}
\KeywordTok{library}\NormalTok{(arm)}
\KeywordTok{coefplot}\NormalTok{(M1)}
\end{Highlighting}
\end{Shaded}
The implications for workflow are faily substantial. Authors can rather
quickly display the code they used to run the analyses in the document
itself (likely in the appendix). As such, there's little guesswork for
reviewers and editors in understanding what the author did in the
analyses reported in the manuscript.
It doesn't end there. In fact, here's what happens when
\texttt{eval=FALSE} is omitted or changed to \texttt{eval=TRUE}. Now,
the code runs within R. Observe.
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(stevemisc)}
\KeywordTok{data}\NormalTok{(uniondensity)}
\NormalTok{M1 <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(union }\OperatorTok{~}\StringTok{ }\NormalTok{left }\OperatorTok{+}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{concen, }\DataTypeTok{data =}\NormalTok{ uniondensity)}
\KeywordTok{library}\NormalTok{(arm)}
\KeywordTok{coefplot}\NormalTok{(M1)}
\end{Highlighting}
\end{Shaded}
\begin{figure}
\centering
\includegraphics{figs/unnamed-chunk-3.pdf}
\caption{A Coefficient Plot}
\end{figure}
To get \texttt{knitr} to present the results of a table, add
\texttt{results="asis"} to the brackets to start the R code chunk. The
ensuing output will look like this (though the table may come on the
next page).
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(stevemisc)}
\KeywordTok{data}\NormalTok{(uniondensity)}
\KeywordTok{library}\NormalTok{(stargazer)}
\NormalTok{M1 <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(union }\OperatorTok{~}\StringTok{ }\NormalTok{left }\OperatorTok{+}\StringTok{ }\NormalTok{size }\OperatorTok{+}\StringTok{ }\NormalTok{concen, }\DataTypeTok{data =}\NormalTok{ uniondensity)}
\KeywordTok{stargazer}\NormalTok{(M1, }\DataTypeTok{title =} \StringTok{"A Handsome Table"}\NormalTok{, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}
\begin{table}[!htbp] \centering 
  \caption{A Handsome Table} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & union \\ 
\hline \\[-1.8ex] 
 left & 0.270$^{***}$ \\ 
  & (0.076) \\ 
  & \\ 
 size & $-$6.460 \\ 
  & (3.794) \\ 
  & \\ 
 concen & 0.351 \\ 
  & (19.253) \\ 
  & \\ 
 Constant & 97.591 \\ 
  & (57.481) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 20 \\ 
R$^{2}$ & 0.743 \\ 
Adjusted R$^{2}$ & 0.694 \\ 
Residual Std. Error & 10.369 (df = 16) \\ 
F Statistic & 15.381$^{***}$ (df = 3; 16) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}
Adding \texttt{echo="FALSE"} inside the brackets to start the R chunk
will omit the presentation of the R commands. It will just present the
table. This provides substantial opportunity for authors in doing their
analyses. Now, the analysis and presentation in the form of a polished
manuscript can be effectively simultaneous.\footnote{I'm not sure if I'm
  ready to commit to this myself since my workflow is still largely
  derived from
  \href{http://robjhyndman.com/hyndsight/workflow-in-r/}{Rob J.
  Hyndman's example}. However, \emph{knitr} has endless potential,
  especially when analyses can stored in cache, saved as chunks, or
  loaded in the preamble of a document to reference later in the
  manuscript.}
\newpage
\singlespacing 
\bibliography{master.bib}
\end{document}

### 2020-03-24 Analysis.R

@@ -1,453 +1,0 @@
-# MSE Problem Framing 
-# Literature Review Data Analysis
-# 1/4/2018, JWC
-
-##### 
-# Meta Data and setup
-# Description: Import and analyze MSE documentation from database
-# Value: TBD
-
-# load libraries
-library(tidyverse) # upgrade to base R
-library(RODBC) # connect to databases
-library(shiny) # Shiny App
-library(shinythemes) # web themes for Shiny
-library(DT) # Data Table
-library(purrr) # for loop alternative
-library(ggmap) # map plotting
-library(mapdata) # basemap generation
-
-##### 
-# Script
-
-#####
-###-- Data Processing --###
-wd<-getwd()
-# Establish connection to Access database
-channel<-odbcDriverConnect(paste0("Driver={Microsoft Access Driver (*.mdb, *.accdb)};
-                           DBQ=",wd,"/DB files - Access/MSE Problem Framing_be.accdb"))
-
-# Obtain data tables
-study<-as_tibble(sqlFetch(channel,sqtable = "tblStudy"))
-mgmt<-as_tibble(sqlFetch(channel,sqtable = "tblStudyManagementTools"))
-obj<-as_tibble(sqlFetch(channel,sqtable = "tblStudyObjectives"))
-fields<-as_tibble(sqlFetch(channel,sqtable = "tblStudyFields"))
-odbcCloseAll()
-
-#####
-# Edit tables for joining
-study<-study %>% 
-  unite(Citation,c(Authors,YearPub),sep=" ",remove=F)
-
-study.join<-study %>% 
-  select(ID,Citation)
-
-mgmt.join<-mgmt %>%
-  group_by(fkStudyID) %>%
-  summarise(ManagementTool = toString(sort(unique(MPManagementTool))),
-            AlternativesEvaluated = toString(sort(unique(MPAlternativesEvaluated))))
-
-obj.join<-obj %>% 
-  group_by(fkStudyID) %>%
-  summarise(ObjectiveCategories = toString(sort(unique(ObjCategory))))
-
-# Join tables
-data<-full_join(study,mgmt.join,by=c("ID"="fkStudyID"))
-data<-full_join(data,obj.join,by=c("ID"="fkStudyID")) %>% 
-  select(-c("ID"))
-obj.data<-right_join(study.join,obj,by=c("ID"="fkStudyID"))%>% 
-  select(-c("ID","ID.y"))
-
-# Move comment column to the end
-data<-data %>% 
-  select(-'Comments', 'Comments')
-
[…]

### 2020-03-24 AnalysisExcel.R

@@ -1,450 +1,0 @@
-# MSE Problem Framing 
-# Literature Review Data Analysis
-# 2/14/2019, JWC
-
-##### 
-# Meta Data and setup
-# Description: Import and analyze MSE documentation from database
-# Value: TBD
-
-# load libraries
-library(tidyverse) # upgrade to base R
-library(readxl) # read in excel files
-library(shiny) # Shiny App
-library(shinythemes) # web themes for Shiny
-library(DT) # Data Table
-library(purrr) # for loop alternative
-library(ggmap) # map plotting
-library(mapdata) # basemap generation
-
-##### 
-# Script
-
-# Get map background for plotting the map
-world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
-
-
-#####
-###-- Data Processing --###
-
-# Obtain data tables
-study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
-mgmt<-read_xlsx("DB files - Excel/tblStudyManagementTools.xlsx")
-mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
-obj<-read_xlsx("DB files - Excel/tblStudyObjectives.xlsx")
-obj$fkStudyID<-parse_integer(obj$fkStudyID)
-fields<-read_xlsx("DB files - Excel/tblStudyFields.xlsx")
-
-#####
-# Edit tables for joining
-study<-study %>% 
-  unite(Citation,c(Authors,YearPub),sep=" ",remove=F)
-
-study.join<-study %>% 
-  select(ID,Citation)
-
-mgmt.join<-mgmt %>% 
-  group_by(fkStudyID) %>%
-  summarise(ManagementTool = toString(sort(unique(MPManagementTool))),
-            AlternativesEvaluated = toString(sort(unique(MPAlternativesEvaluated))))
-
-obj.join<-obj %>% 
-  group_by(fkStudyID) %>%
-  summarise(ObjectiveCategories = toString(sort(unique(ObjCategory))))
-
-# Join tables
-data<-full_join(study,mgmt.join,by=c("ID"="fkStudyID"))
-data<-full_join(data,obj.join,by=c("ID"="fkStudyID")) %>% 
-  select(-c("ID"))
-obj.data<-right_join(study.join,obj,by=c("ID"="fkStudyID"))%>% 
-  select(-c("ID","ID.y"))
-
-# Move comment column to the end
-data<-data %>% 
-  select(-'Comments', 'Comments')
-
-# Get columns whose width needs editing
[…]

### 2020-03-24 R/Analysis.R

@@ -1,0 +1,453 @@
+# MSE Problem Framing 
+# Literature Review Data Analysis
+# 1/4/2018, JWC
+
+##### 
+# Meta Data and setup
+# Description: Import and analyze MSE documentation from database
+# Value: TBD
+
+# load libraries
+library(tidyverse) # upgrade to base R
+library(RODBC) # connect to databases
+library(shiny) # Shiny App
+library(shinythemes) # web themes for Shiny
+library(DT) # Data Table
+library(purrr) # for loop alternative
+library(ggmap) # map plotting
+library(mapdata) # basemap generation
+
+##### 
+# Script
+
+#####
+###-- Data Processing --###
+wd<-getwd()
+# Establish connection to Access database
+channel<-odbcDriverConnect(paste0("Driver={Microsoft Access Driver (*.mdb, *.accdb)};
+                           DBQ=",wd,"/DB files - Access/MSE Problem Framing_be.accdb"))
+
+# Obtain data tables
+study<-as_tibble(sqlFetch(channel,sqtable = "tblStudy"))
+mgmt<-as_tibble(sqlFetch(channel,sqtable = "tblStudyManagementTools"))
+obj<-as_tibble(sqlFetch(channel,sqtable = "tblStudyObjectives"))
+fields<-as_tibble(sqlFetch(channel,sqtable = "tblStudyFields"))
+odbcCloseAll()
+
+#####
+# Edit tables for joining
+study<-study %>% 
+  unite(Citation,c(Authors,YearPub),sep=" ",remove=F)
+
+study.join<-study %>% 
+  select(ID,Citation)
+
+mgmt.join<-mgmt %>%
+  group_by(fkStudyID) %>%
+  summarise(ManagementTool = toString(sort(unique(MPManagementTool))),
+            AlternativesEvaluated = toString(sort(unique(MPAlternativesEvaluated))))
+
+obj.join<-obj %>% 
+  group_by(fkStudyID) %>%
+  summarise(ObjectiveCategories = toString(sort(unique(ObjCategory))))
+
+# Join tables
+data<-full_join(study,mgmt.join,by=c("ID"="fkStudyID"))
+data<-full_join(data,obj.join,by=c("ID"="fkStudyID")) %>% 
+  select(-c("ID"))
+obj.data<-right_join(study.join,obj,by=c("ID"="fkStudyID"))%>% 
+  select(-c("ID","ID.y"))
+
+# Move comment column to the end
+data<-data %>% 
+  select(-'Comments', 'Comments')
+
[…]

### 2020-03-24 R/AnalysisExcel.R

@@ -1,0 +1,450 @@
+# MSE Problem Framing 
+# Literature Review Data Analysis
+# 2/14/2019, JWC
+
+##### 
+# Meta Data and setup
+# Description: Import and analyze MSE documentation from database
+# Value: TBD
+
+# load libraries
+library(tidyverse) # upgrade to base R
+library(readxl) # read in excel files
+library(shiny) # Shiny App
+library(shinythemes) # web themes for Shiny
+library(DT) # Data Table
+library(purrr) # for loop alternative
+library(ggmap) # map plotting
+library(mapdata) # basemap generation
+
+##### 
+# Script
+
+# Get map background for plotting the map
+world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
+
+
+#####
+###-- Data Processing --###
+
+# Obtain data tables
+study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
+mgmt<-read_xlsx("DB files - Excel/tblStudyManagementTools.xlsx")
+mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
+obj<-read_xlsx("DB files - Excel/tblStudyObjectives.xlsx")
+obj$fkStudyID<-parse_integer(obj$fkStudyID)
+fields<-read_xlsx("DB files - Excel/tblStudyFields.xlsx")
+
+#####
+# Edit tables for joining
+study<-study %>% 
+  unite(Citation,c(Authors,YearPub),sep=" ",remove=F)
+
+study.join<-study %>% 
+  select(ID,Citation)
+
+mgmt.join<-mgmt %>% 
+  group_by(fkStudyID) %>%
+  summarise(ManagementTool = toString(sort(unique(MPManagementTool))),
+            AlternativesEvaluated = toString(sort(unique(MPAlternativesEvaluated))))
+
+obj.join<-obj %>% 
+  group_by(fkStudyID) %>%
+  summarise(ObjectiveCategories = toString(sort(unique(ObjCategory))))
+
+# Join tables
+data<-full_join(study,mgmt.join,by=c("ID"="fkStudyID"))
+data<-full_join(data,obj.join,by=c("ID"="fkStudyID")) %>% 
+  select(-c("ID"))
+obj.data<-right_join(study.join,obj,by=c("ID"="fkStudyID"))%>% 
+  select(-c("ID","ID.y"))
+
+# Move comment column to the end
+data<-data %>% 
+  select(-'Comments', 'Comments')
+
+# Get columns whose width needs editing
[…]

### 2020-03-24 R/app.R

@@ -1,0 +1,602 @@
+# MSE Problem Framing 
+# Literature Review Data Analysis
+# 3/8/2019, JWC
+
+##### Meta Data and setup #####
+# Description: Import and analyze MSE documentation from database
+# Value: TBD
+
+# load libraries
+library(tidyverse) # upgrade to base R
+library(readxl) # read in excel files
+library(shiny) # Shiny App
+library(shinythemes) # web themes for Shiny
+library(DT) # Data Table
+library(purrr) # for loop alternative
+library(ggmap) # map plotting
+library(mapdata) # basemap generation
+library(odbc) # database connection 
+library(RMySQL) # MySQL scripting in R
+library(DBI) # database interface
+
+#To open a connection to the database:
+# Copy into command prompt:
+## cloud_sql_proxy -instances=alpine-tracker-230222:us-east1:mse-review=tcp:3306
+
+# # Create connection function
+# getSqlConnection <- function(){
+#   cn <-
+#     dbConnect(
+#       RMySQL::MySQL(),
+#       dbname = 'MSEreview',
+#       host = '127.0.0.1',
+#       port=3306,
+#       username = 'root',
+#       password = '2u8aNDdur1aFdb8z')
+#   return(cn)
+# }
+# # Open connection
+# con <- getSqlConnection()
+# 
+# # Obtain data tables
+# study<-dbReadTable(con,"tblStudy")
+# # study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
+# mgmt<-dbReadTable(con,"tblStudyManagement")
+# # mgmt<-read_xlsx("DB files - Excel/tblStudyManagementTools.xlsx")
+# # mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
+# obj<-dbReadTable(con,"tblStudyObjectives")
+# # obj<-read_xlsx("DB files - Excel/tblStudyObjectives.xlsx")
+# # obj$fkStudyID<-parse_integer(obj$fkStudyID)
+# fields<-dbReadTable(con,"tblStudyFields")
+# # fields<-read_xlsx("DB files - Excel/tblStudyFields.xlsx")
+# 
+# # To close the connection
+# dbDisconnect(con)
+
+load("MSEreview.RData")
+
+
+# Get map background for plotting the map
+world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
+
+
[…]

### 2020-03-24 app.R

@@ -1,602 +1,0 @@
-# MSE Problem Framing 
-# Literature Review Data Analysis
-# 3/8/2019, JWC
-
-##### Meta Data and setup #####
-# Description: Import and analyze MSE documentation from database
-# Value: TBD
-
-# load libraries
-library(tidyverse) # upgrade to base R
-library(readxl) # read in excel files
-library(shiny) # Shiny App
-library(shinythemes) # web themes for Shiny
-library(DT) # Data Table
-library(purrr) # for loop alternative
-library(ggmap) # map plotting
-library(mapdata) # basemap generation
-library(odbc) # database connection 
-library(RMySQL) # MySQL scripting in R
-library(DBI) # database interface
-
-#To open a connection to the database:
-# Copy into command prompt:
-## cloud_sql_proxy -instances=alpine-tracker-230222:us-east1:mse-review=tcp:3306
-
-# # Create connection function
-# getSqlConnection <- function(){
-#   cn <-
-#     dbConnect(
-#       RMySQL::MySQL(),
-#       dbname = 'MSEreview',
-#       host = '127.0.0.1',
-#       port=3306,
-#       username = 'root',
-#       password = '2u8aNDdur1aFdb8z')
-#   return(cn)
-# }
-# # Open connection
-# con <- getSqlConnection()
-# 
-# # Obtain data tables
-# study<-dbReadTable(con,"tblStudy")
-# # study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
-# mgmt<-dbReadTable(con,"tblStudyManagement")
-# # mgmt<-read_xlsx("DB files - Excel/tblStudyManagementTools.xlsx")
-# # mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
-# obj<-dbReadTable(con,"tblStudyObjectives")
-# # obj<-read_xlsx("DB files - Excel/tblStudyObjectives.xlsx")
-# # obj$fkStudyID<-parse_integer(obj$fkStudyID)
-# fields<-dbReadTable(con,"tblStudyFields")
-# # fields<-read_xlsx("DB files - Excel/tblStudyFields.xlsx")
-# 
-# # To close the connection
-# dbDisconnect(con)
-
-load("MSEreview.RData")
-
-
-# Get map background for plotting the map
-world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
-
-
[…]

### 2020-03-24 R/app.R
Add here() package to aid new file structure.

@@ -18,6 +18,7 @@ library(mapdata) # basemap generation
 library(odbc) # database connection 
 library(RMySQL) # MySQL scripting in R
 library(DBI) # database interface
+library(here) # file directory assistance
 
 #To open a connection to the database:
 # Copy into command prompt:
@@ -53,7 +54,7 @@ library(DBI) # database interface
 # # To close the connection
 # dbDisconnect(con)
 
-load("MSEreview.RData")
+load(here("data/MSEreview.RData"))
 
 
 # Get map background for plotting the map
@@ -180,7 +181,7 @@ ui <- fluidPage(
       alternative management strategies and to assess the consequences of
       uncertainty for achieving management goals' (Punt et al. 2014). Thus,
       MSE is a compelling tool to assess fisheries management procedures.  For
-      example, in our changing oceans and increasing climate change impacts MSE
+      example, in our changing oceans and with increasing climate change impacts MSE
       can evaluate climate-ready options for fisheries management decisions."),
       p("Adaptive management arose to address uncertainties and accelerate 
       progress towards meeting management objectives. In the case of fisheries

### 2020-03-26 R/MSE manuscript.Rmd
Add use of here() package; correct number of climate change MSEs

@@
 library(kableExtra) # extra formating of the tables
+library(here) # file directory assistance
 
-load("MSEreview.RData")
+load(here("data/MSEreview.RData"))
 
@@
 We reviewed a random sample of 30 articles that document a MSE, removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
-Of the 264 articles, 11 included climate change as a driver of fishery status, and after removing other articles 154 articles document a MSE.
+After removing other articles from the original set of 264 articles, 154 articles document a MSE and 13 of these included climate change as a driver of fishery status.
+To give as full a picture of the climate change MSE to date as possible we have also reviewed the additional climate change MSE
 
@@
 | Random Sample | 30 |
-| Climate Change | 11 |  
+| Climate Change | 13 |  
 *Estimated by extrapolating from the assessed MSEs to the full set for now, the submission will contain an final count.

### 2020-03-26 R/MSE manuscript.Rmd
update text of the abstract and introduction

@@
 bibliography: MSEreview.bib
-abstract: The aim and vision for fisheries management that comes from management strategy evaluation and adaptive management is a nobel and aspirational vision.
+abstract: The aim and vision for fisheries management that comes from management strategy evaluation and adaptive management is a noble and aspirational vision.
 We review MSE documentation in the published literature to appraise the achievement of this vision, finding that reality doesn't fully reflect this vision.
-The steps taken to complete the MSE process often are not documented.
-Scientists often fulled the role of fishery stakeholders, seemingly selecting the objectives and alternatives to evaluate in most MSEs.
-Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.
+The steps taken to complete the management strategy evaluation process often are not documented.
+Scientists often filled the role of fishery stakeholders, seemingly selecting the objectives and alternatives to evaluate in most management strategy evaluations.
+Facilitators and decision analysts -- participants focused on the decision making process rather than making the decision -- were rarely participants in MSE processes.
 Conservation, Economic, and Yield focused objectives were commonly considered in MSE processes, while broader social objectives were rarely considered.
-The picture painted by MSE documentation is primarily one of scientist driven simulation studies rather than management driven decision processes, with limited process documentation.
+The picture painted by MSE documentation is primarily one of scientist driven simulation studies rather than management driven decision processes, with limited process documentation occurring.
+We hope providing a repository for reviews of previous management strategy evaluatons will aid and speed learning among practitioners of management strategy evaluation.
 ---
@@
 For example, management strategy evaluation (MSE) is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@Punt].
-The positive aspects of MSE that have been listed include that it is “A tool that scientists and managers can use to simulate the workings of a fisheries system and allow them to test whether potential harvest strategies—or management procedures— can achieve pre-agreed management objectives.”[@Pew], that it “can be used to identify a ‘best’ management strategy among a set of candidate strategies" [@PuntLab] and “MSE has the advantage of being able to reveal the trade-offs among a range of possible management decisions, and should assist our Commissioners in determining whether the objectives should be weighted differently.
+The positive aspects of MSE that have been listed include that it is “A tool that scientists and managers can use to simulate the workings of a fisheries system and allow them to test whether potential harvest strategies — or management procedures — can achieve pre-agreed management objectives.”[@Pew], that it “can be used to identify a ‘best’ management strategy among a set of candidate strategies" [@PuntLab] and “MSE has the advantage of being able to reveal the trade-offs among a range of possible management decisions, and should assist our Commissioners in determining whether the objectives should be weighted differently.
 Specifically, to provide the information on which to base a rational decision, given objectives, preferences, and attitudes to risk.”  [@IPHC].
@@
 
-In terms of the MSE process, it is described as being "at the interface between science and policy.” [@PuntLab] and being an “Undertaking [that] requires scientists, managers, and stakeholders to be involved throughout the process." [@IPHC] The International Pacific Halibut Commission (2017) also states, "While the scientists do the modelling, managers must offer extensive input.
+Are the results of MSE living up to their descriptions?
+Are MSEs evaluating trade-offs and the consequences of uncertainty?
+Do they test potential harvest strategies?
+Were the management objectives used to evaluate management proceedures established and agreed to?
+Was a 'best' management strategy identified?
+Was consideration given to the weights on objectives, preferences, and attitudes to risk?
+
+In the descriptions of the MSE process, the process itself is described as being "at the interface between science and policy.” [@PuntLab] and being an “Undertaking [that] requires scientists, managers, and stakeholders to be involved throughout the process." [@IPHC] The International Pacific Halibut Commission (2017) also states, "While the scientists do the modelling, managers must offer extensive input.
 Because of the many steps and the iterative process, communication among parties is critical for achieving buy-in on the results of the management strategy evaluation.”
+
+ARe MSE processes being conducted in the manner described?
+Do MSE processes consider and account for the interface between science and policy?
+Are scientists, managers and stakeholders involved and communicating through and iterative process that creates buy-in?
 
@@
 ### Objectives 
-Our overall objectives is to determine the extent that MSE documentation can support learning within the MSE practitioner community?
+Our overall objectives is to determine the extent that MSE achieves the aspirational aims outlined above, and whether MSE documentation supports learning within the MSE practitioner community.
 Given that scientific publication is the primary means of communicating the results of MSE, we review the scientific literature to assess the extent to which the process and aims of MSE, and adaptive management are being achieved and supported by the scientific literature.

### 2020-03-26 R/MSE manuscript.Rmd
Edit through objectives

@@
 Do they test potential harvest strategies?
-Were the management objectives used to evaluate management proceedures established and agreed to?
+Were the management objectives used to evaluate management procedures established and agreed to?
 Was a 'best' management strategy identified?
@@
 
-ARe MSE processes being conducted in the manner described?
+Are MSE processes being conducted in the manner described?
 Do MSE processes consider and account for the interface between science and policy?

### 2020-03-27 R/MSE manuscript.Rmd
Edit objective text

@@
 ### Objectives 
-Our overall objectives is to determine the extent that MSE achieves the aspirational aims outlined above, and whether MSE documentation supports learning within the MSE practitioner community.
+Our overall objective is to determine the extent that MSE achieves the aspirational aims outlined above, and whether MSE documentation supports learning within the MSE practitioner community.
 Given that scientific publication is the primary means of communicating the results of MSE, we review the scientific literature to assess the extent to which the process and aims of MSE, and adaptive management are being achieved and supported by the scientific literature.
@@
 * Identify the objectives and alternatives considered in MSE
-* Determine the prevalence of explicit documentation and use of decision making steps in the MSE process, and 
+* Determine the prevalence of a) explicit documentation and b) use of decision making steps in the MSE process, and 
 * Identify participants in MSE processes.
 
-We also conduct a focused analysis of MSE processes that address the influence of climate change on fisheries management, highlighting unique aspects of MSE processes that model climate change as a driver of system status.
+We also conduct a focused analysis of MSE processes that address the influence of climate change on fisheries management, highlighting aspects of MSE processes that model climate change as a driver of system status.
 

### 2020-04-02 R/MSE manuscript.Rmd
Edit objectives and methods for climate change citations and counts

@@
 
+An incresingly recognized challenge in fisheries management is responding to changes in the productivity of fisheries stocks as a result of climate change [@busch_climate_2016, @brander_global_2007].
 We also conduct a focused analysis of MSE processes that address the influence of climate change on fisheries management, highlighting aspects of MSE processes that model climate change as a driver of system status.
@@
 This search returned 264 results.
-We reviewed a random sample of 30 articles that document a MSE, removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
-After removing other articles from the original set of 264 articles, 154 articles document a MSE and 13 of these included climate change as a driver of fishery status.
-To give as full a picture of the climate change MSE to date as possible we have also reviewed the additional climate change MSE
+We reviewed a random sample of 30 articles that document a MSE (supplement 2), removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
+After removing other articles from the original set of 264 articles, 154 articles document a MSE.
+We reviewed 16 climate change MSE articles (supplement 2).
+Twelve of these articles came from our original search.
+To give as full a picture of the climate change MSE articles to date as possible we also reviewed the 4 additional climate change MSE articles published in 2019.
 
@@
 | Random Sample | 30 |
-| Climate Change | 13 |  
-*Estimated by extrapolating from the assessed MSEs to the full set for now, the submission will contain an final count.
+| Climate Change | 12 |
 

### 2020-04-06 R/MSE manuscript.Rmd
Edit search methods

@@
 ## Finding and Sampling the MSE Literature 
-We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019.
+We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019 (Table 1).
 This search returned 264 results.
@@
 We reviewed 16 climate change MSE articles (supplement 2).
-Twelve of these articles came from our original search.
-To give as full a picture of the climate change MSE articles to date as possible we also reviewed the 4 additional climate change MSE articles published in 2019.
+Twelve of these articles came from our original January 8th, 2019 Web of Science search.
+To give as full a picture of the climate change MSE articles to date as possible we also reviewed the 4 climate change MSE articles published in 2019.
 
+Table 1. Januaray 8, 2019 Web of Science "management strategy evaluation" search results
 | MSE type | Count |
@@
 | Random Sample | 30 |
-| Climate Change | 12 |
+| Climate Change | 12* |  
+* We also reviewed 4 additional climate change MSE articles published in 2019 for a total of 16.
 

### 2020-04-06 R/MSE manuscript.Rmd
Make first table reference dynamic and use kable to produce table

@@
 ## Finding and Sampling the MSE Literature 
-We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019 (Table 1).
+We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019 (Table \@ref(tab:MSE count)).
 This search returned 264 results.
@@
 To give as full a picture of the climate change MSE articles to date as possible we also reviewed the 4 climate change MSE articles published in 2019.
+
+```{r MSE count}
+MSE_Type<-c("Published","Random Sample","Climate Change")
+Count<-c(154,30,12)
+Table1.data<-as.tibble(cbind(MSE_Type,Count))
+  Table1.data %>%
+  kable(caption = "Januaray 8, 2019 Web of Science 'management strategy evaluation' search results") %>% 
+  kable_styling()
+```
 

### 2020-04-06 R/MSE manuscript.Rmd
Covert markdown table of MSE counts to kable table produced with R code.

--- 2020-04-06 R/MSE manuscript.Rmd
@@
 ## Finding and Sampling the MSE Literature 
-We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019 (Table 1).
+We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019 (Table \@ref(tab:MSEcount)).
 This search returned 264 results.
@@
 
-Table 1. Januaray 8, 2019 Web of Science "management strategy evaluation" search results
-| MSE type | Count |
-|:--:|:--:|
-| Published | 154 |
-| Random Sample | 30 |
-| Climate Change | 12* |  
-* We also reviewed 4 additional climate change MSE articles published in 2019 for a total of 16.
+```{r MSEcount}
+MSE_Type<-c("Published","Random Sample","Climate Change")
+Count<-c(154,30,"12*")
+Table1.data<-as_tibble(cbind(MSE_Type,Count))
+  Table1.data %>%
+    rename("MSE Type"=MSE_Type) %>% 
+  kable(caption = "Januaray 8, 2019 Web of Science 'management strategy evaluation' search results") %>% 
+  kable_styling() %>% 
+    add_footnote(c("We also reviewed 4 additional climate change MSE articles published in 2019 for a total of 16."), notation = "symbol")
+```
 
@@
 ```{r, SDMfigure, fig.cap='Review framework - structured decision making process'}
-knitr::include_graphics('figs/SDMProcessFramework.png')
+knitr::include_graphics(here('figs/SDMProcessFramework.png'))
 ```

### 2020-04-06 R/MSE manuscript.Rmd
Edit methods and results text until the start of the climate change insert.

@@
 summary.data_pub<-data_pub %>%
-    select(summary.col)
+    select(all_of(summary.col))
 summary.data_CC<-data_CC %>%
-  select(summary.col)
+  select(all_of(summary.col))
 ```
@@
 freq.data_pub<-data_pub %>%
-    select(freq.col) %>%
+    select(all_of(freq.col)) %>%
     rename("Process"="ProcessExplicit",
@@
 freq.data_CC<-data_CC %>%
-  select(freq.col) %>%
+  select(all_of(freq.col)) %>%
   rename("Process"="ProcessExplicit",
@@
 freq2.data_pub<-data_pub %>%
-    select(freq2.col) %>%
+    select(all_of(freq2.col)) %>%
     rename("Roles"="RolesExplicit",
@@
 part.data_pub<-data_pub %>%
-    select(part.col) %>%
+    select(all_of(part.col)) %>%
     rename("Process"="Leader",
@@
 part.data_CC<-data_CC %>%
-  select(part.col) %>%
+  select(all_of(part.col)) %>%
   rename("Process"="Leader",
@@
 obj.data_pub<-obj.data_pub %>%
-  select(obj.col) %>%
+  select(all_of(obj.col)) %>%
   purrr::map(table) %>%
@@
 obj.data_CC<-obj.data_CC %>%
-  select(obj.col) %>%
+  select(all_of(obj.col)) %>%
   purrr::map(table) %>%
@@
 map.data<-rbind(data_pub,data_CC) %>%
-  select(map.col) %>%
+  select(all_of(map.col)) %>%
   mutate(Drivers=str_extract(Drivers, "Climate Change")) %>%
@@
 We report the drivers included as the primary factors influencing the predictions produced by the MSE model, listing each driver a MSE included.
-We also report how the consequence predictions were made, with possible values of simulation modeling, dynamic programming, expert elicitation, mental models, and unknown.
+We also report how the consequence predictions were made, with possible values of simulation modeling, dynamic programming, expert elicitation, and mental models, or unknown when it is unclear what methodology was used to predict the status of the system for management evaluation.
 
@@
 We record each objective considered, and report the type and category of objectives included in each MSE.
-Natural resource management problems typically include four types of objectives: conservation, resource use, i.e., yield in the fisheries context, economic, and social objectives [@runge2013].
+Natural resource management problems typically include four types of objectives: conservation, resource use, i.e., yield in the fisheries context, as well as economic, and social objectives [@runge2013].
 We sort objectives into four types: strategic objectives - which are concerned with the overall mission of the system in which the decision evaluated by the MSE falls, process objectives - which are concerned with how the management decision addressed by the MSE would be made, fundamental objectives - which address the goals of the management decision, and means objectives - which address means to achieving the fundamental objectives of the management decision.
@@
 In the stages group we address how the MSE stages were completed.
-We report whether the stages of the structured decision making framework were explicitly completed and documented.
+We report whether the stages of the structured decision making framework, i.e., problem, objectives, alternatives, tradeoffs, were explicitly completed and the methods used to complete them were documented.
 We also report whether the MSE results were adopted in subsequent management.
@@
 In the participant group we address how participation in the MSE process was structured and who participated.
-We report who lead the MSE, who participated in it, and if the roles of participants were established and documented.
+We report who lead the MSE process, who participated in it, and if the roles of participants were established and documented.
 In many cases the process used to elicit objectives and alternatives was not explicit.
@@
 ### What questions and management objectives do MSEs address? 
-We found that uncertainty was the main driver of the system state included in management strategy evaluations, with uncertainty included in 21 of the 30 evaluations we reviewed (Table \@ref(tab:drivers)).
+We found that uncertainty was the main driver of the system state included in management strategy evaluations, with uncertainty included in 21 of the 30 evaluations we reviewed in our random sample (Table \@ref(tab:drivers)).
 In addition to uncertainty about the system model itself, monitoring methodology was the most prevalent driver in MSE models, but a variety of drivers were considered across the random sample of MSEs.
@@
 <br>
-Conservation objectives were present in nearly all of the MSEs we reviewed (Table \@ref(tab:objectiveCategories)).
-Yield and economic objectives were present in about three quarters of the MSEs, while social objectives were only included in 2 of the MSEs.
+Conservation objectives were present in nearly all of the MSEs we sampled (Table \@ref(tab:objectiveCategories)).
+Yield and economic objectives were present in about three quarters of the MSEs, while social objectives were only included in 2 of the sampled MSEs.
 
@@
 <br>
-No strategic or process objectives were documented in the MSEs we reviewed.  There were roughly the same number of fundamental and means included in the MSEs (Table \@ref(tab:objectiveTypes)).
+No strategic or process objectives were documented in the MSEs.  There were roughly the same number of fundamental and means included in the MSEs (Table \@ref(tab:objectiveTypes)).
 ```{r, objectiveTypes}
@@
 The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Table \@ref(tab:explicit)).
-The consequence prediction stage -- which is the primary focus of MSE publication -- was always addressed.
-Apart from the consequence stage, the most documented stage was the alternative stage with 6 MSEs documenting this stage of the process.
+Apart from the consequence stage, which is the primary focus of MSE publication and therefore was always addressed, the most documented stage was the alternative stage with 6 MSEs documenting this stage of the process.
+The MSE process, problem, and objectives methodology were each documented in 5 of the MSEs.
 Three of the 30 MSEs explicitly documented a trade-off analysis and 2 of the 30 documented a management decision.
@@
 The primary participants and leaders of the MSE processes were scientists, while some MSE process were led by governments and management agencies (Figure \@ref(fig:participants)).
-Scientist led 73 percent of the MSE processes, with another 23 percent lacking documentation that made the leader of the process clear.
+Scientists were leaders, or the sole participants and therefore seemingly the leaders, in 73 percent of the MSE processes, while 23 percent of process lacked enough documentation of participants and leadership such that we were unable to identify a presumed leader of the process.
 
@@
 In most cases the source of the objectives and alternatives was not explicit.
-In those cases scientists selected these components most frequently, while in some cases the source was unknown.
+In those subjective cases scientists seemingly selected these components most frequently, while in some cases the source was unknown.
 

### 2020-04-07 R/AnalysisExcel.R
Fix file paths for folder reorganization

@@ -28,12 +28,12 @@ world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a
 ###-- Data Processing --###
 
 # Obtain data tables
-study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
-mgmt<-read_xlsx("DB files - Excel/tblStudyManagementTools.xlsx")
+study<-read_xlsx("data/DB files - Excel/tblStudy.xlsx")
+mgmt<-read_xlsx("data/DB files - Excel/tblStudyManagementTools.xlsx")
 mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
-obj<-read_xlsx("DB files - Excel/tblStudyObjectives.xlsx")
+obj<-read_xlsx("data/DB files - Excel/tblStudyObjectives.xlsx")
 obj$fkStudyID<-parse_integer(obj$fkStudyID)
-fields<-read_xlsx("DB files - Excel/tblStudyFields.xlsx")
+fields<-read_xlsx("data/DB files - Excel/tblStudyFields.xlsx")
 
 #####
 # Edit tables for joining

### 2020-04-07 R/MSE manuscript.Rmd
Fix names of fields taken from Rdata; edit climate change text box

@@
   group_by(fkStudyID) %>%
-  summarise(ManagementType = toString(sort(unique(ManagementType))),
-            AlternativesEvaluated = toString(sort(unique(AlternativesEvaluated))))
+  summarise(ManagementType = toString(sort(unique(MPManagementTool))),
+            AlternativesEvaluated = toString(sort(unique(MPAlternativesEvaluated))))
 
@@
 
-We review published MSEs including climate change as a driver against the stages in a SDM process to:  Highlight unique aspects of MSEs that model climate change as a driver to help inform fisheries management
+We review published MSEs including climate change as a driver against the stages in a SDM process to highlight unique aspects of MSEs that model climate change as a driver to help inform fisheries management
 
-We filtered the original set of MSE by searching for the term "climate change" within the set.
-We found that, as of the end of December 2018, 11 MSEs included climate change as a driver.
+We filtered the original set of MSE by searching for the term "climate change" within the set, reviewing 16 published MSEs included climate change as a driver as of December 31st, 2019.
 We reviewed the climate change MSEs using the same methodology described above.

### 2020-04-07 R/app.R
Fix names of fields taken from Rdata

@@ -72,8 +72,8 @@ study.join<-study %>%
 
 mgmt.join<-mgmt %>% 
   group_by(fkStudyID) %>%
-  summarise(ManagementType = toString(sort(unique(ManagementType))),
-            AlternativesEvaluated = toString(sort(unique(AlternativesEvaluated))))
+  summarise(ManagementType = toString(sort(unique(MPManagementTool))),
+            AlternativesEvaluated = toString(sort(unique(MPAlternativesEvaluated))))
 
 obj.join<-obj %>% 
   group_by(fkStudyID) %>%

### 2020-04-07 R/MSE manuscript.Rmd
Edit publication and CC filters to match updated database column names; update frequency column selection to match updated database column names

@@
 # Filter data by study type
-data_pub<-filter(data,IncludeInPublication==TRUE)
-data_CC<-filter(data,str_detect(Drivers,"Climate Change"))
+data_pub<-filter(data,RandomSample==TRUE)
+data_CC<-filter(data,str_detect(Drivers,"Climate Change")&UseInPublication==TRUE)
 
@@
             "ResultsAdopted")
-freq2.col<-c("RolesExplicit",
+freq2.col<-c("RoleSpecification",
              "OpenMeetings")
@@
     select(all_of(freq2.col)) %>%
-    rename("Roles"="RolesExplicit",
+    rename("Roles"="RoleSpecification",
            "Open Meetings"="OpenMeetings") %>%

### 2020-04-07 R/app.R
Update data filters to match updated database columns for IncludeInPublication and Climate Change columns

@@ -89,8 +89,8 @@ data<-data %>%
   select(-'Comments', 'Comments')
 
 # Filter data to create separate pub and climate change objects
-data_pub<-filter(data,IncludeInPublication==TRUE)
-data_CC<-filter(data,str_detect(Drivers,"Climate Change"))
+data_pub<-filter(data,RandomSample==TRUE)
+data_CC<-filter(data,str_detect(Drivers,"Climate Change")&UseInPublication==TRUE)
 
 data_pub.join<-data_pub %>%
   select(ID,Citation)
@@ -143,7 +143,7 @@ freq.col<-c("ProcessExplicit",
             "AlternativesExplicit",
             "TradeOffsExplicit",
             "DecisionExplicit",
-            "RolesExplicit",
+            "RoleSpecification",
             "OpenMeetings",
             "ResultsAdopted")
 # Get columns for participant analysis
@@ -322,15 +322,15 @@ server <- function(input, output, session) {   # code to create output using ren
   # Filter data to include only those reviewed for the publication or all MSEs
   data_reviewed<-reactive({
     if(input$data_filter=="pub"){
-      filter(data,IncludeInPublication==TRUE)
+      filter(data,RandomSample==TRUE)
     } 
     else 
       if(input$data_filter=="CC"){
-        filter(data,str_detect(Drivers,"Climate Change"))
+        filter(data,str_detect(Drivers,"Climate Change")&UseInPublication==TRUE)
       }
     else 
       if(input$data_filter=="all"){
-        filter(data,IncludeInPublication==TRUE|IncludeInPublication==FALSE)
+        filter(data,UseInPublication==TRUE|UseInPublication==FALSE)
     }
   })
   
@@ -354,7 +354,7 @@ server <- function(input, output, session) {   # code to create output using ren
            "Alternatives"="AlternativesExplicit",
            "Tradeoffs"="TradeOffsExplicit",
            "Decision"="DecisionExplicit",
-           "Roles"="RolesExplicit",
+           "Roles"="RoleSpecification",
            "Open Meetings"="OpenMeetings",
            "Adopted"="ResultsAdopted") %>%
     summarise_all(funs(sum)) %>%

### 2020-04-07 R/MSE manuscript.Rmd
Fix alternative data join; fix string split for part.date_pub

@@
 
-alt.data_pub<-left_join(data_pub.join,mgmt,by=c("ID"="fkStudyID"))%>%
-  select(-c("ID","ID.y"))
-alt.data_CC<-left_join(data_CC.join,mgmt,by=c("ID"="fkStudyID"))%>%
-  select(-c("ID","ID.y"))
+alt.data_pub<-left_join(data_pub.join,mgmt.join,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID"))
+alt.data_CC<-left_join(data_CC.join,mgmt.join,by=c("ID"="fkStudyID"))%>%
+  select(-c("ID"))
 
@@
 part.data_pub<- part.data_pub %>%
-    purrr::map(~ strsplit(as.character(.),split=",")) %>%
+    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
     purrr::map(unlist) %>%
@@
 part.data_CC<- part.data_CC %>%
-  purrr::map(~ strsplit(as.character(.),split=",")) %>%
+  purrr::map(~ strsplit(as.character(.),split=", ")) %>%
   purrr::map(unlist) %>%

### 2020-04-07 R/MSE manuscript.Rmd
Fis all strsplit by changing "," to ", "; remove uneccessary funs() from summarise_all calls; Add grey circle around points in the map; comment out part.data$order specification

@@
            "Results Adopted"="ResultsAdopted") %>%
-    summarise_all(funs(sum))%>%
+    summarise_all(sum)%>%
     gather(Explicit) %>%
@@
          "Results Adopted"="ResultsAdopted") %>%
-  summarise_all(funs(sum))%>%
+  summarise_all(sum)%>%
   gather(Explicit) %>%
@@
            "Open Meetings"="OpenMeetings") %>%
-    summarise_all(funs(sum)) %>%
+    summarise_all(sum) %>%
     gather(Explicit) %>%
@@
     select(Drivers) %>%
-    purrr::map(~ strsplit(as.character(.),split=",")) %>%
+    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
     purrr::map(unlist) %>%
@@
   select(Drivers) %>%
-  purrr::map(~ strsplit(as.character(.),split=",")) %>%
+  purrr::map(~ strsplit(as.character(.),split=", ")) %>%
   purrr::map(unlist) %>%
@@
     select(ManagementType) %>%
-    purrr::map(~ strsplit(as.character(.),split=",")) %>%
+    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
     purrr::map(unlist) %>%
@@
   mutate(order = row_number())
-part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,
-                   16,17,18,20,21,22,23,24,25,21,23,26,25,26,27,27,28,29,30,31,29,32,30,33,
-                   32,33,34,35,36,37,38,39,39)
+# part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,
+#                    20,16,17,18,20,21,22,23,24,25,21,23,26,25,26,27,27,28,29,30,31,29,32,
+#                    30,33,32,33,34,35,36,37,38,39,39)
 drive.data_pub<-mutate(drive.data_pub,Analysis="Random Sample") %>% 
@@
 MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
-  geom_point(size=2.5) + scale_color_manual(values=c("#66CCCC","#006666")) +  theme_void() +
+  geom_point(size=2.5) + geom_point(size = 2.5, colour = "grey", shape = 1) +
+  scale_color_manual(values=c("#66CCCC","#006666"))+theme_void() +
   theme(legend.position = c(0.15, 0.25))

### 2020-04-07 R/app.R
remove unneccessary funs() in summarise_all; make all strsplit split arguments ", "; add grey border to map points; all all_of() to select

@@ -357,7 +357,7 @@ server <- function(input, output, session) {   # code to create output using ren
            "Roles"="RoleSpecification",
            "Open Meetings"="OpenMeetings",
            "Adopted"="ResultsAdopted") %>%
-    summarise_all(funs(sum)) %>%
+    summarise_all(sum) %>%
     gather(Explicit) %>%
     mutate(Percent=value/n_mse()*100) %>%
     mutate(Explicit=factor(Explicit,levels=
@@ -376,7 +376,7 @@ server <- function(input, output, session) {   # code to create output using ren
              "Sub Alternatives"="ProcedureElicitation_Sub")
   })
   part.data_reviewed2<- reactive({part.data_reviewed() %>%
-    purrr::map(~ strsplit(as.character(.),split=",")) %>%
+    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
     purrr::map(unlist) %>%
     purrr::map(table)
   })
@@ -417,7 +417,7 @@ server <- function(input, output, session) {   # code to create output using ren
   # What drivers are considered
   drive.data<-reactive({data_reviewed() %>%
     select(Drivers) %>%
-    purrr::map(~ strsplit(as.character(.),split=",")) %>%
+    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
     purrr::map(unlist) %>%
     purrr::map(table) %>%
     plyr::ldply(data.frame) %>%
@@ -475,7 +475,7 @@ server <- function(input, output, session) {   # code to create output using ren
     select(ManagementType) %>%
     purrr::map(~ strsplit(as.character(.),split=", ")) %>%
     purrr::map(unlist) %>%
-    purrr::map(~ strsplit(as.character(.),split=",")) %>% 
+    purrr::map(~ strsplit(as.character(.),split=", ")) %>% 
     purrr::map(unlist) %>%
     purrr::map(table) %>%
     plyr::ldply(data.frame) %>%
@@ -490,7 +490,7 @@ server <- function(input, output, session) {   # code to create output using ren
 
   # Where MSEs have occured
   map.data<-reactive({data_reviewed() %>%
-    select(map.col)
+    select(all_of(map.col))
   })
 
   # Tab 1 - Filtering
@@ -499,7 +499,7 @@ server <- function(input, output, session) {   # code to create output using ren
   output$mse.map <- renderPlot({
     # plot MSEs on map
[…]

### 2020-04-08 R/app.R
Fix all_of() in select arguments; fix spelling error; fix climate change count

@@ -220,7 +220,7 @@ ui <- fluidPage(
          "Show results from:",
          choices = c(
            "30 articles reviewed for Cummings et. al. Publication, "="pub",
-           "MSE included climate change as a driver"="CC",
+           "MSE including climate change as a driver"="CC",
            "All MSEs (entered in the database to date)"="all"),
          width='400px'),
         textOutput("radio")
@@ -336,18 +336,18 @@ server <- function(input, output, session) {   # code to create output using ren
   
   # Select study summary data_reviewed
   summary.data<-reactive({data_reviewed() %>%
-    select(summary.col)
+    select(all_of(summary.col))
   })
 
   # Select study problem and driver data_reviewed
   prob.data<-reactive({data_reviewed() %>%
-    select(prob.col)
+    select(all_of(prob.col))
   })
   
   # Frequency of method
   n_mse<-reactive({nrow(data_reviewed())})
   freq.data<-reactive({data_reviewed() %>%
-    select(freq.col) %>%
+    select(all_of(freq.col)) %>%
     rename("Process"="ProcessExplicit",
            "Problem"="ProblemDefinitionExplicit",
            "Objectives"="ObjectivesExplicit",
@@ -368,7 +368,7 @@ server <- function(input, output, session) {   # code to create output using ren
 
   # Who participates
   part.data_reviewed<-reactive({data_reviewed() %>%
-    select(part.col) %>%
+    select(all_of(part.col)) %>%
       rename("Process"="Leader",
              "Doc Objectives"="ObjElicitationSource_Exp",
              "Doc Alternatives"="ProcedureElicitation_Exp",
@@ -442,7 +442,7 @@ server <- function(input, output, session) {   # code to create output using ren
 
   # How were objectives defined
   obj.data2<-reactive({obj.data %>%
-    select(obj.col) %>%
+    select(all_of(obj.col)) %>%
     purrr::map(table)
   })
   
@@ -463,7 +463,7 @@ server <- function(input, output, session) {   # code to create output using ren
 
   obj.data_table2<-reactive({
     d<-obj.data %>%
-      select(obj.col) %>%
+      select(all_of(obj.col)) %>%
       group_by(ObjType,ObjCategory,ObjDirection,ObjScale) %>%
       summarize(n())
     colnames(d)<-c("Type","Category","Direction","Scale","Number")
@@ -596,7 +596,7 @@ server <- function(input, output, session) {   # code to create output using ren
[…]

### 2020-04-09 R/app.R
rearrange files to better enable publishing the shiny app

@@ -1,603 +1,0 @@
-# MSE Problem Framing 
-# Literature Review Data Analysis
-# 3/8/2019, JWC
-
-##### Meta Data and setup #####
-# Description: Import and analyze MSE documentation from database
-# Value: TBD
-
-# load libraries
-library(tidyverse) # upgrade to base R
-library(readxl) # read in excel files
-library(shiny) # Shiny App
-library(shinythemes) # web themes for Shiny
-library(DT) # Data Table
-library(purrr) # for loop alternative
-library(ggmap) # map plotting
-library(mapdata) # basemap generation
-library(odbc) # database connection 
-library(RMySQL) # MySQL scripting in R
-library(DBI) # database interface
-library(here) # file directory assistance
-
-#To open a connection to the database:
-# Copy into command prompt:
-## cloud_sql_proxy -instances=alpine-tracker-230222:us-east1:mse-review=tcp:3306
-
-# # Create connection function
-# getSqlConnection <- function(){
-#   cn <-
-#     dbConnect(
-#       RMySQL::MySQL(),
-#       dbname = 'MSEreview',
-#       host = '127.0.0.1',
-#       port=3306,
-#       username = 'root',
-#       password = '2u8aNDdur1aFdb8z')
-#   return(cn)
-# }
-# # Open connection
-# con <- getSqlConnection()
-# 
-# # Obtain data tables
-# study<-dbReadTable(con,"tblStudy")
-# # study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
-# mgmt<-dbReadTable(con,"tblStudyManagement")
-# # mgmt<-read_xlsx("DB files - Excel/tblStudyManagementTools.xlsx")
-# # mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
-# obj<-dbReadTable(con,"tblStudyObjectives")
-# # obj<-read_xlsx("DB files - Excel/tblStudyObjectives.xlsx")
-# # obj$fkStudyID<-parse_integer(obj$fkStudyID)
-# fields<-dbReadTable(con,"tblStudyFields")
-# # fields<-read_xlsx("DB files - Excel/tblStudyFields.xlsx")
-# 
-# # To close the connection
-# dbDisconnect(con)
-
-load(here("data/MSEreview.RData"))
-
-
-# Get map background for plotting the map
-world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
[…]

### 2020-04-09 R/app/app.R

@@ -1,0 +1,602 @@
+# MSE Problem Framing 
+# Literature Review Data Analysis
+# 4/9/2020, JWC
+
+##### Meta Data and setup #####
+# Description: Import and analyze MSE documentation from database
+# Value: TBD
+
+# load libraries
+library(tidyverse) # upgrade to base R
+library(readxl) # read in excel files
+library(shiny) # Shiny App
+library(shinythemes) # web themes for Shiny
+library(DT) # Data Table
+library(purrr) # for loop alternative
+library(ggmap) # map plotting
+library(mapdata) # basemap generation
+library(odbc) # database connection 
+library(RMySQL) # MySQL scripting in R
+library(DBI) # database interface
+library(here) # file directory assistance
+
+#To open a connection to the database:
+# Copy into command prompt:
+## cloud_sql_proxy -instances=alpine-tracker-230222:us-east1:mse-review=tcp:3306
+
+# # Create connection function
+# getSqlConnection <- function(){
+#   cn <-
+#     dbConnect(
+#       RMySQL::MySQL(),
+#       dbname = 'MSEreview',
+#       host = '127.0.0.1',
+#       port=3306,
+#       username = 'root',
+#       password = '2u8aNDdur1aFdb8z')
+#   return(cn)
+# }
+# # Open connection
+# con <- getSqlConnection()
+# 
+# # Obtain data tables
+# study<-dbReadTable(con,"tblStudy")
+# # study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
+# mgmt<-dbReadTable(con,"tblStudyManagement")
+# # mgmt<-read_xlsx("DB files - Excel/tblStudyManagementTools.xlsx")
+# # mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
+# obj<-dbReadTable(con,"tblStudyObjectives")
+# # obj<-read_xlsx("DB files - Excel/tblStudyObjectives.xlsx")
+# # obj$fkStudyID<-parse_integer(obj$fkStudyID)
+# fields<-dbReadTable(con,"tblStudyFields")
+# # fields<-read_xlsx("DB files - Excel/tblStudyFields.xlsx")
+# 
+# # To close the connection
+# dbDisconnect(con)
+
+# Load data for app (data is obtained via the analysisexcel.R script)
+load(here("MSEreview.RData"))
+
+# Get map background for plotting the map
[…]

### 2020-04-09 R/app/app.R
Remove here() from data load statement

@@ -55,7 +55,7 @@ library(here) # file directory assistance
 # dbDisconnect(con)
 
 # Load data for app (data is obtained via the analysisexcel.R script)
-load(here("MSEreview.RData"))
+load("MSEreview.RData")
 
 # Get map background for plotting the map
 world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders

### 2020-04-09 R/MSE manuscript.Rmd
Edit results text

@@
 
-An incresingly recognized challenge in fisheries management is responding to changes in the productivity of fisheries stocks as a result of climate change [@busch_climate_2016, @brander_global_2007].
+An incresingly recognized challenge in fisheries management is responding to changes in the productivity of fisheries stocks as a result of climate change [@busch_climate_2016; @brander_global_2007].
 We also conduct a focused analysis of MSE processes that address the influence of climate change on fisheries management, highlighting aspects of MSE processes that model climate change as a driver of system status.
@@
 We collected additional information about the MSEs we reviewed beyond what is reported here.
-For a full description of the information recorded in the relational database see the supplemental information (supplement), and for the full data set and results see our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
+For a full description of the information recorded in the relational database see the supplemental information (supplement 1), and for the full data set and results see our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
 
@@
 We found that uncertainty was the main driver of the system state included in management strategy evaluations, with uncertainty included in 21 of the 30 evaluations we reviewed in our random sample (Table \@ref(tab:drivers)).
-In addition to uncertainty about the system model itself, monitoring methodology was the most prevalent driver in MSE models, but a variety of drivers were considered across the random sample of MSEs.
+In addition to uncertainty about the system model itself, monitoring methodology, how and what data is collected, was the most prevalent driver in MSE models, but a variety of drivers were considered across the random sample of MSEs.
 
@@
 <br>
-No strategic or process objectives were documented in the MSEs.  There were roughly the same number of fundamental and means included in the MSEs (Table \@ref(tab:objectiveTypes)).
+No strategic or process objectives were documented in the MSEs.
+There were roughly the same number of fundamental and means objectives included in the MSEs (Table \@ref(tab:objectiveTypes)).
+An average of 6 objectives were evaluated by each management strategy evaluation.
 ```{r, objectiveTypes}
@@
 ### Who is involved in MSEs? 
-MSEs were rarely explicit about the roles participants played in MSE processes, as only one MSE documented participant roles (Table \@ref(tab:roles)).
-None of the MSEs noted whether or not public meetings were held in which the public could participate in the MSE process.
+MSEs were rarely explicit about the roles participants played in MSE processes, as only one MSE documented the roles participants played in the process (Table \@ref(tab:roles)).
+None of the MSEs noted whether or not public meetings were held in which the public was invited to participate.
 
@@
 
-The primary participants and leaders of the MSE processes were scientists, while some MSE process were led by governments and management agencies (Figure \@ref(fig:participants)).
-Scientists were leaders, or the sole participants and therefore seemingly the leaders, in 73 percent of the MSE processes, while 23 percent of process lacked enough documentation of participants and leadership such that we were unable to identify a presumed leader of the process.
+The primary participants and leaders of the MSE processes were scientists, while some MSE process were led by governments and management agencies (Figure \@ref(fig:participants) Process).
+Scientists were exlicitly the leaders -- or the sole participants and therefore seemingly the leaders -- in 73 percent of the MSE processes, while 27 percent of processes lacked enough documentation of participants and leadership such that we were unable to identify a presumed leader of the process.
 
-Scientists were participants in nearly all of the reviewed MSE processes.
+Scientists were participants in nearly all of the reviewed MSE processes (Figure \@ref(fig:participants) Participants).
 Other stakeholders in MSE processes were less frequent participants.
@@
 Decision makers -- those responsible for selecting the management plan to implement -- were participants in 13 percent of MSEs.
-Experts, a possible source of data or predictions, participated in 3 percent of the MSEs reviewed.
-Participants involved to the assist the MSE process itself, i.e., facilitators and analysts, participated in 10 and 3 percent of MSE processes respectively.
+Experts -- a possible source of data or predictions -- participated in 3 percent of the MSEs reviewed.
+Participants involved to the assist the MSE process itself, i.e., facilitators and decision analysts, participated in 10 and 3 percent of MSE processes respectively.
 
 Objectives and alternatives were elicited from a variety of participants.
-About 10 percent of the objectives and alternatives explicitly were produced by stakeholders in the fisheries, coming from management, government, or fishery sources.
-In most cases the source of the objectives and alternatives was not explicit.
-In those subjective cases scientists seemingly selected these components most frequently, while in some cases the source was unknown.
+Management, government, or fishery participants explicitly provided objectives and alternatives in about 10 percent of the MSE processes (Figure \@ref(fig:participants) Explicit).
+However, in most cases the source of the objectives and alternatives was not explicit.
+The source of objectives and alternatives was not explicit in more than 80% of the MSE processes, 25 of 30 objective sources were not explicit and 24 of 30 alternative sources.
+In those subjective cases scientists -- being the only or primary participants -- seemingly selected these components most frequently (Figure \@ref(fig:participants) Subjective).
 
-```{r participants, fig.cap='Who guided, participated in, or provided input during the specified steps of the MSE process', fig.height=6, out.width="100%"}
+```{r participants, fig.cap='Who guided (Process), participated in (Participants), or provided explicitly documented input (Explicit), or seemingly provided input (Subjective) during the specified steps of the MSE process', fig.height=6, out.width="100%"}
 Part.plot

### 2020-04-09 R/MSE-manuscript.Rmd
fix inadvertent renaming

--- 2020-04-09 R/MSE manuscript.Rmd
@@
   filter(Objective=="Type") %>%
-  select(-Analysis, -Objective) %>% 
+  select(Type, Percent, Number, 'Per MSE', -Analysis, -Objective) %>% 
   kable(caption = "Objective Types") %>% 
@@
   mutate("Percent"=round(Percent,0)) %>% 
+  select(Explicit,Percent,Number) %>% 
   kable(caption = "Prevalence of explicit documentation of stages in the MSE process") %>% 
@@
   mutate("Percent"=round(Percent,0)) %>%
+  select(Explicit,Percent,Number) %>% 
   kable(caption = "Prevalence of explicit roles for MSE participants and meetings that were open to the public during the MSE process") %>% 
@@
 Fisheries management has mostly focused on fishing impacts with ecosystem status viewed as a background constant.
-The increasing rate of climate change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
+The increasing rate of climate change and its influence on fisheries is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
 
-We review published MSEs including climate change as a driver against the stages in a SDM process to highlight unique aspects of MSEs that model climate change as a driver to help inform fisheries management
-
-We filtered the original set of MSE by searching for the term "climate change" within the set, reviewing 16 published MSEs included climate change as a driver as of December 31st, 2019.
+Here we review published MSEs the include climate change as a driver of the fisheries model using the stages in a SDM process to highlight unique aspects of these MSEs.
+We filtered the original set of MSEs published by as of December 31st, 2019 searching for the term "climate change" within the set that resulted from searching for "management strategy evaluation".
+This resulted in our reviewing 16 published MSEs that included climate change as a driver.
 We reviewed the climate change MSEs using the same methodology described above.
@@
 
-Uncertainty and monitoring methodology were also included in some of the climate change MSEs (Table \@ref(tab:driversCC)), but not as often.
-Other biological drivers were also included in some of the climate change MSEs.
-Overall climate change MSEs considered fewer drivers on average, 1.64, than the random sample of MSEs, which included 1.87.
+As with the random sample of reviewed MSEs, uncertainty was the most common driver, apart from climate change, included climate change MSEs (Table \@ref(tab:driversCC)), while other biological drivers were also included in some of the climate change MSEs.
+Overall climate change MSEs considered fewer drivers on average, 1.4, than the random sample of MSEs, which included 1.8.
 
@@
 
-More objectives were considered per MSE in the random sample of MSEs than in the climate change MSEs (Table \@ref(tab:CCPerMSE)).
-Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:CCobjectiveCategories)).
+Climate change MSEs averaged the same number of objectives per MSE.
+The categories of the objectives differed however, with more frequent inclusion of yield objectives, but less frequent consideration of economic objectives and no consideration of social objectives (Table \@ref(tab:CCobjectiveCategories)) in the climate change MSEs.
 ```{r, CCPerMSE}
-color_Pub <- which(per.MSE$Analysis=='Random Sample')
-color_CC <- which(per.MSE$Analysis=='Climate Change')
+# color_Pub <- which(per.MSE$Analysis=='Random Sample')
+# color_CC <- which(per.MSE$Analysis=='Climate Change')
+# 
+# per.MSE %>%
+#   kable(caption = "Number of objectives considered per management strategy evaluation") %>%
+#   row_spec(color_CC, color = "#66CCCC", background = "white") %>%
+#   row_spec(color_Pub, color = "#006666", background = "white")%>% 
+#   kable_styling()
+```
 
-per.MSE %>%
-  kable(caption = "Number of objectives considered per management strategy evaluation") %>%
+```{r, CCobjectiveCategories}
+objcat.data_CC<-objcat.data %>%
+  filter(`Objective Category`!="Utility") %>% 
+  add_column(order=c(1,2,5,6,7,3,4)) %>% 
+  arrange(order)
+
+color_Pub <- which(objcat.data_CC$Analysis=='Random Sample')
+color_CC <- which(objcat.data_CC$Analysis=='Climate Change')
+
+objcat.data_CC %>%
+  select(-order) %>% 
+  kable(caption = "Objective categories considered in management strategy evaluations") %>%
   row_spec(color_CC, color = "#66CCCC", background = "white") %>%
-  row_spec(color_Pub, color = "#006666", background = "white")%>% 
+  row_spec(color_Pub, color = "#006666", background = "white") %>% 
   kable_styling()
@@
 
-```{r, CCobjectiveCategories}
-color_Pub <- which(objcat.data$Analysis=='Random Sample')
-color_CC <- which(objcat.data$Analysis=='Climate Change')
-
-objcat.data %>%
-  kable(caption = "Objective categories considered in management strategy evaluations") %>%
-  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
-  row_spec(color_Pub, color = "#006666", background = "white")%>% 
-  kable_styling()
-```
-
-Climate change MSEs were more likely to explicitly document the alternative elicitation and trade-off analysis stages, but otherwise were less likely to document stages in the decision process.
+Climate change MSEs were more likely to explicitly document the trade-off analysis stage, equaly likely to document the alternatives stages, but were less likely to document the other stages in the decision process.
 ```{r, CCstages}
 freq.data_CC %>%
-  select(-Analysis) %>% 
   mutate("Percent"=round(Percent,0)) %>% 
+  select(Explicit, Percent, Number, -Analysis) %>% 
   kable(caption = "Prevalence of explicit documentation of stages in climate change MSE processes") %>%
@@
 
-Including climate change in a MSE appears to reduce the complexity of other stages in the MSE process.
-On average climate change MSEs consider fewer drivers, consider fewer and less diverse objectives, and explicitly document fewer stages of the MSE process.
-If the MSE process is time or effort limited, the inclusion of climate change may utlizing time and effort that in other case was invested elsewhere in the MSE process.
+Including climate change in a MSE appears to reduce the complexity of other stages in the MSE process, in particular those stages that utilize input from additional particapants.
+On average climate change MSEs consider fewer drivers, consider less diverse objectives, and explicitly document fewer stages of the MSE process.
+If the MSE process is time, effort, or participant limited the inclusion of climate change may be utlizing time, effort, or outside participants to focus on climate change rather than other aspects of the MSE process.
 
@@
 Management strategy evaluations typically do assess the consequences of uncertainty for achieving management goals, but other factors influencing the consequences of management strategies get more sporadic treatment.
-Data limitation is a frequent concern in fisheries management.
-Next to accounting for uncertainty, accessing the influence of monitoring methodology was the most prevalent driver included in MSEs.
-Further inclusion of this driver, and additional use of value of information analyses could expand the use of MSE as a tool to design cost-effective fisheries monitoring plans.
+Data limitation is a frequent concern in fisheries management, and contributes to the degree of uncertainty present in fisheries management.
+Next to accounting for uncertainty, accessing the influence of data quality and availablity, i.e.,  monitoring methodology was the most prevalent driver included in MSEs.
+This likely points to a desire for the design and use of cost-effective fisheries monitoring plans, suggesting future MSEs may be able to learn from the set including monitoring methodology as a driver.
+Additionally, the use of value of information analyses could expand upon the benefits of MSE as a tool to direct scientific resources [@runge_2011; @moore_2012].
 While discussions of fisheries research and management seem to be moving toward ecosystem effects, spatial models, and multiple species models, these characteristics were rarely considered in the set of MSEs we reviewed.

### 2020-04-10 R/MSE-manuscript.Rmd
Edit text through the discussion!

@@
 
-Management strategy evaluations typically do assess the consequences of uncertainty for achieving management goals, but other factors influencing the consequences of management strategies get more sporadic treatment.
+We found that management strategy evaluations do typically assess the consequences of uncertainty for achieving management goals, but other factors influencing the consequences of management strategies get more sporadic treatment.
 Data limitation is a frequent concern in fisheries management, and contributes to the degree of uncertainty present in fisheries management.
@@
 
-Managing natural resources entails achieving the objectives of people interacting with a natural resource, so decision processes are likely to produce more desirable results when they include peoples objectives.
-These objectives are likely to include conservation, yield, economic, and social concerns, however the MSEs we reviewed don't cover all categories of objectives, as yield and economic objectives were not included in about one quarter of MSEs, and social objectives were absent from all but 2 of the 30 evaluations.
-While we are unable to know the cause for this with more explicit documentatin of MSE processes, our results may provide some clues.
+Managing natural resources entails achieving the objectives of people interacting with a natural resource, so decision processes are likely to produce more desirable results when they include objectives that reflect stakeholders' and society's values.
+These objectives are likely to include conservation, yield, economic, and social concerns.
+As noted, MSEs are described as a tool to evaluate what management strategy best achieves objectives, however the MSEs we reviewed don't cover all categories of objectives.
+Yield and economic objectives were not included in about one quarter of MSEs, and broader social objectives were absent from all but 2 of the 30 evaluations.
+While we are unable to know the cause for this without more explicit documentation of MSE processes, our results may provide some clues.
 
-Scientists led, and participated in most of the MSEs, while decision-makers, decision making institutions, and stakeholders were relatively infrequent leaders and participants in the process.
-The lack of participation by stakeholders may result in less frequent inclusion of social objectives.
+Scientists led, and participated in most of the MSEs, while decision-makers, decision making institutions, and stakeholders were relatively infrequent leaders or participants in the process.
+The lack of participation by stakeholders may result in less frequent inclusion of social objectives, as well as yield and economic objectives specific to a particular fishery.
 Another factor may be that MSE models are likely constructed mainly by ecosystem modelers, who are likely to be more comfortable modeling the consequences of conservation and yield objectives than economic or social objectives.
@@
 There also appears to be some degree of borrowing occurring in which scientists use a set of generic fishery objectives when conducting a MSE.
-Perhaps because their were few social objectives included in these MSE, there was little inclusion of experts to aid in the prediction of consequences, or consequence prediction techniques other than simulation modeling.
+Perhaps because their were few social objectives included in these MSEs, there was little inclusion of experts to aid in the prediction of consequences, or consequence prediction techniques other than simulation modeling.
 
@@
 The inclusion of facilitators in decision processes aid information gain and conflict resolution, increase trust between participants, and lead to more beneficial environmental and social outcomes [@Vente].
-Facilitators and decision analysts also aid decision processes by helping to diagnose the problem, which can result in selecting more appropriate analysis techniques for the problem type.
-Finally, the additional structure and focus on the decision process would likely result in more explicit documentation when facilitators and decision analysts participate in MSE, as resulted in some of the MSEs reviewed here [@williams2011; @smith2013; @kolody2008].
+The additional structure and focus on the decision process would likely result in more explicit documentation when facilitators and decision analysts participate in MSE, as resulted in some of the MSEs reviewed here [@williams2011; @smith2013; @kolody2008].
+Facilitators and decision analysts can also aid decision processes by helping to diagnose the problem, which can result in selecting more appropriate analysis techniques for the problem type, elicit a more complete and representative set of objectives, and select analysis tools to best address decision impediments to identify the optimal management strategy [@Cummings_inprep].
+
+The MSE process may be aided by utilizing the techniques from the structured decision making approach to decision analysis, such as expanding the focus on problem framing at the outset of MSEs, eliciting fully representative sets of objectives, and utilizing multiple attribute utility theory to evaluate tradeoffs between objectives (Cummings, in prep).
+Utilizing the structured decision making approach also eases the documentation process by breaking decision making into more discrete stages.
+However, there may be practical considerations, such as time or budgetary constraints that limit how complex the stages of a MSE are, and the degree of documentation that occurs.
+Where the expertise and budget is available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.
 
@@
 Different decision makers can rationally look at the same prediction of consequences and select a different preferred alternative if they treat the importance of the objectives differently.
-Therefore, MSE documentation often is not supporting replication or learning due to the frequent absence of trade-off and decision documentation.
-
-The MSE process may be aided by utilizing the techniques from the structured decision making approach to decision analysis (Cummings, in prep).
-Utilizing the structured decision making approach eases the documentation process by breaking decision making into more discrete stages.
-There may be practical considerations, such as time or budgetary constraints that limit how complex the stages of a MSE are, and the degree of documentation that occurs.
-Where the expertise and budget is available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.
+Therefore, MSE documentation often is not fully supporting replication or learning due to the frequent absence of trade-off and decision documentation.
 
 ### Explore and expand upon this review
-To aid future learning from the MSEs that have been conducted to date we produced a shiny application [@R-Shiny] that enables interactive results to accompany this article and entry of additional MSE information (<https://jonathancummings.shinyapps.io/MSEreview/>).
-We hope that as MSE are conducted in the future they are entered here so that MSE practitioners can learn from each other.
+To aid learning from the documentation that has occurred, and provide ready access to future MSE documentation we produced a shiny application [@R-Shiny] to enable an interactive means to explore the MSE literature (<https://jonathancummings.shinyapps.io/MSEreview/>).
+The MSE review shiny application enables user to interactively explore the results presented in this article.
+The application includes additional analyses that may be of interest, such as an analysis of what journals MSEs have been published in, who is authoring MSEs, and more.
+Users may also filter by driver, location, species, or other characteristics of MSEs to explore a set of MSEs of interest and relevance to aid learning in the MSE practitioner community.
+We hope that as MSE are conducted they are entered here so that MSE practitioners can further learn from each other in the future.
 

### 2020-04-10 R/MSE-manuscript.Rmd
Spell check

@@
 The picture painted by MSE documentation is primarily one of scientist driven simulation studies rather than management driven decision processes, with limited process documentation occurring.
-We hope providing a repository for reviews of previous management strategy evaluatons will aid and speed learning among practitioners of management strategy evaluation.
+We hope providing a repository for reviews of previous management strategy evaluations will aid and speed learning among practitioners of management strategy evaluation.
 ---
@@
 
-An incresingly recognized challenge in fisheries management is responding to changes in the productivity of fisheries stocks as a result of climate change [@busch_climate_2016; @brander_global_2007].
+An increasingly recognized challenge in fisheries management is responding to changes in the productivity of fisheries stocks as a result of climate change [@busch_climate_2016; @brander_global_2007].
 We also conduct a focused analysis of MSE processes that address the influence of climate change on fisheries management, highlighting aspects of MSE processes that model climate change as a driver of system status.
@@
     rename("MSE Type"=MSE_Type) %>% 
-  kable(caption = "Januaray 8, 2019 Web of Science 'management strategy evaluation' search results") %>% 
+  kable(caption = "January 8, 2019 Web of Science 'management strategy evaluation' search results") %>% 
   kable_styling() %>% 
@@
 The primary participants and leaders of the MSE processes were scientists, while some MSE process were led by governments and management agencies (Figure \@ref(fig:participants) Process).
-Scientists were exlicitly the leaders -- or the sole participants and therefore seemingly the leaders -- in 73 percent of the MSE processes, while 27 percent of processes lacked enough documentation of participants and leadership such that we were unable to identify a presumed leader of the process.
+Scientists were explicitly the leaders -- or the sole participants and therefore seemingly the leaders -- in 73 percent of the MSE processes, while 27 percent of processes lacked enough documentation of participants and leadership such that we were unable to identify a presumed leader of the process.
 
@@
 
-Climate change MSEs were more likely to explicitly document the trade-off analysis stage, equaly likely to document the alternatives stages, but were less likely to document the other stages in the decision process.
+Climate change MSEs were more likely to explicitly document the trade-off analysis stage, equally likely to document the alternatives stages, but were less likely to document the other stages in the decision process.
 ```{r, CCstages}
@@
 
-Including climate change in a MSE appears to reduce the complexity of other stages in the MSE process, in particular those stages that utilize input from additional particapants.
+Including climate change in a MSE appears to reduce the complexity of other stages in the MSE process, in particular those stages that utilize input from additional participants.
 On average climate change MSEs consider fewer drivers, consider less diverse objectives, and explicitly document fewer stages of the MSE process.
-If the MSE process is time, effort, or participant limited the inclusion of climate change may be utlizing time, effort, or outside participants to focus on climate change rather than other aspects of the MSE process.
+If the MSE process is time, effort, or participant limited the inclusion of climate change may be utilizing time, effort, or outside participants to focus on climate change rather than other aspects of the MSE process.
 
@@
 Data limitation is a frequent concern in fisheries management, and contributes to the degree of uncertainty present in fisheries management.
-Next to accounting for uncertainty, accessing the influence of data quality and availablity, i.e.,  monitoring methodology was the most prevalent driver included in MSEs.
+Next to accounting for uncertainty, accessing the influence of data quality and availability, i.e.,  monitoring methodology was the most prevalent driver included in MSEs.
 This likely points to a desire for the design and use of cost-effective fisheries monitoring plans, suggesting future MSEs may be able to learn from the set including monitoring methodology as a driver.

### 2020-04-10 R/MSE-manuscript.Rmd
minor text edits for spelling and grammar

@@
 output:
+  bookdown::word_document2:
   bookdown::html_document2:
     number_sections: false
-  bookdown::word_document2:
   bookdown::pdf_document2:
@@
 Given that scientific publication is the primary means of communicating the results of MSE, we review the scientific literature to assess the extent to which the process and aims of MSE, and adaptive management are being achieved and supported by the scientific literature.
-To assess this we determine how MSEs utilize standard components of decision making processes.
+To assess this, we determine how MSEs utilize standard components of decision making processes.
 We use the structured decision making (SDM) process [@SDM] -- the decision making framework in which adaptive management occurs -- as our framework for evaluation.
-Using this framework we review published MSEs to:
+Using this framework, we review published MSEs to:
 
@@
 We used the structured decision making framework [@conroy2013decision; @gregory2012structured] as our framework for evaluating management strategy evaluation documents (Figure \@ref(fig:SDMfigure)).
-The structured decision making processes is composed of the six process stages: problem definition, objective elicitation, alternative development, consequence prediction, trade-off analysis, and decision implementation.
+The structured decision making process is composed of the six process stages: problem definition, objective elicitation, alternative development, consequence prediction, trade-off analysis, and decision implementation.
 
@@
 In many cases the process used to elicit objectives and alternatives was not explicit.
-In these cases we used what information was provided to glean who participated in the production of those components of the MSE.
+In these cases, we used what information was provided to glean who participated in the production of those components of the MSE.
 We record and report who those participants were, referring to them as subjective participants in the objectives or alternatives process.
@@
 
-Climate change MSEs were more likely to explicitly document the trade-off analysis stage, equally likely to document the alternatives stages, but were less likely to document the other stages in the decision process.
+Climate change MSEs were more likely to explicitly document the trade-off analysis stage, equally likely to document the alternatives stages, but were less likely to document the other stages in the decision process (Table \@ref(tab:CCstages)).
 ```{r, CCstages}
@@
 However, there may be practical considerations, such as time or budgetary constraints that limit how complex the stages of a MSE are, and the degree of documentation that occurs.
-Where the expertise and budget is available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.
+Where the expertise and budget are available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.
 

### 2020-04-16 R/MSE-manuscript.Rmd
fix typo
change supplement to appendix; increase heading level of objectives; fix plot.margin to avoid cropping figures

@@
     labels=c("Decision","Tradeoffs",
-             "Alternatives","Objectives","Problem","Process")) + theme_bw()
+             "Alternatives","Objectives","Problem","Process")) + theme_bw() + theme(plot.margin = margin(0, 0.5, 0, 0.25, "cm"))
 
@@
   scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
-  ylab("Percent")+xlab(NULL)+coord_flip() + theme_bw()
+  ylab("Percent")+xlab(NULL)+coord_flip() + theme_bw() + theme(plot.margin = margin(0, 0.5, 0, 0.25, "cm"))
 
@@
   scale_y_continuous(limits=c(0,100),expand = c(0,0))+
-  ylab("Percent")+xlab(NULL)+coord_flip()+ theme_bw()
+  ylab("Percent")+xlab(NULL)+coord_flip()+ theme_bw()+ theme(plot.margin = margin(0, 0.5, 0, 0.25, "cm"))
 ```
@@
 
-### Objectives 
+## Objectives 
 Our overall objective is to determine the extent that MSE achieves the aspirational aims outlined above, and whether MSE documentation supports learning within the MSE practitioner community.
@@
 This search returned 264 results.
-We reviewed a random sample of 30 articles that document a MSE (supplement 2), removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
+We reviewed a random sample of 30 articles that document a MSE (Appendix B), removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
 After removing other articles from the original set of 264 articles, 154 articles document a MSE.
-We reviewed 16 climate change MSE articles (supplement 2).
+We reviewed 16 climate change MSE articles (Appendix B).
 Twelve of these articles came from our original January 8th, 2019 Web of Science search.
@@
 We collected additional information about the MSEs we reviewed beyond what is reported here.
-For a full description of the information recorded in the relational database see the supplemental information (supplement 1), and for the full data set and results see our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
+For a full description of the information recorded in the relational database see the supplemental information (Appendix A), and for the full data set and results see our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
 
@@
 To aid learning from the documentation that has occurred, and provide ready access to future MSE documentation we produced a shiny application [@R-Shiny] to enable an interactive means to explore the MSE literature (<https://jonathancummings.shinyapps.io/MSEreview/>).
-The MSE review shiny application enables user to interactively explore the results presented in this article.
+The MSE review shiny application enables users to interactively explore the results presented in this article.
 The application includes additional analyses that may be of interest, such as an analysis of what journals MSEs have been published in, who is authoring MSEs, and more.

### 2020-04-20 R/MSE-manuscript.Rmd
Edit map to produce greyscale image; edit header levels.

@@
 # Get map background for plotting the map
-world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
+world <- borders("world", colour="gray85", fill="gray96", alpha=0.75) # create a layer of borders
 ```
@@
 MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
-  geom_point(size=2.5) + geom_point(size = 2.5, colour = "grey", shape = 1) +
-  scale_color_manual(values=c("#66CCCC","#006666"))+theme_void() +
+  geom_point(size=2.5) + geom_point(size = 2.5, colour = "gray55", shape = 1) +
+  scale_color_manual(values=c("gray10","gray55"))+theme_void() +
   theme(legend.position = c(0.15, 0.25))
@@
 # Results 
-### What questions and management objectives do MSEs address? 
+## What questions and management objectives do MSEs address? 
 We found that uncertainty was the main driver of the system state included in management strategy evaluations, with uncertainty included in 21 of the 30 evaluations we reviewed in our random sample (Table \@ref(tab:drivers)).
@@
 
-### Who is involved in MSEs? 
+## Who is involved in MSEs? 
 MSEs were rarely explicit about the roles participants played in MSE processes, as only one MSE documented the roles participants played in the process (Table \@ref(tab:roles)).
@@
 
-### Explore and expand upon this review
+## Explore and expand upon this review
 To aid learning from the documentation that has occurred, and provide ready access to future MSE documentation we produced a shiny application [@R-Shiny] to enable an interactive means to explore the MSE literature (<https://jonathancummings.shinyapps.io/MSEreview/>).

### 2020-04-20 R/Figures.R

@@ -20,56 +20,62 @@ library(odbc) # database connection
 library(RMySQL) # MySQL scripting in R
 library(DBI) # database interface
 library(dbplyr) # database interface for plyr
+library(kableExtra) # extra formating of the tables
+library(here) # file directory assistance
 
 ###### To open a connection to the database: #####
 # Copy into command prompt:
 ## cloud_sql_proxy -instances=alpine-tracker-230222:us-east1:mse-review=tcp:3306
 
-##### Connect to database and get data #####
-# Create connection function
-getSqlConnection <- function(){
-  cn <-
-    dbConnect(
-      RMySQL::MySQL(),
-      dbname = 'MSEreview',
-      host = '127.0.0.1',
-      port=3306,
-      username = 'root',
-      password = '2u8aNDdur1aFdb8z')
-  return(cn)
-}
-# Open connection
-con <- getSqlConnection()
+# ##### Connect to database and get data #####
+# # Create connection function
+# getSqlConnection <- function(){
+#   cn <-
+#     dbConnect(
+#       RMySQL::MySQL(),
+#       dbname = 'MSEreview',
+#       host = '127.0.0.1',
+#       port=3306,
+#       username = 'root',
+#       password = '2u8aNDdur1aFdb8z')
+#   return(cn)
+# }
+# # Open connection
+# con <- getSqlConnection()
+# 
+# # load("C:/Users/jcummings/OneDrive - UMASS Dartmouth/SMAST/Research/MSE Problem Framing/MSEreview/MSEreview.RData")
+# 
+# # Obtain data tables
+# study<-dbReadTable(con,"tblStudy")
+# # study<-read_xlsx("DB files - Excel/tblStudy.xlsx")
+# mgmt<-dbReadTable(con,"tblStudyManagement")
+# # mgmt<-read_xlsx("DB files - Excel/tblStudyManagementTools.xlsx")
+# # mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
+# obj<-dbReadTable(con,"tblStudyObjectives")
+# # obj<-read_xlsx("DB files - Excel/tblStudyObjectives.xlsx")
+# # obj$fkStudyID<-parse_integer(obj$fkStudyID)
+# fields<-dbReadTable(con,"tblStudyFields")
+# # fields<-read_xlsx("DB files - Excel/tblStudyFields.xlsx")
[…]

### 2020-04-23 R/Figures.R
Add analysis of the degree of randomness of the random sample

@@ -623,3 +623,49 @@ as_tibble(altcat.data_CC)
 #   summarise('AVG'=sum(Number)) %>% 
 #   mutate(count=c(11,30)) %>% 
 #   mutate('Per MSE'=AVG/count)
+
+#Analyze the year that MSEs and the random sample of MSEs were published
+
+# Read in data on all publications
+tblWOS <- read_excel("data/DB files - Excel/tblWOS.xlsx")
+# Rename columns, filter to only MSEs, and select year published and whether the publication was in the 
+# random sample
+SampleCheck<-tblWOS %>%
+  rename(IsMSE="Is MSE?") %>% 
+  filter(IsMSE==TRUE) %>% 
+  select(YearPub,IncludeInPub)
+# Plot result
+ggplot(SampleCheck,aes(x=YearPub,color=IncludeInPub,fill=IncludeInPub))+geom_histogram(binwidth = 1)+
+  scale_color_grey()+scale_fill_manual(values=c("gray35","gray50"))+theme_bw()+
+  labs(col="Random Sample",fill="Random Sample")
+# make a table of the result
+SampleCount<-SampleCheck %>% 
+  group_by(YearPub,IncludeInPub) %>% 
+  count() %>% 
+  pivot_wider(names_from = IncludeInPub,values_from = n) %>% 
+  rename(Selected="TRUE",
+         NotSelected="FALSE") %>%
+  mutate(percent=Selected/(Selected+NotSelected)*100)
+
+# Rename columns, filter to only MSEs, and select journal and whether the publication was in the 
+# random sample
+SampleJournals<-tblWOS %>%
+  rename(IsMSE="Is MSE?") %>% 
+  filter(IsMSE==TRUE) %>% 
+  select(Journal,IncludeInPub) %>%
+  group_by(Journal,IncludeInPub) %>% 
+  count() %>% 
+  filter(n>2) %>% 
+  drop_na() %>% 
+  arrange(-n)
+# Plot result
+ggplot(SampleJournals,aes(x=reorder(Journal, n),y=n,color=IncludeInPub,fill=IncludeInPub))+geom_col()+
+  scale_color_grey()+scale_fill_manual(values=c("gray35","gray50"))+theme_bw()+
+  labs(col="Random Sample",fill="Random Sample")+coord_flip()
+# make a table of the result
+SampleJournalsTable<-SampleJOurnals %>% 
[…]

### 2020-04-23 R/AnalysisExcel.R
Add tblWOS to Rdata file

@@ -20,10 +20,6 @@ library(mapdata) # basemap generation
 ##### 
 # Script
 
-# Get map background for plotting the map
-world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
-
-
 #####
 ###-- Data Processing --###
 
@@ -34,6 +30,7 @@ mgmt$fkStudyID<-parse_integer(mgmt$fkStudyID)
 obj<-read_xlsx("data/DB files - Excel/tblStudyObjectives.xlsx")
 obj$fkStudyID<-parse_integer(obj$fkStudyID)
 fields<-read_xlsx("data/DB files - Excel/tblStudyFields.xlsx")
+tblWOS <- read_excel("data/DB files - Excel/tblWOS.xlsx")
 
 #####
 # Edit tables for joining
@@ -107,6 +104,9 @@ alt.col<-c("ManagementTool",
 map.col<-c("Latitude",
            "Longitude",
            "Citation")
+
+# Get map background for plotting the map
+world <- borders("world", colour="gray50", fill="gray50", alpha=0.75) # create a layer of borders
 
 #####
 ###-- Shiny App --###

### 2020-04-23 R/Figures.R
read in tblWOS from RData file

@@ -626,18 +626,17 @@ as_tibble(altcat.data_CC)
 
 #Analyze the year that MSEs and the random sample of MSEs were published
 
-# Read in data on all publications
-tblWOS <- read_excel("data/DB files - Excel/tblWOS.xlsx")
 # Rename columns, filter to only MSEs, and select year published and whether the publication was in the 
 # random sample
 SampleCheck<-tblWOS %>%
   rename(IsMSE="Is MSE?") %>% 
-  filter(IsMSE==TRUE) %>% 
+  filter(IsMSE==TRUE&YearPub<2019) %>% 
   select(YearPub,IncludeInPub)
-# Plot result
+# Plot result as a histogram
 ggplot(SampleCheck,aes(x=YearPub,color=IncludeInPub,fill=IncludeInPub))+geom_histogram(binwidth = 1)+
   scale_color_grey()+scale_fill_manual(values=c("gray35","gray50"))+theme_bw()+
   labs(col="Random Sample",fill="Random Sample")
+
 # make a table of the result
 SampleCount<-SampleCheck %>% 
   group_by(YearPub,IncludeInPub) %>% 
@@ -645,7 +644,14 @@ SampleCount<-SampleCheck %>%
   pivot_wider(names_from = IncludeInPub,values_from = n) %>% 
   rename(Selected="TRUE",
          NotSelected="FALSE") %>%
-  mutate(percent=Selected/(Selected+NotSelected)*100)
+  mutate_at(vars(Selected,NotSelected),~replace_na(., 0)) %>% 
+  mutate(percent=Selected/(Selected+NotSelected)*100) %>% 
+  pivot_longer(cols = c(Selected,NotSelected),names_to = "RandomSample") %>% 
+  arrange(value)
+# Plot result as area
+ggplot(SampleCount,aes(x=YearPub,y=value,color=RandomSample,fill=RandomSample))+geom_area()+geom_point(position = "stack")+
+  scale_color_grey()+scale_fill_manual(values=c("gray35","gray50"))+theme_bw()+
+  labs(y="Publication Count",x="Year Published")
 
 # Rename columns, filter to only MSEs, and select journal and whether the publication was in the 
 # random sample
@@ -654,18 +660,28 @@ SampleJournals<-tblWOS %>%
   filter(IsMSE==TRUE) %>% 
   select(Journal,IncludeInPub) %>%
[…]

### 2020-04-23 R/Figures.R
Fix Not Selected and NotSelected column and value naming for wrangling and plotting

@@ -630,31 +630,32 @@ as_tibble(altcat.data_CC)
 # random sample
 SampleCheck<-tblWOS %>%
   rename(IsMSE="Is MSE?") %>% 
-  filter(IsMSE==TRUE&YearPub<2019) %>% 
+  filter(IsMSE==TRUE&YearPub<2019) %>%
+  mutate(IncludeInPub = replace(IncludeInPub, IncludeInPub == TRUE, "Selected")) %>% 
+  mutate(IncludeInPub = replace(IncludeInPub, IncludeInPub == FALSE, "Not Selected")) %>% 
   select(YearPub,IncludeInPub)
 # Plot result as a histogram
 ggplot(SampleCheck,aes(x=YearPub,color=IncludeInPub,fill=IncludeInPub))+geom_histogram(binwidth = 1)+
   scale_color_grey()+scale_fill_manual(values=c("gray35","gray50"))+theme_bw()+
-  labs(col="Random Sample",fill="Random Sample")
+  labs(col="Random Sample",fill="Random Sample",x="Publication Year")+theme(legend.position = c(0.15, 0.85))
 
-# make a table of the result
-SampleCount<-SampleCheck %>% 
-  group_by(YearPub,IncludeInPub) %>% 
-  count() %>% 
-  pivot_wider(names_from = IncludeInPub,values_from = n) %>% 
-  rename(Selected="TRUE",
-         NotSelected="FALSE") %>%
-  mutate_at(vars(Selected,NotSelected),~replace_na(., 0)) %>% 
-  mutate(percent=Selected/(Selected+NotSelected)*100) %>% 
-  pivot_longer(cols = c(Selected,NotSelected),names_to = "RandomSample") %>% 
-  arrange(value)
-# Plot result as area
-ggplot(SampleCount,aes(x=YearPub,y=value,color=RandomSample,fill=RandomSample))+geom_area()+geom_point(position = "stack")+
-  scale_color_grey()+scale_fill_manual(values=c("gray35","gray50"))+theme_bw()+
-  labs(y="Publication Count",x="Year Published")
+# # make a table of the result
+# SampleCount<-SampleCheck %>% 
+#   group_by(YearPub,IncludeInPub) %>% 
+#   count() %>% 
+#   pivot_wider(names_from = IncludeInPub,values_from = n) %>% 
+#   rename(Selected="TRUE",
+#          NotSelected="FALSE") %>%
[…]

### 2020-04-23 R/MSE-manuscript.Rmd
Add year and publication histograms to the method section

@@
 After removing other articles from the original set of 264 articles, 154 articles document a MSE.
+The number of MSEs published each year is trending higher (Figure \@ref(fig:YearHistogram)), with the ICES Journal of Marine Science and Fisheries Research as the leading publication outlets ((Figure \@ref(fig:PublicationHistogram))).
+
 We reviewed 16 climate change MSE articles (Appendix B).
@@
     add_footnote(c("We also reviewed 4 additional climate change MSE articles published in 2019 for a total of 16."), notation = "symbol")
+```
+
+```{r YearHistogram}
+# Analyze the year that MSEs were published and sampled from
+# Wrangle data
+SampleCheck<-tblWOS %>%
+  rename(IsMSE="Is MSE?") %>% # rename column
+  filter(IsMSE==TRUE&YearPub<2019) %>% #filter data
+  mutate(IncludeInPub = replace(IncludeInPub, IncludeInPub == TRUE, "Selected")) %>% 
+  mutate(IncludeInPub = replace(IncludeInPub, IncludeInPub == FALSE, "Not Selected")) %>% #rename values
+  select(YearPub,IncludeInPub) # select desired columns
+# Plot result as a histogram
+YearHistogram<-ggplot(SampleCheck,aes(x=YearPub,color=IncludeInPub,fill=IncludeInPub))+
+  geom_histogram(binwidth = 1)+scale_color_grey()+scale_fill_manual(values=c("gray35","gray50"))+theme_bw()+
+  labs(col="Random Sample",fill="Random Sample",x="Publication Year")+theme(legend.position = c(0.15, 0.85))
+YearHistogram
+```
+
+```{r PublicationHistogram}
+# Analyze the journal that MSEs were published in and sampled from
+# Wrangle data
+SampleJournals<-tblWOS %>%
+  rename(IsMSE="Is MSE?") %>% # rename column
+  filter(IsMSE==TRUE&YearPub<2019) %>% # filter data
+  select(Journal,IncludeInPub) %>% # select columns
+  group_by(Journal,IncludeInPub) %>% # group summary by
+  count() %>% # count by groups
+  arrange(-n) %>% # sort
+  ungroup() %>% # ungroup
+  mutate(IncludeInPub = replace(IncludeInPub, IncludeInPub == TRUE, "Selected")) %>% 
+  mutate(IncludeInPub = replace(IncludeInPub, IncludeInPub == FALSE, "Not Selected")) %>% # rename values
+  drop_na() %>% # Drop entries without journal sources
+  pivot_wider(names_from = IncludeInPub,values_from = n) %>% # pivot data
+  mutate_at(vars(Selected,"Not Selected"),~replace_na(., 0)) %>% # replace NA with 0
+  rename(NotSelected="Not Selected") %>% 
+  mutate(Total=Selected+NotSelected) %>% # calculate Total published for each journal
+  filter(Total>1) %>% # filter to remove journals with only one publication
+  rename("Not Selected"=NotSelected) %>%
+  pivot_longer(cols = c("Not Selected","Selected"),names_to = "IncludeInPub",values_to = "n") %>%  #pivot data
+  select(Journal,IncludeInPub,n) # select desired columns
+PublicationHistogram<-ggplot(SampleJournals,aes(x=reorder(Journal, n),y=n,color=IncludeInPub,fill=IncludeInPub))+
+  geom_col()+  scale_color_grey()+scale_fill_manual(values=c("gray35","gray50"))+theme_bw()+
+  labs(col="Random Sample",fill="Random Sample")+coord_flip()+
+  theme(legend.position = c(0.75, 0.125),axis.title.y=element_blank())+
+  labs(y="Publication Count")
+PublicationHistogram
 ```

### 2020-04-23 R/MSE-manuscript.Rmd
Fix strsplit split value that I broke AGAIN, ugh!

@@
 part.data_pub<- part.data_pub %>%
-    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+    purrr::map(~ strsplit(as.character(.),split=";")) %>%
     purrr::map(unlist) %>%
@@
 part.data_CC<- part.data_CC %>%
-  purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+  purrr::map(~ strsplit(as.character(.),split=";")) %>%
   purrr::map(unlist) %>%
@@
     select(Drivers) %>%
-    purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+    purrr::map(~ strsplit(as.character(.),split=";")) %>%
     purrr::map(unlist) %>%
@@
   select(Drivers) %>%
-  purrr::map(~ strsplit(as.character(.),split=", ")) %>%
+  purrr::map(~ strsplit(as.character(.),split=";")) %>%
   purrr::map(unlist) %>%

### 2020-04-23 R/MSE-manuscript.Rmd
Conver lat and long from type character to double

@@
   mutate(Drivers=str_extract(Drivers, "Climate Change")) %>%
-  mutate(Drivers=replace_na(Drivers, "Random Sample"))
+  mutate(Drivers=replace_na(Drivers, "Random Sample")) %>% 
+  type_convert()
 

### 2020-04-23 R/MSE-manuscript.Rmd
Edit methods to include year published and journal of publication figures

@@
 After removing other articles from the original set of 264 articles, 154 articles document a MSE.
-The number of MSEs published each year is trending higher (Figure \@ref(fig:YearHistogram)), with the ICES Journal of Marine Science and Fisheries Research as the leading publication outlets ((Figure \@ref(fig:PublicationHistogram))).
-
-We reviewed 16 climate change MSE articles (Appendix B).
-Twelve of these articles came from our original January 8th, 2019 Web of Science search.
-To give as full a picture of the climate change MSE articles to date as possible we also reviewed the 4 climate change MSE articles published in 2019.
+The number of MSEs published each year is trending higher (Figure \@ref(fig:YearHistogram)), with the ICES Journal of Marine Science and Fisheries Research as the leading publication outlets (Figure \@ref(fig:PublicationHistogram)).
 
@@
 
-```{r YearHistogram}
+```{r YearHistogram, fig.cap="Management Strategy Evaluations published per year for those selected and not selected in our random sample"}
 # Analyze the year that MSEs were published and sampled from
@@
 
-```{r PublicationHistogram}
+We reviewed 16 climate change MSE articles (Appendix B).
+Twelve of these articles came from our original January 8th, 2019 Web of Science search.
+To give as full a picture of the climate change MSE articles to date as possible we also reviewed the 4 climate change MSE articles published in 2019.
+
+```{r PublicationHistogram, , fig.cap="Count of management strategy evaluations published by journal for those journal that have published at least two management strategy evaluations"}
 # Analyze the journal that MSEs were published in and sampled from
@@
 
-```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=4, out.width="100%"}
+```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=6, out.width="100%"}
 MSE.map +

### 2020-05-04 R/Figures.R
Examining the years of climate change MSE publications

@@ -687,3 +687,16 @@ ggplot(SampleJournalsFiltered,aes(x=reorder(Journal, n),y=n,color=IncludeInPub,f
   labs(col="Random Sample",fill="Random Sample")+coord_flip()+
   theme(legend.position = c(0.75, 0.125),axis.title.y=element_blank())+
   labs(y="Publication Count")
+
+# Examining the years of climate change MSE publications
+SampleClimate<-data_CC %>%
+  select(YearPub) %>% 
+  group_by(YearPub) %>% 
+  count() %>% 
+  ungroup() %>% 
+  add_row(YearPub=c("1999","2000","2001","2002","2003","2004","2005","2006","2007","2008",
+                    "2012","2014","2015"),n=0)
+ggplot(SampleClimate,aes(x=YearPub,y=n))+geom_col()+
+  scale_color_grey()+theme_bw()+
+  labs(x="Publication Year")
+

### 2020-05-04 R/Figures.R
Also examine the journal climate change publications

@@ -688,7 +688,7 @@ ggplot(SampleJournalsFiltered,aes(x=reorder(Journal, n),y=n,color=IncludeInPub,f
   theme(legend.position = c(0.75, 0.125),axis.title.y=element_blank())+
   labs(y="Publication Count")
 
-# Examining the years of climate change MSE publications
+# Examining the years and journals of climate change MSE publications
 SampleClimate<-data_CC %>%
   select(YearPub) %>% 
   group_by(YearPub) %>% 
@@ -699,4 +699,11 @@ SampleClimate<-data_CC %>%
 ggplot(SampleClimate,aes(x=YearPub,y=n))+geom_col()+
   scale_color_grey()+theme_bw()+
   labs(x="Publication Year")
+SampleClimateJournal<-data_CC %>%
+  select(Journal) %>% 
+  group_by(Journal) %>% 
+  count() %>% 
+  arrange(-n)
+ggplot(SampleClimateJournal,aes(x=reorder(Journal,n),y=n))+geom_col()+  scale_color_grey()+theme_bw()+  coord_flip()+
+  theme(axis.title.y=element_blank())+  labs(y="Publication Count")
 

### 2020-05-04 R/MSE-manuscript.Rmd
Change the shape of the climate change and random sample points in MSE.map

@@
 # plot MSEs on map
-MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
-  geom_point(size=2.5) + geom_point(size = 2.5, colour = "gray55", shape = 1) +
-  scale_color_manual(values=c("gray10","gray55"))+theme_void() +
+MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,group=Drivers)) + world +
+  geom_point(aes(shape=Drivers,color=Drivers,fill=Drivers),size=2.5) + 
+  geom_point(aes(shape=Drivers),size = 2.5, colour = "gray45") +
+  scale_shape_manual(values=c(21,22))+scale_fill_manual(values=c("gray10","gray80"))+theme_void() +
   theme(legend.position = c(0.15, 0.25))

### 2020-05-04 R/Figures.R

@@ -468,9 +468,10 @@ per.MSE<-obj.dataTable %>%
 world <- borders("world", colour="gray85", fill="gray96", alpha=0.75) # create a layer of borders
 
 # plot MSEs on map
-MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
-  geom_point(size=2.5) + geom_point(size = 2.5, colour = "gray55", shape = 1) +
-  scale_color_manual(values=c("gray10","gray55"))+theme_void() +
+MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,group=Drivers)) + world +
+  geom_point(aes(shape=Drivers,color=Drivers,fill=Drivers),size=2.5) + 
+  geom_point(aes(shape=Drivers),size = 2.5, colour = "gray45") +
+  scale_shape_manual(values=c(21,22))+scale_fill_manual(values=c("gray10","gray80"))+theme_void() +
   theme(legend.position = c(0.15, 0.25))
 
 # plot explicit documentation of steps or components of MSE processes 
@@ -689,6 +690,7 @@ ggplot(SampleJournalsFiltered,aes(x=reorder(Journal, n),y=n,color=IncludeInPub,f
   labs(y="Publication Count")
 
 # Examining the years and journals of climate change MSE publications
+# Year of publication
 SampleClimate<-data_CC %>%
   select(YearPub) %>% 
   group_by(YearPub) %>% 
@@ -696,14 +698,14 @@ SampleClimate<-data_CC %>%
   ungroup() %>% 
   add_row(YearPub=c("1999","2000","2001","2002","2003","2004","2005","2006","2007","2008",
                     "2012","2014","2015"),n=0)
-ggplot(SampleClimate,aes(x=YearPub,y=n))+geom_col()+
-  scale_color_grey()+theme_bw()+
-  labs(x="Publication Year")
+ggplot(SampleClimate,aes(x=YearPub,y=n))+geom_col()+scale_color_grey()+theme_bw()+labs(x="Publication Year")+
+  theme(axis.title.y=element_blank())
+# Journal
 SampleClimateJournal<-data_CC %>%
   select(Journal) %>% 
   group_by(Journal) %>% 
   count() %>% 
   arrange(-n)
 ggplot(SampleClimateJournal,aes(x=reorder(Journal,n),y=n))+geom_col()+  scale_color_grey()+theme_bw()+  coord_flip()+
[…]

### 2020-05-07 R/MSE-manuscript.Rmd
Update to refect text edits in manuscript submission

@@
 abstract: The aim and vision for fisheries management that comes from management strategy evaluation and adaptive management is a noble and aspirational vision.
-We review MSE documentation in the published literature to appraise the achievement of this vision, finding that reality doesn't fully reflect this vision.
-The steps taken to complete the management strategy evaluation process often are not documented.
-Scientists often filled the role of fishery stakeholders, seemingly selecting the objectives and alternatives to evaluate in most management strategy evaluations.
-Facilitators and decision analysts -- participants focused on the decision making process rather than making the decision -- were rarely participants in MSE processes.
+We review MSE documentation in the published literature to appraise the achievement of this vision, finding that reality doesn’t fully reflect this vision.
+As a scientific research tool to evaluate management performance, exemplary applications of MSE in the literature abound.
+Examples of full integration of the MSE process as a stakeholder-based participatory, management decision support framework are less common in the literature, and these processes are not as well documented as the technical modeling detail.
+The steps taken to complete the management strategy evaluation process often are not documented and scientists often filled the role of fishery stakeholders, seemingly selecting the objectives and alternatives to evaluate in most management strategy evaluations.
+Facilitators and decision analysts – participants who may aid the process by focusing on the decision-making process rather than making the decision – were rarely participants.
 Conservation, Economic, and Yield focused objectives were commonly considered in MSE processes, while broader social objectives were rarely considered.
-The picture painted by MSE documentation is primarily one of scientist driven simulation studies rather than management driven decision processes, with limited process documentation occurring.
-We hope providing a repository for reviews of previous management strategy evaluations will aid and speed learning among practitioners of management strategy evaluation.
+While the picture painted by MSE documentation is primarily one of scientist driven simulation studies rather than management driven decision processes, with limited process documentation occurring.
+Here we provide a repository for reviews of previous management strategy evaluations with hopes this will aid and speed learning among practitioners of management strategy evaluation.
 ---
@@
 # Introduction 
-The calls for fisheries management to utilize management strategy evaluation and adaptive management have been both aspirational and estimable in their aim and vision.
-For example, management strategy evaluation (MSE) is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@Punt].
-The positive aspects of MSE that have been listed include that it is “A tool that scientists and managers can use to simulate the workings of a fisheries system and allow them to test whether potential harvest strategies — or management procedures — can achieve pre-agreed management objectives.”[@Pew], that it “can be used to identify a ‘best’ management strategy among a set of candidate strategies" [@PuntLab] and “MSE has the advantage of being able to reveal the trade-offs among a range of possible management decisions, and should assist our Commissioners in determining whether the objectives should be weighted differently.
+The calls for fisheries management to use management strategy evaluation and adaptive management have been both aspirational and estimable in their aim and vision.
+For example, management strategy evaluation (MSE) is “widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@Punt].
+The positive aspects of MSE that have been listed include that it is “A tool that scientists and managers can use to simulate the workings of a fisheries system and allow them to test whether potential harvest strategies — or management procedures — can achieve pre-agreed management objectives.”[@Pew], that it “can be used to identify a ‘best’ management strategy among a set of candidate strategies” [@PuntLab] and “MSE has the advantage of being able to reveal the trade-offs among a range of possible management decisions, and should assist our Commissioners in determining whether the objectives should be weighted differently.
 Specifically, to provide the information on which to base a rational decision, given objectives, preferences, and attitudes to risk.” [@IPHC].
@@
 
-Are the results of MSE living up to their descriptions?
-Are MSEs evaluating trade-offs and the consequences of uncertainty?
+Are MSEs meeting the aspirational standards for aiding management-driven processes as outlined in the literature?” Do MSEs evaluate trade-offs and the consequences of uncertainty?
 Do they test potential harvest strategies?
@@
 Was consideration given to the weights on objectives, preferences, and attitudes to risk?
-
-In the descriptions of the MSE process, the process itself is described as being "at the interface between science and policy.” [@PuntLab] and being an “Undertaking [that] requires scientists, managers, and stakeholders to be involved throughout the process." [@IPHC] The International Pacific Halibut Commission (2017) also states, "While the scientists do the modelling, managers must offer extensive input.
+These questions prompted the review that follows.
+Practitioners of MSE have described it as a method that lies “at the interface between science and policy” [@PuntLab] and an “Undertaking [that] requires scientists, managers, and stakeholders to be involved throughout the process” [@IPHC] also states, that “While the scientists do the modelling, managers must offer extensive input.
 Because of the many steps and the iterative process, communication among parties is critical for achieving buy-in on the results of the management strategy evaluation.”
 
+These statements raise additional questions for our review.
 Are MSE processes being conducted in the manner described?
@@
 
-Adaptive management developed in fisheries management as a means to improve the achievement of objectives through time, when learning can aid achieving those objectives [@ADM].
+Similar to MSE, adaptive management was developed in fisheries management as a means to improve the achievement of objectives through time, when learning can aid achieving those objectives [@ADM].
 For example, the Millennium Ecosystem Assessment [-@MEA] defines adaptive management as “A systematic process for continually improving management policies and practices by learning from the outcomes of previously employed policies and practices.
 In active adaptive management, management is treated as a deliberate experiment for purposes of learning.” We contend that scientific publication is essentially a process intended to support adaptive management, in that researchers gain knowledge from the decisions made and outcomes produced by previous work to inform their decisions.
-In the case of fisheries management, are scientific publications of management strategy evaluations supporting adaptive management?
+This led to our review’s final question, are scientific publications of management strategy evaluations supporting adaptive management?
 
 ## Objectives 
-Our overall objective is to determine the extent that MSE achieves the aspirational aims outlined above, and whether MSE documentation supports learning within the MSE practitioner community.
-Given that scientific publication is the primary means of communicating the results of MSE, we review the scientific literature to assess the extent to which the process and aims of MSE, and adaptive management are being achieved and supported by the scientific literature.
-To assess this, we determine how MSEs utilize standard components of decision making processes.
-We use the structured decision making (SDM) process [@SDM] -- the decision making framework in which adaptive management occurs -- as our framework for evaluation.
-Using this framework, we review published MSEs to:
+Our overall objective for this review was to determine the extent to which MSE achieves the aspirational aims outlined above and evaluate whether MSE documentation supports learning within the MSE practitioner community.
+Given that scientific publication is the primary means of communicating the results of MSE, we reviewed the scientific literature to assess this objective.
+We evaluated whether MSEs use standard components of decision making processes using the structured decision making (SDM) process [@SDM] -- the decision making framework in which adaptive management occurs -- as our framework for evaluation.
+Using SDM as our framework, we:
 
-* Determine what questions MSE processes are addressing
-* Identify the objectives and alternatives considered in MSE
-* Determine the prevalence of a) explicit documentation and b) use of decision making steps in the MSE process, and 
-* Identify participants in MSE processes.
+* Determined what questions MSE processes are addressing
+* Identified the objectives and alternatives considered in MSE
+* Determined the prevalence of a) explicit documentation and b) use of decision making steps in the MSE process, and 
+* Identified participants in MSE processes.
 
-An increasingly recognized challenge in fisheries management is responding to changes in the productivity of fisheries stocks as a result of climate change [@busch_climate_2016; @brander_global_2007].
-We also conduct a focused analysis of MSE processes that address the influence of climate change on fisheries management, highlighting aspects of MSE processes that model climate change as a driver of system status.
+As MSE expands and seeks to address new challenges, the insights from past MSEs may become increasingly valuable.
+Fisheries management has mostly focused on fishing impacts with ecosystem status viewed as a background constant.
+The increasing rate of climate change and its influence on fisheries is changing this dynamic, bringing ecosystem status to the forefront of fisheries management [@busch_climate_2016; @brander_global_2007].
+We use climate change focused MSEs as a case study into the use of MSE in an emerging area of study and conduct a focused analysis of MSE processes that address the influence of climate change on fisheries management, highlighting aspects of MSE processes that model climate change as a driver of system status.
 
@@
 This search returned 264 results.
-We reviewed a random sample of 30 articles that document a MSE (Appendix B), removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
+We reviewed a random sample of 30 articles that document a MSE (Appendix A), removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample.
 After removing other articles from the original set of 264 articles, 154 articles document a MSE.
+Searches for MSEs by other names i.e., “management procedure approach” and “management procedure evaluation” returned 9 and 1 results respectively, none of which document a MSE.
 The number of MSEs published each year is trending higher (Figure \@ref(fig:YearHistogram)), with the ICES Journal of Marine Science and Fisheries Research as the leading publication outlets (Figure \@ref(fig:PublicationHistogram)).
@@
 ```
-
-We reviewed 16 climate change MSE articles (Appendix B).
-Twelve of these articles came from our original January 8th, 2019 Web of Science search.
-To give as full a picture of the climate change MSE articles to date as possible we also reviewed the 4 climate change MSE articles published in 2019.
 
@@
 
+We reviewed 16 climate change MSE articles (Appendix A).
+Twelve of these articles came from our original January 8th, 2019 Web of Science search.
+To give as full a picture of the climate change MSE articles to date as possible we also reviewed the 4 climate change MSE articles published in 2019.
+These MSEs have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).
+
+```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=6, out.width="100%"}
+MSE.map +
+  theme(text = element_text(size=14))
+```
+
 ## Reviewing MSE documentation 
@@
 
-We present the review methodology and results in four groups: consequences - which focuses on what was considered during the prediction stage, objectives - which focuses on what categories of objectives were considered, stages - which focuses on whether and how the stages of the process were completed, and participants - which focuses on how participation in the process was structured.
-These results were entered into a relational database for storage and evaluation, and is available via our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
+We present the review methodology and results in four groups: stages - which focuses on whether and how the stages of the process were completed, objectives - which focuses on what categories of objectives were considered, participants - which focuses on how participation in the process was structured, and consequences - which focuses on what was considered during the prediction stage.
+These results were entered into a relational database for storage and evaluation, and are available via our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
 
-The consequence group addresses what question the MSE was designed to address, and how it was addressed.
-We report the drivers included as the primary factors influencing the predictions produced by the MSE model, listing each driver a MSE included.
-We also report how the consequence predictions were made, with possible values of simulation modeling, dynamic programming, expert elicitation, and mental models, or unknown when it is unclear what methodology was used to predict the status of the system for management evaluation.
-
-In the objectives group we address what the concerns, or goals of the MSE were.
-We record each objective considered, and report the type and category of objectives included in each MSE.
-Natural resource management problems typically include four types of objectives: conservation, resource use, i.e., yield in the fisheries context, as well as economic, and social objectives [@runge2013].
-We sort objectives into four types: strategic objectives - which are concerned with the overall mission of the system in which the decision evaluated by the MSE falls, process objectives - which are concerned with how the management decision addressed by the MSE would be made, fundamental objectives - which address the goals of the management decision, and means objectives - which address means to achieving the fundamental objectives of the management decision.
-
-In the stages group we address how the MSE stages were completed.
+In the stages group we identify how the MSE stages were completed.
 We report whether the stages of the structured decision making framework, i.e., problem, objectives, alternatives, tradeoffs, were explicitly completed and the methods used to complete them were documented.
@@
 
-In the participant group we address how participation in the MSE process was structured and who participated.
-We report who lead the MSE process, who participated in it, and if the roles of participants were established and documented.
+In the objectives group we identify the concerns, or goals of each MSE.
+We record each objective considered, and report the type and category of objectives included in each MSE.
+Natural resource management problems typically include four categories of objectives: conservation, resource use, i.e., yield in the fisheries context, as well as economic, and social objectives [@runge2013].
+At times multiple objectives categories are combined into a single metric, which we considered a fifth objective category, namely utility objectives.
+We sort objectives into four types: 1) strategic objectives - which are related to the overall mission of management in the system where the decision evaluated by the MSE occurs, e.g., objectives reaching beyond the scope of the problem addressed, 2) process objectives - which are concerned with how the management decision addressed by the MSE would be made, e.g., achieving stakeholder buy-in to project outcomes, 3) fundamental objectives - which address the goals of the management decision, e.g., maximizing revenue from the fishery, and 4) means objectives - which address means to achieving the fundamental objectives of the management decision, e.g.
+maximizing catch per unit effort, which is a means to maximizing revenue.
+
+In the participant group we describe how participation in each MSE process was structured and who participated.
+We report who led the MSE process, who participated in it, and if the roles of participants were established and documented.
 In many cases the process used to elicit objectives and alternatives was not explicit.
@@
 We record and report who those participants were, referring to them as subjective participants in the objectives or alternatives process.
-In other cases where there was not enough information to glean who participated we report the participants as unknown.
+In cases where there was not enough information to glean who participated we report the participants as unknown.
 
+The consequence group identifies what question each MSE was designed to address, and what methods were employed.
+We report the drivers – the characteristics of the model that drive the behavior of and determine the predictions produced by the MSE model – listing each driver a given MSE included.
+We also report whether consequence predictions were made, using simulation modeling, dynamic programming, expert elicitation, mental models, or unknown prediction methods when it was unclear what methodology was used to predict the status of the system for management evaluation.
 We collected additional information about the MSEs we reviewed beyond what is reported here.
-For a full description of the information recorded in the relational database see the supplemental information (Appendix A), and for the full data set and results see our shiny application (https://jonathancummings.shinyapps.io/MSEreview/).
+For a full description of the information recorded in the relational database see our appendix (Appendix B), and for the full data set and results see our web-based shiny application in R  (https://jonathancummings.shinyapps.io/MSEreview/).
 
 # Results 
-## What questions and management objectives do MSEs address? 
+## Are structured decision making steps explicit in MSEs? 
+The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework (Figure \@ref(fig:SDMfigure)) -- were completed (Table \@ref(tab:explicit), Random Sample).
+Apart from the consequence stage, which is the primary focus of MSE publication and therefore was always addressed, the most documented stage was the alternative stage with 6 MSEs documenting this stage of the process.
+The MSE process, problem, and objectives methodology were each documented in 5 of the MSEs.
+Three of the 30 MSEs explicitly documented a trade-off analysis and 2 of the 30 documented a management decision.
+One MSE documented that the results of the MSE were adopted.
+Climate change MSEs were more likely to explicitly document the trade-off analysis stage, equally likely to document the alternatives stages, but less likely to document the other stages in the decision process (Table \@ref(tab:explicit), Climate Change).
+
+```{r explicit}
+freq.data_pub %>%
+  select(-Analysis) %>% 
+  mutate("Percent"=round(Percent,0)) %>% 
+  select(Explicit,Percent,Number) %>% 
+  kable(caption = "Prevalence of explicit documentation of stages in the MSE process") %>% 
+  kable_styling()
+```
+
+## What management objectives do MSEs address? 
 We found that uncertainty was the main driver of the system state included in management strategy evaluations, with uncertainty included in 21 of the 30 evaluations we reviewed in our random sample (Table \@ref(tab:drivers)).
@@
 Yield and economic objectives were present in about three quarters of the MSEs, while social objectives were only included in 2 of the sampled MSEs.
+The climate change MSE objectives differed, with more frequent inclusion of yield objectives, but less frequent consideration of economic objectives and no consideration of social objectives.
 
@@
 <br>
-No strategic or process objectives were documented in the MSEs.
+No strategic or process objectives were documented in the sampled MSEs.
 There were roughly the same number of fundamental and means objectives included in the MSEs (Table \@ref(tab:objectiveTypes)).
 An average of 6 objectives were evaluated by each management strategy evaluation.
+On average climate change MSEs included the same number of objectives per MSE as the random sample.
 ```{r, objectiveTypes}
@@
 
-## Are structured decision making steps explicit in MSEs? 
-The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Table \@ref(tab:explicit)).
-Apart from the consequence stage, which is the primary focus of MSE publication and therefore was always addressed, the most documented stage was the alternative stage with 6 MSEs documenting this stage of the process.
-The MSE process, problem, and objectives methodology were each documented in 5 of the MSEs.
-Three of the 30 MSEs explicitly documented a trade-off analysis and 2 of the 30 documented a management decision.
-One MSE documented that the results of the MSE were adopted.
-
-```{r explicit}
-freq.data_pub %>%
-  select(-Analysis) %>% 
-  mutate("Percent"=round(Percent,0)) %>% 
-  select(Explicit,Percent,Number) %>% 
-  kable(caption = "Prevalence of explicit documentation of stages in the MSE process") %>% 
-  kable_styling()
-```
-
 ## Who is involved in MSEs? 
-MSEs were rarely explicit about the roles participants played in MSE processes, as only one MSE documented the roles participants played in the process (Table \@ref(tab:roles)).
+MSEs were rarely explicit about the roles participants played in MSE processes, with only one MSE documenting these (Table \@ref(tab:roles)).
 None of the MSEs noted whether or not public meetings were held in which the public was invited to participate.
@@
 
-The primary participants and leaders of the MSE processes were scientists, while some MSE process were led by governments and management agencies (Figure \@ref(fig:participants) Process).
-Scientists were explicitly the leaders -- or the sole participants and therefore seemingly the leaders -- in 73 percent of the MSE processes, while 27 percent of processes lacked enough documentation of participants and leadership such that we were unable to identify a presumed leader of the process.
+Inferring from scientifically authored publications in the absence of explicit statements indicating otherwise, the primary participants and leaders of the MSE processes were scientists, while some MSE processes were led by governments and management agencies (Figure \@ref(fig:participants) Process).
+Scientists were explicitly the leaders – or the sole participants and therefore seemingly the primary participants and leaders – in 73 percent of the MSE processes, while 27 percent of processes lacked enough documentation of participants and leadership such that we were unable to identify a presumed leader of the process.
 
-Scientists were participants in nearly all of the reviewed MSE processes (Figure \@ref(fig:participants) Participants).
+Scientists were participants in all of the reviewed MSE processes (Figure \@ref(fig:participants) Participants).
 Other stakeholders in MSE processes were less frequent participants.
-Members of the fishery participated in 30 percent, management and government representatives in 23 percent, representatives of independent institutions in 7 percent, and the members of the public in 3 percent of reviewed MSEs.
-Decision makers -- those responsible for selecting the management plan to implement -- were participants in 13 percent of MSEs.
-Experts -- a possible source of data or predictions -- participated in 3 percent of the MSEs reviewed.
-Participants involved to the assist the MSE process itself, i.e., facilitators and decision analysts, participated in 10 and 3 percent of MSE processes respectively.
+Members of the fishing community participated in 30 percent, management and government representatives in 23 percent, representatives of independent institutions in 7 percent, and the members of the public in 3 percent of reviewed MSEs.
+Decision makers – those responsible for selecting the management plan to implement – were participants in 13 percent of MSEs.
+Experts – a possible source of data or predictions – participated in 3 percent of the MSEs reviewed.
+Participants involved to assist the MSE process itself, i.e., facilitators and decision analysts, participated in 10 and 3 percent of MSE processes respectively.
 
@@
 Management, government, or fishery participants explicitly provided objectives and alternatives in about 10 percent of the MSE processes (Figure \@ref(fig:participants) Explicit).
-However, in most cases the source of the objectives and alternatives was not explicit.
-The source of objectives and alternatives was not explicit in more than 80% of the MSE processes, 25 of 30 objective sources were not explicit and 24 of 30 alternative sources.
-In those subjective cases scientists -- being the only or primary participants -- seemingly selected these components most frequently (Figure \@ref(fig:participants) Subjective).
+However, in most cases the source of the objectives and alternatives was not explicitly documented.
+The source of objectives and alternatives was not explicit in more than 80% of the MSE processes, 25 of 30 MSE processes used objectives without explicitly documenting the source for the objectives, i.e., who provided them, and 24 of 30 MSEs did not explicitly document who provided the alternatives.
+In those subjective cases scientists – being the only or primary participants – seemingly selected these components, making them the most frequent providers of objectives and alternatives in MSE processes (Figure \@ref(fig:participants) Subjective).
 
@@
 
-# Climate Change in MSE (As *Box Insert*) 
-Fisheries management has mostly focused on fishing impacts with ecosystem status viewed as a background constant.
-The increasing rate of climate change and its influence on fisheries is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.
-
-Here we review published MSEs the include climate change as a driver of the fisheries model using the stages in a SDM process to highlight unique aspects of these MSEs.
-We filtered the original set of MSEs published by as of December 31st, 2019 searching for the term "climate change" within the set that resulted from searching for "management strategy evaluation".
-This resulted in our reviewing 16 published MSEs that included climate change as a driver.
-We reviewed the climate change MSEs using the same methodology described above.
-These MSEs have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).
-
-```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=6, out.width="100%"}
-MSE.map +
-  theme(text = element_text(size=14))
-```
-
-As with the random sample of reviewed MSEs, uncertainty was the most common driver, apart from climate change, included climate change MSEs (Table \@ref(tab:driversCC)), while other biological drivers were also included in some of the climate change MSEs.
-Overall climate change MSEs considered fewer drivers on average, 1.4, than the random sample of MSEs, which included 1.8.
+## What questions do MSEs address? 
+All sampled MSEs used simulation modeling for consequence prediction.
+We found that uncertainty was the main driver of the system state included in management strategy evaluations, with uncertainty included in 21 of the 30 evaluations we reviewed (Table \@ref(tab:driversCC)).
+Monitoring methodology, how and what data was collected, was the next most prevalent driver in MSE models, but a variety of drivers were considered across the random sample of MSEs.
+As with the random sample of reviewed MSEs, uncertainty was the most common driver, apart from climate change, included in climate change MSEs, while other biological drivers were also included in some of the climate change MSEs.
+Overall, climate change MSEs considered fewer drivers on average (1.4) than the random sample of MSEs (1.8).
 
@@
 
-Climate change MSEs averaged the same number of objectives per MSE.
-The categories of the objectives differed however, with more frequent inclusion of yield objectives, but less frequent consideration of economic objectives and no consideration of social objectives (Table \@ref(tab:CCobjectiveCategories)) in the climate change MSEs.
-```{r, CCPerMSE}
-# color_Pub <- which(per.MSE$Analysis=='Random Sample')
-# color_CC <- which(per.MSE$Analysis=='Climate Change')
-# 
-# per.MSE %>%
-#   kable(caption = "Number of objectives considered per management strategy evaluation") %>%
-#   row_spec(color_CC, color = "#66CCCC", background = "white") %>%
-#   row_spec(color_Pub, color = "#006666", background = "white")%>% 
-#   kable_styling()
-```
+# Discussion 
+The primary picture painted by MSE documentation in the scientific literature is one of scientist driven simulation studies rather than management driven decision processes.
+Documentation about the modeling conducted in a MSE was explicit, and all MSEs included well documented simulation studies supporting decision making based upon a clear prediction of the future state of the selected performance metrics.
+Documentation about the MSE process itself was often limited, however.
 
-```{r, CCobjectiveCategories}
-objcat.data_CC<-objcat.data %>%
-  filter(`Objective Category`!="Utility") %>% 
-  add_column(order=c(1,2,5,6,7,3,4)) %>% 
-  arrange(order)
+We found that management strategy evaluations typically assess the consequences of uncertainty for achieving management goals, one of the recommended best practices of MSEs [@Punt], but other factors influencing the consequences of management strategies get more sporadic treatment.
+Data limitation is a frequent concern in fisheries management and contributes to the degree of uncertainty present in fisheries management.
+Next to accounting for uncertainty, accessing the influence of data quality and availability, i.e., monitoring methodology was the most prevalent driver included in MSEs.
+This likely points to a desire for the design and use of cost-effective fisheries monitoring plans, suggesting future MSEs may be able to learn from the set of studies identified here that include monitoring methodology as a driver.
+Additionally, using value of information analyses – which directly determine the expected benefits of additional research – could expand upon the benefits of MSE as a tool to direct scientific resources [@runge_2011; @moore_2012].
+While discussions of fisheries research and management seem to be moving toward ecosystem effects, spatial models, and multiple species models, these characteristics were rarely considered in the set of MSEs we reviewed.
+Including climate change in a MSE appears to reduce the complexity of other stages in the MSE process, in particular those stages that use input from additional participants.
+On average climate change MSEs consider fewer drivers, consider less diverse objectives, and explicitly document fewer stages of the MSE process.
+If the MSE process is time, effort, or participant limited, the inclusion of climate change may be utilizing time, effort, or outside participants to focus on climate change rather than other aspects of the MSE process.
 
-color_Pub <- which(objcat.data_CC$Analysis=='Random Sample')
-color_CC <- which(objcat.data_CC$Analysis=='Climate Change')
-
-objcat.data_CC %>%
-  select(-order) %>% 
-  kable(caption = "Objective categories considered in management strategy evaluations") %>%
-  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
-  row_spec(color_Pub, color = "#006666", background = "white") %>% 
-  kable_styling()
-```
-
-Climate change MSEs were more likely to explicitly document the trade-off analysis stage, equally likely to document the alternatives stages, but were less likely to document the other stages in the decision process (Table \@ref(tab:CCstages)).
-```{r, CCstages}
-freq.data_CC %>%
-  mutate("Percent"=round(Percent,0)) %>% 
-  select(Explicit, Percent, Number, -Analysis) %>% 
-  kable(caption = "Prevalence of explicit documentation of stages in climate change MSE processes") %>%
-  kable_styling()
-```
-
-Including climate change in a MSE appears to reduce the complexity of other stages in the MSE process, in particular those stages that utilize input from additional participants.
-On average climate change MSEs consider fewer drivers, consider less diverse objectives, and explicitly document fewer stages of the MSE process.
-If the MSE process is time, effort, or participant limited the inclusion of climate change may be utilizing time, effort, or outside participants to focus on climate change rather than other aspects of the MSE process.
-
-# Discussion 
-The primary picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes.
-While documentation about the modeling conducted in a MSE was explicit, documentation about the MSE process itself was often limited.
-
-We found that management strategy evaluations do typically assess the consequences of uncertainty for achieving management goals, but other factors influencing the consequences of management strategies get more sporadic treatment.
-Data limitation is a frequent concern in fisheries management, and contributes to the degree of uncertainty present in fisheries management.
-Next to accounting for uncertainty, accessing the influence of data quality and availability, i.e.,  monitoring methodology was the most prevalent driver included in MSEs.
-This likely points to a desire for the design and use of cost-effective fisheries monitoring plans, suggesting future MSEs may be able to learn from the set including monitoring methodology as a driver.
-Additionally, the use of value of information analyses could expand upon the benefits of MSE as a tool to direct scientific resources [@runge_2011; @moore_2012].
-While discussions of fisheries research and management seem to be moving toward ecosystem effects, spatial models, and multiple species models, these characteristics were rarely considered in the set of MSEs we reviewed.
-
-Managing natural resources entails achieving the objectives of people interacting with a natural resource, so decision processes are likely to produce more desirable results when they include objectives that reflect stakeholders' and society's values.
+Managing natural resources entails achieving the objectives of people interacting with a natural resource, so decision processes are likely to produce more desirable results when they include objectives that reflect stakeholders’ and society’s values.
 These objectives are likely to include conservation, yield, economic, and social concerns.
-As noted, MSEs are described as a tool to evaluate what management strategy best achieves objectives, however the MSEs we reviewed don't cover all categories of objectives.
+As noted, MSEs are described as a tool to evaluate which management strategies best achieve objectives, however the MSEs we reviewed don’t cover all categories of objectives.
 Yield and economic objectives were not included in about one quarter of MSEs, and broader social objectives were absent from all but 2 of the 30 evaluations.
@@
 The lack of participation by stakeholders may result in less frequent inclusion of social objectives, as well as yield and economic objectives specific to a particular fishery.
-Another factor may be that MSE models are likely constructed mainly by ecosystem modelers, who are likely to be more comfortable modeling the consequences of conservation and yield objectives than economic or social objectives.
+Another factor may be that MSE models are likely constructed mainly by ecological modelers, who are likely to be more comfortable modeling the consequences of conservation and yield objectives than economic or social objectives.
 Given that scientists seem to be providing the objectives in many MSE processes, they may be selecting objectives that they are more comfortable evaluating.
 There also appears to be some degree of borrowing occurring in which scientists use a set of generic fishery objectives when conducting a MSE.
-Perhaps because their were few social objectives included in these MSEs, there was little inclusion of experts to aid in the prediction of consequences, or consequence prediction techniques other than simulation modeling.
+Perhaps because there were few social objectives included in these MSEs, there was little inclusion of experts to aid in the prediction of consequences, or consequence prediction techniques other than simulation modeling.
 
-Participants that would be likely to expand the set of objectives considered in a MSE rarely participated in the MSEs we reviewed.
+Participants that would likely expand the set of objectives considered in a MSE rarely participated in the MSEs we reviewed.
 That is, facilitators and decision analysts were rarely participants in MSE processes.
-The inclusion of facilitators in decision processes aid information gain and conflict resolution, increase trust between participants, and lead to more beneficial environmental and social outcomes [@Vente].
+The inclusion of facilitators in decision processes aid information gain and conflict resolution, increase trust between participants, and lead to more beneficial environmental and social outcomes [@Vente; @Feeney2019].
 The additional structure and focus on the decision process would likely result in more explicit documentation when facilitators and decision analysts participate in MSE, as resulted in some of the MSEs reviewed here [@williams2011; @smith2013; @kolody2008].
@@
 
-The MSE process may be aided by utilizing the techniques from the structured decision making approach to decision analysis, such as expanding the focus on problem framing at the outset of MSEs, eliciting fully representative sets of objectives, and utilizing multiple attribute utility theory to evaluate tradeoffs between objectives (Cummings, in prep).
+The MSE process may be aided by utilizing the techniques from the structured decision making approach to decision analysis, such as expanding the focus on problem framing at the outset of MSEs, eliciting fully representative sets of objectives, and utilizing multiple attribute utility theory to evaluate tradeoffs between objectives [@Cummings_inprep], and more closely adhering to the best practices for MSE [@Punt].
 Utilizing the structured decision making approach also eases the documentation process by breaking decision making into more discrete stages.
-However, there may be practical considerations, such as time or budgetary constraints that limit how complex the stages of a MSE are, and the degree of documentation that occurs.
+However, there may be practical considerations, such as time, budgetary constraints, or available participants that limit how complex the stages of a MSE are, and the degree or form of documentation that occurs.
 Where the expertise and budget are available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.
+
+We did not include grey literature in our evaluation.
+It is possible that more of the MSE process is documented in grey literature publications, and less is documented in the scientific publications because they are primarily produced by scientific authors.
+However, we are aware of only one grey literature publication focused on aspects of an MSE process, namely grey literature documentation of the Atlantic herring MSE process.
+We also note that grey literature publications are less likely to reach an audience of MSE practitioners and perhaps should be included as an appendix or supplement to MSE publications where available.
 
@@
 Different decision makers can rationally look at the same prediction of consequences and select a different preferred alternative if they treat the importance of the objectives differently.
-Therefore, MSE documentation often is not fully supporting replication or learning due to the frequent absence of trade-off and decision documentation.
+Therefore, MSE documentation in the published, peer-reviewed literature often is not fully supporting replication or learning due to the frequent absence of trade-off and decision documentation.
 
 ## Explore and expand upon this review
-To aid learning from the documentation that has occurred, and provide ready access to future MSE documentation we produced a shiny application [@R-Shiny] to enable an interactive means to explore the MSE literature (<https://jonathancummings.shinyapps.io/MSEreview/>).
+To aid learning from the documentation that has occurred, and to provide ready access to future MSE documentation we produced a shiny application [@R-Shiny] to enable an interactive means to explore the MSE literature (https://jonathancummings.shinyapps.io/MSEreview/).
 The MSE review shiny application enables users to interactively explore the results presented in this article.
@@
 Users may also filter by driver, location, species, or other characteristics of MSEs to explore a set of MSEs of interest and relevance to aid learning in the MSE practitioner community.
-We hope that as MSE are conducted they are entered here so that MSE practitioners can further learn from each other in the future.
+We hope that as MSEs are conducted they are entered here so that MSE practitioners can further learn from each other in the future.
 
